# Get Started {#sec-start}

## Installation

`sconjoint` runs on the `torch` deep-learning backend, which bundles its
own copy of `libtorch` (downloaded on first use). Install `torch` first
and confirm its native library loads:

```{r install-torch, eval=FALSE}
install.packages("torch")
torch::install_torch()        # downloads libtorch the first time
torch::torch_is_installed()   # should return TRUE
```

Then install `sconjoint` from CRAN. It has been submitted; until it is
accepted this line stays commented out --- install a development version
instead (see @sec-versions):

```{r install-cran, eval=FALSE}
# install.packages("sconjoint")
```

On CPU-only machines this is all you need. On a CUDA-enabled GPU, follow
<https://torch.mlverse.org> to install the CUDA build of `libtorch`;
`scfit(device = "cuda")` then uses it automatically. (The bit-exact
determinism guarantee --- identical `sc_fit` output regardless of
`n_cores` --- holds only on CPU.)

## Versions {#sec-versions}

Until the CRAN release, install from GitHub. The development branches
carry features ahead of CRAN:

| Source | Version | What it adds |
|:--------------------|:------------|:------------------------------------------------------------|
| CRAN | 0.2.1 | The released version (default), pending acceptance |
| GitHub `@dev` | 0.2.1 | Development version; currently matches the CRAN submission |
| GitHub `@feat/torch-threads` | 0.2.1.9000 | Newest in-progress work: the `torch_threads` Windows-stability control and the attribute-interaction extension, pending merge to `dev` after the release |

Install a specific source with the `@branch` suffix:

```{r install-github, eval=FALSE}
# install.packages("remotes")
remotes::install_github("xuyiqing/sconjoint@dev")                 # development version
remotes::install_github("xuyiqing/sconjoint@feat/torch-threads")  # newest features
```

Check which version is installed on this computer:

```{r check-version}
packageVersion("sconjoint")
```

Once `sconjoint` is on CRAN, `install.packages("sconjoint")` gives the
latest release and `remotes::install_version("sconjoint", "0.2.1")` pins a
specific one. Installing `sconjoint` pulls in `ggplot2`, `ggridges`, and
related plotting packages automatically; no extra setup is needed for the
visualization methods in the Plot Options chapter.

::: {.callout-warning}
## Troubleshooting: the R session aborts when calling `scfit()`

A session *abort* (the R process exits, rather than throwing an R error)
almost always comes from the native `libtorch` backend, not from
`sconjoint` itself. It has been reported on some Windows systems while the
same code runs fine on macOS. Things to try, in order:

1. **Confirm `torch` works in isolation.** In a fresh session, run
   `library(torch); x <- torch_randn(3, 3); x %*% x`. If *that* aborts, the
   problem is the `torch` / `libtorch` installation or your environment, not
   `sconjoint`: reinstall with `torch::install_torch()`, and check whether
   security or endpoint-protection software is blocking the `libtorch` DLLs
   (a common cause on managed corporate Windows machines).

2. **Cap torch's threads with `torch_threads = 1`.** torch's default
   multi-threaded CPU backend aborts the session on some Windows
   configurations, and running it single-threaded often avoids it. The
   `torch_threads` argument to `scfit()` is the targeted fix for this
   ([issue #7](https://github.com/xuyiqing/sconjoint/issues/7)). It
   currently lives on the `feat/torch-threads` branch, pending merge to
   `dev` after the next release, so install that branch first:

   ```r
   remotes::install_github("xuyiqing/sconjoint@feat/torch-threads")
   library(sconjoint)
   fit <- scfit(formula, data, ..., torch_threads = 1)
   ```

   `torch_threads = 1` changes only the threading, not the results, and
   torch honors the thread count only on the *first* such call of a
   session --- so set it before running any other torch work.

3. **Try plain `R`** (`Rterm` or the console) rather than an IDE, to rule
   out an IDE-specific interaction.
:::
