# Get Started {#sec-start}

`sconjoint` depends on the `torch` R package for the deep-learning
backend. `torch` in turn bundles its own copy of `libtorch` which is
downloaded on first use. Before installing `sconjoint` itself,
confirm that `torch` can load its native library:

```r
install.packages("torch")
torch::install_torch()  # downloads libtorch the first time it is run
torch::torch_is_installed()
```

On CPU-only machines the default installation is sufficient. On
machines with a CUDA-enabled GPU, follow the instructions at
<https://torch.mlverse.org> to install the CUDA build of `libtorch`;
`sconjoint` will then pick up GPU acceleration automatically when you
pass `device = "cuda"` to `scfit()`. Note that the package's
bit-exact determinism guarantee --- identical `sc_fit` output regardless
of `n_cores` --- applies only on CPU; on GPU, cuDNN introduces
nondeterminism that the package cannot fully eliminate.

Then install `sconjoint`. The table summarizes the available sources; pick
one and run the matching code.

| Source | Version | Notes |
|--------|---------|-------|
| CRAN | 0.2.1 | Submitted; pending acceptance --- use a GitHub source until then |
| GitHub `@dev` | 0.2.1 | Development version (currently matches the CRAN submission) |
| GitHub `@feat/torch-threads` | 0.2.1.9000 | Newest in-progress features: the `torch_threads` Windows-stability control and the attribute-interaction extension, pending merge to `dev` after the release |

Once `sconjoint` is on CRAN you can install the released version directly.
It has been submitted; this is commented out until it is accepted:

```{r install-cran, eval=FALSE}
# install.packages("sconjoint")
```

For the development version, install from the `dev` branch on GitHub:

```{r install-dev, eval=FALSE}
# install.packages("remotes")
remotes::install_github("xuyiqing/sconjoint@dev")
```

After installing, check the version to confirm you have the features you
expect:

```{r check-version}
packageVersion("sconjoint")
```

The package pulls in `ggplot2`, `ggridges`, and related plotting
infrastructure as Imports; no extra setup is required for the
visualization methods described in the Plot Options chapter.

::: {.callout-warning}
## Troubleshooting: the R session aborts when calling `scfit()`

A session *abort* (the R process exits, rather than throwing an R error)
almost always comes from the native `libtorch` backend, not from
`sconjoint` itself. It has been reported on some Windows systems while the
same code runs fine on macOS. Things to try, in order:

1. **Confirm `torch` works in isolation.** In a fresh session, run
   `library(torch); x <- torch_randn(3, 3); x %*% x`. If *that* aborts, the
   problem is the `torch` / `libtorch` installation or your environment, not
   `sconjoint`: reinstall with `torch::install_torch()`, and check whether
   security or endpoint-protection software is blocking the `libtorch` DLLs
   (a common cause on managed corporate Windows machines).

2. **Cap torch's threads with `torch_threads = 1`.** torch's default
   multi-threaded CPU backend aborts the session on some Windows
   configurations, and running it single-threaded often avoids it. The
   `torch_threads` argument to `scfit()` is the targeted fix for this
   ([issue #7](https://github.com/xuyiqing/sconjoint/issues/7)). It
   currently lives on the `feat/torch-threads` branch, pending merge to
   `dev` after the next release, so install that branch first:

   ```r
   remotes::install_github("xuyiqing/sconjoint@feat/torch-threads")
   library(sconjoint)
   fit <- scfit(formula, data, ..., torch_threads = 1)
   ```

   `torch_threads = 1` changes only the threading, not the results, and
   torch honors the thread count only on the *first* such call of a
   session --- so set it before running any other torch work.

3. **Try plain `R`** (`Rterm` or the console) rather than an IDE, to rule
   out an IDE-specific interaction.
:::
