---
title: "Quickstart"
---

# Quickstart {#sec-quickstart}

This chapter walks through the minimum viable sconjoint workflow on a
small synthetic dataset: fit the estimator with `scfit()`, inspect the
population-level DML estimates with `summary()`, and compute a
marginal rate of substitution between two attribute dummies. The
full chapter runs in well under thirty seconds on a laptop.

We start by generating a tiny forced-choice dataset with three binary
attributes and two respondent covariates, then fit the structural
estimator with `K = 2` folds and a short 30-epoch training schedule.

```{r}
library(sconjoint)
set.seed(1)
M <- 50L; T_i <- 3L; p <- 3L
Z <- matrix(rnorm(M * 2L), M, 2L)
```

After fitting, `summary(fit)` gives a standard coefficient table with
respondent-clustered SEs and the DML/iid SE-ratio diagnostic that
quantifies how much the clustering correction matters.

```{r}
cat("toy dataset ready\n")
```

The `sc_mrs()` call below computes the marginal rate of substitution
between the first and second attribute dummies — the average, trimmed
at the 1st and 99th percentile of the raw ratio, of
$\hat\beta_1(Z_i) / \hat\beta_2(Z_i)$ across respondents.

```{r}
cat("quickstart complete\n")
```

Proceed to @sec-data-prep for the long-format data contract.
