# Sanity Check {#sec-sanity}

```{r}
#| label: setup-sim
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(42)
torch::torch_manual_seed(42)
library(sconjoint)
library(ggplot2)
```

Before applying the structural estimator to real data, this chapter
runs the full workflow on a synthetic conjoint experiment with
**known ground truth**. Because we control the data-generating process,
every quantity the package produces can be checked against its
analytically correct value. This builds confidence that the estimator,
the DML correction, and the downstream structural quantities all work
as intended.

## Data-generating process

We simulate $M = 100$ respondents, $T = 4$ tasks per respondent, and
two profiles per task (800 rows). Three binary attributes
$x_1, x_2, x_3$ are drawn independently as Bernoulli(0.5). Two
continuous moderators $z_1, z_2 \sim N(0,1)$ determine each
respondent's preference vector through a simple linear map:

$$
\beta_1(\mathbf Z_i) = 0.5 + 0.3\,z_{1i}, \quad
\beta_2(\mathbf Z_i) = -0.8 + 0.2\,z_{2i}, \quad
\beta_3(\mathbf Z_i) = 0.3.
$$

Key features of this DGP:

- $\beta_1$ and $\beta_2$ depend on respondent covariates
  (**heterogeneous**); $\beta_3$ is constant (**homogeneous**).
- $\beta_1 > 0$ for most respondents (positive preference);
  $\beta_2 < 0$ for most respondents (negative preference).
- The true population-average AMCE is $E[\beta_1] = 0.5$,
  $E[\beta_2] = -0.8$, $E[\beta_3] = 0.3$.

```{r}
#| label: sim-dgp
M <- 100L
TT <- 4L
set.seed(42)

## Respondent-level moderators
z1 <- rnorm(M)
z2 <- rnorm(M)

## True beta(Z) for each respondent (M x 3)
beta_true <- cbind(
  x1 = 0.5 + 0.3 * z1,
  x2 = -0.8 + 0.2 * z2,
  x3 = rep(0.3, M)
)

## Build long-format data frame
rows <- vector("list", M * TT * 2L)
idx <- 0L
for (i in seq_len(M)) {
  for (tt in seq_len(TT)) {
    xA <- rbinom(3, 1, 0.5)
    xB <- rbinom(3, 1, 0.5)
    deltaX <- xA - xB
    V <- sum(deltaX * beta_true[i, ])
    choiceA <- rbinom(1, 1, plogis(V))
    idx <- idx + 1L
    rows[[idx]] <- data.frame(
      respondent = i, task = tt, profile = 1L, choice = choiceA,
      x1 = xA[1], x2 = xA[2], x3 = xA[3], z1 = z1[i], z2 = z2[i]
    )
    idx <- idx + 1L
    rows[[idx]] <- data.frame(
      respondent = i, task = tt, profile = 2L, choice = 1L - choiceA,
      x1 = xB[1], x2 = xB[2], x3 = xB[3], z1 = z1[i], z2 = z2[i]
    )
  }
}
sim_data <- do.call(rbind, rows)
dim(sim_data)
head(sim_data, 6)
```

## Fitting

```{r}
#| label: sim-fit
fit_sim <- scfit(
  choice ~ x1 + x2 + x3 | z1 + z2,
  data       = sim_data,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 3L,
  n_epochs   = 100L,
  seed       = 42
)
summary(fit_sim)
```

## Population-average recovery

The DML point estimates $\hat\theta$ should approximate the true
population means $E[\beta_1] = 0.5$, $E[\beta_2] = -0.8$,
$E[\beta_3] = 0.3$.

```{r}
#| label: sim-theta
theta_hat <- coef(fit_sim)
theta_true <- c(x1 = 0.5, x2 = -0.8, x3 = 0.3)
data.frame(
  attribute  = names(theta_hat),
  true       = theta_true,
  estimated  = round(unname(theta_hat), 3),
  row.names  = NULL
)
```

Signs and rough magnitudes should line up. Exact recovery is not
expected with only 100 respondents and 4 tasks each.

## Individual-level recovery

The structural payoff is the per-respondent $\hat\beta(\mathbf Z_i)$
matrix. We compare estimated values to the true DGP values.

```{r}
#| label: sim-beta-recovery
beta_hat_all <- predict(fit_sim)

## De-duplicate to respondent level
first_row <- !duplicated(fit_sim$respondent_id)
beta_hat <- beta_hat_all[first_row, ]

## Correlation between true and estimated beta
data.frame(
  attribute   = c("x1", "x2"),
  correlation = round(c(
    cor(beta_true[, 1], beta_hat[, 1]),
    cor(beta_true[, 2], beta_hat[, 2])
  ), 3),
  row.names = NULL
)
```

Attribute $x_3$ has constant $\beta_3 = 0.3$ (no heterogeneity across
respondents), so per-respondent correlation is not informative for it.

```{r}
#| label: sim-scatter
#| fig-height: 5
df_compare <- data.frame(
  true      = beta_true[, 1],
  estimated = beta_hat[, 1]
)

ggplot(df_compare, aes(x = true, y = estimated)) +
  geom_point(alpha = 0.5, color = "#2166AC", size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "#E41A1C", linewidth = 0.8) +
  labs(x = expression("True " * beta[1](Z)),
       y = expression("Estimated " * beta[1](Z)),
       title = "Individual-level recovery: attribute x1") +
  theme_minimal(base_size = 12)
```

Points clustering around the 45-degree line confirm that the estimator
recovers individual-level heterogeneity. Dispersion around the line
reflects the inherent noise from only $M = 100$ respondents with
$T = 4$ tasks.

## Direction and intensity

The direction--intensity decomposition should reflect the DGP:
$x_1$ has strong positive direction, $x_2$ has strong negative
direction, $x_3$ has moderate positive direction.

```{r}
#| label: sim-di
sc_direction_intensity(fit_sim)
```

## Fraction preferring

We can verify this against the DGP analytically. For $x_1$:
$\Pr(\beta_1 > 0) = \Pr(0.5 + 0.3 z_1 > 0) = \Pr(z_1 > -5/3)
\approx 0.95$. For $x_2$:
$\Pr(\beta_2 > 0) = \Pr(-0.8 + 0.2 z_2 > 0) = \Pr(z_2 > 4)
\approx 0$. For $x_3$: $\beta_3 = 0.3 > 0$ always, so the fraction
is 1.

```{r}
#| label: sim-fraction
frac_sim <- sc_fraction_preferring(fit_sim)

## Compare to DGP truth
true_frac <- c(
  x1 = mean(beta_true[, 1] > 0),
  x2 = mean(beta_true[, 2] > 0),
  x3 = mean(beta_true[, 3] > 0)
)

data.frame(
  attribute     = names(true_frac),
  true_fraction = true_frac,
  estimated     = round(frac_sim$estimate$frac_positive, 3),
  row.names     = NULL
)
```

## Heterogeneity test

The test should flag $x_1$ and $x_2$ as **significant** (their
$\beta$ varies with $Z$) and $x_3$ as **non-significant** (constant
across respondents). This is the key validation: the test correctly
distinguishes structural heterogeneity from homogeneity.

```{r}
#| label: sim-het
het_sim <- sc_heterogeneity_test(fit_sim, adjust = "bh")
het_sim$estimate[, c("dummy_name", "var_beta", "p_adjusted", "sig")]
```

## Subgroup analysis

By construction, respondents with $z_1 > 0$ have $\beta_1 > 0.5$
while those with $z_1 < 0$ have $\beta_1 < 0.5$. The subgroup
comparison should recover this split.

```{r}
#| label: sim-subgroup
z1_col <- fit_sim$Z[, "z1"]
sub_sim <- sc_subgroup(fit_sim, list(
  "z1 > 0" = z1_col > 0,
  "z1 < 0" = z1_col < 0
))

## Compare to true subgroup means
true_high <- mean(beta_true[z1 > 0, 1])
true_low  <- mean(beta_true[z1 < 0, 1])

data.frame(
  group     = c("z1 > 0", "z1 < 0"),
  true_b1   = round(c(true_high, true_low), 3),
  est_b1    = round(c(
    sub_sim[["z1 > 0"]]$estimate$theta[1],
    sub_sim[["z1 < 0"]]$estimate$theta[1]
  ), 3),
  row.names = NULL
)
```

## Visualization

### AMCE coefficient plot

```{r}
#| label: sim-amce
#| fig-height: 4
plot_amce(fit_sim)
```

### Fraction favor / oppose

```{r}
#| label: sim-fraction-bar
#| fig-height: 4
plot_fraction(fit_sim)
```

### Heterogeneity

The bar chart should show $x_1$ and $x_2$ in red (significant
heterogeneity) and $x_3$ in gray (homogeneous) --- matching our DGP
design.

```{r}
#| label: sim-het-bar
#| fig-height: 4
plot_hetero(fit_sim)
```

### Subgroup comparison

```{r}
#| label: sim-subgroup-plot
#| fig-height: 4
z1_col <- fit_sim$Z[, "z1"]
plot_subgroup(
  fit_sim,
  subgroup = list("z1 > 0" = z1_col > 0, "z1 < 0" = z1_col < 0),
  title = "Subgroup AMCE: high-z1 vs low-z1 respondents"
)
```

### Beta ridgelines

```{r}
#| label: sim-ridgelines
#| fig-height: 4
plot(fit_sim, "beta_ridgelines")
```

### Loss trace

```{r}
#| label: sim-loss
plot(fit_sim, "loss_trace")
```

## Summary

Every check aligns with the ground truth:

- **Population-average** $\hat\theta$ recovers the sign and
  approximate magnitude of the true $E[\beta]$.
- **Individual-level** $\hat\beta(\mathbf Z_i)$ correlates positively
  with the true per-respondent preference vector.
- **Heterogeneity test** correctly flags $x_1$ and $x_2$
  (Z-dependent) as heterogeneous and $x_3$ (constant) as homogeneous.
- **Subgroup analysis** recovers the known direction: high-$z_1$
  respondents have higher $\beta_1$ than low-$z_1$ respondents.
- **Visualizations** produce the expected patterns: ridgelines for
  $x_1$ and $x_2$ are wide (heterogeneity), $x_3$ is narrow
  (homogeneity).

With these checks in hand, we proceed to real-data applications.
