# Sanity Check {#sec-sanity}

```{r}
#| label: setup-sim
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(42)
torch::torch_manual_seed(42)
library(sconjoint)
library(ggplot2)
```

Before applying the structural estimator to real data, this chapter
runs the full workflow on a synthetic conjoint experiment with
**known ground truth**. Because we control the data-generating process,
every quantity the package produces can be checked against its
analytically correct value.

## Data-generating process

The package ships `simdata`, a synthetic conjoint with $M = 1{,}000$
respondents, $T = 6$ tasks per respondent, and two profiles per task
(12,000 rows). Three binary attributes $x_1, x_2, x_3$ are drawn
independently as Bernoulli(0.5). Two continuous moderators
$z_1, z_2 \sim N(0,1)$ determine each respondent's preference vector:

$$
\beta_1(\mathbf Z_i) = 0.5 + 0.3\,z_{1i}, \quad
\beta_2(\mathbf Z_i) = -0.8 + 0.2\,z_{2i}, \quad
\beta_3(\mathbf Z_i) = 0.3.
$$

Key features:

- $\beta_1$ and $\beta_2$ are **heterogeneous** (vary with $Z$);
  $\beta_3$ is **homogeneous** (constant across respondents).
- $\beta_1 > 0$ for most respondents; $\beta_2 < 0$ for most.
- True population averages: $E[\beta_1] = 0.5$, $E[\beta_2] = -0.8$,
  $E[\beta_3] = 0.3$.

```{r}
#| label: sim-load
data(simdata, package = "sconjoint")
dim(simdata)
head(simdata, 6)

## True beta matrix stored as attribute
beta_true <- attr(simdata, "beta_true")
dim(beta_true)
```

::: {.callout-note}
### Why does `scfit()` require at least one Z column?

The DML estimator is defined in terms of a nuisance function
$\hat\beta(\mathbf Z)$ that maps respondent-level covariates into
preference vectors. Without any moderators the mapping is trivial and
the DML correction degenerates to a plain plug-in estimator. Users who
truly want a homogeneous model should fit a standard
`glm(choice ~ deltaX, family = binomial)` instead.
:::

## Fitting

With 1,000 respondents and 6 tasks each, the estimator has enough signal
to recover both population-level and individual-level parameters
cleanly.

```{r}
#| label: sim-fit
fit_sim <- scfit(
  choice ~ x1 + x2 + x3 | z1 + z2,
  data       = simdata,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 5L,
  n_epochs   = 200L,
  seed       = 42
)
summary(fit_sim)
```

::: {.callout-tip}
### How do I know my `sc_fit` is well-trained?

Check two things: (1) `summary(fit)` reports the DML/iid SE ratio ---
values close to one mean the clustering correction is small and the
DNN's per-respondent $\hat\beta$ has little additional
within-respondent information beyond the population mean.
(2) `plot(fit, "loss_trace")` shows per-fold training loss curves ---
they should visually level off before the last epoch. If they are
still descending, increase `n_epochs`.
:::

::: {.callout-tip}
### Performance tips

- Set `device = "cuda"` on a GPU machine; bit-exact determinism is
  lost but wall-clock time drops substantially.
- Use `parallel = TRUE` with `n_cores = parallelly::availableCores()`
  on CPU for large datasets --- the package guarantees bit-exact
  identity between parallel and sequential runs.
- Prefer `hidden = "auto"` as a starting point and tune only when
  clearly needed.
:::

## Population-average recovery

The DML point estimates $\hat\theta$ should approximate the true
population means.

```{r}
#| label: sim-theta
theta_hat <- coef(fit_sim)
theta_true <- c(x1 = 0.5, x2 = -0.8, x3 = 0.3)
data.frame(
  attribute  = names(theta_hat),
  true       = theta_true,
  estimated  = round(unname(theta_hat), 3),
  row.names  = NULL
)
```

All three signs match and the magnitudes are close. The small
discrepancy is expected: the DNN does not impose the linear-in-$Z$
structure of the DGP, so it needs to learn it from data.

## Individual-level recovery

The per-respondent $\hat\beta(\mathbf Z_i)$ matrix is the structural
payoff --- it gives a complete preference vector for each respondent.

```{r}
#| label: sim-beta-recovery
beta_hat_all <- predict(fit_sim)

## De-duplicate to respondent level
first_row <- !duplicated(fit_sim$respondent_id)
beta_hat <- beta_hat_all[first_row, ]

## Correlation between true and estimated beta
data.frame(
  attribute   = c("x1", "x2", "x3"),
  correlation = round(c(
    cor(beta_true[, 1], beta_hat[, 1]),
    cor(beta_true[, 2], beta_hat[, 2]),
    cor(beta_true[, 3], beta_hat[, 3])
  ), 3),
  row.names = NULL
)
```

High correlations for $x_1$ and $x_2$ confirm that the DNN recovers
the individual-level mapping $\beta(Z)$. The correlation for $x_3$ is
low because $\beta_3$ is constant across respondents --- there is no
heterogeneity to recover, so the correlation reflects sampling noise
rather than a failure of the estimator.

```{r}
#| label: sim-scatter
#| fig-height: 5
df_compare <- data.frame(
  true      = beta_true[, 1],
  estimated = beta_hat[, 1]
)

ggplot(df_compare, aes(x = true, y = estimated)) +
  geom_point(alpha = 0.3, color = "#2166AC", size = 1.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "#E41A1C", linewidth = 0.8) +
  labs(x = expression("True " * beta[1](Z)),
       y = expression("Estimated " * hat(beta)[1](Z)),
       title = "Individual-level recovery: attribute x1") +
  theme_minimal(base_size = 12)
```

The point cloud tracks the 45-degree line, confirming that the
structural estimator recovers both the **direction** and **magnitude**
of individual-level preference heterogeneity.

## Direction and intensity

The direction--intensity decomposition should show: $x_1$ has strong
positive direction (most $\beta_1 > 0$), $x_2$ has strong negative
direction (most $\beta_2 < 0$), and $x_3$ has perfect positive
direction ($\beta_3 = 0.3$ always).

```{r}
#| label: sim-di
sc_direction_intensity(fit_sim)
```

This matches the DGP: $x_1$ and $x_3$ both have positive mean
direction, while $x_2$ has negative mean direction. The intensity
column captures the average magnitude $\overline{|\beta_k|}$.

## Fraction preferring

We can verify against closed-form DGP values. For $x_1$:
$\Pr(\beta_1 > 0) = \Pr(z_1 > -5/3) \approx 0.95$. For $x_2$:
$\Pr(\beta_2 > 0) = \Pr(z_2 > 4) \approx 0$. For $x_3$: $\beta_3 =
0.3 > 0$ always, so the fraction is 1.

```{r}
#| label: sim-fraction
frac_sim <- sc_fraction_preferring(fit_sim)

true_frac <- c(
  x1 = mean(beta_true[, 1] > 0),
  x2 = mean(beta_true[, 2] > 0),
  x3 = mean(beta_true[, 3] > 0)
)

data.frame(
  attribute     = names(true_frac),
  true_fraction = true_frac,
  estimated     = round(frac_sim$estimate$frac_positive, 3),
  row.names     = NULL
)
```

The estimated fractions closely match the DGP truth. Nearly all
respondents favor $x_1$ and $x_3$; nearly all oppose $x_2$.

## Heterogeneity test

This is the most important validation. The test should flag $x_1$ and
$x_2$ as **significant** (their $\beta$ varies with $Z$) and $x_3$ as
**non-significant** (constant across respondents). Getting this
distinction right confirms that the test correctly separates structural
heterogeneity from sampling noise.

```{r}
#| label: sim-het
het_sim <- sc_heterogeneity_test(fit_sim, adjust = "bh")
het_sim$estimate[, c("dummy_name", "var_beta", "p_adjusted", "sig")]
```

The variance of $\hat\beta$ for $x_1$ and $x_2$ is large and highly
significant, while $x_3$ has near-zero variance and a non-significant
p-value --- exactly as the DGP predicts.

## Subgroup analysis

By construction, respondents with $z_1 > 0$ have $\beta_1 > 0.5$
while those with $z_1 < 0$ have $\beta_1 < 0.5$. The subgroup analysis
should recover this split without re-estimating the model.

```{r}
#| label: sim-subgroup
z1_col <- fit_sim$Z[, "z1"]
sub_sim <- sc_subgroup(fit_sim, list(
  "z1 > 0" = z1_col > 0,
  "z1 < 0" = z1_col < 0
))

## z1 at respondent level (de-duplicated, same order as beta_true)
z1_resp <- fit_sim$Z[first_row, "z1"]
true_high <- mean(beta_true[z1_resp > 0, 1])
true_low  <- mean(beta_true[z1_resp < 0, 1])

data.frame(
  group     = c("z1 > 0", "z1 < 0"),
  true_b1   = round(c(true_high, true_low), 3),
  est_b1    = round(c(
    sub_sim[["z1 > 0"]]$estimate$theta[1],
    sub_sim[["z1 < 0"]]$estimate$theta[1]
  ), 3),
  row.names = NULL
)
```

The estimated subgroup AMCEs track the true subgroup means, confirming
that the structural model recovers group-level differences from the
same fit.

## Visualization

### AMCE coefficient plot

```{r}
#| label: sim-amce
#| fig-height: 4
plot_amce(fit_sim)
```

The coefficient plot shows the expected pattern: $x_1$ and $x_3$ are
positive, $x_2$ is negative, with confidence intervals comfortably
excluding zero.

### Fraction favor / oppose

```{r}
#| label: sim-fraction-bar
#| fig-height: 4
plot_fraction(fit_sim)
```

The diverging bar chart makes the asymmetry vivid: nearly all
respondents favor $x_1$ and $x_3$, while nearly all oppose $x_2$.

### Heterogeneity

```{r}
#| label: sim-het-bar
#| fig-height: 4
plot_hetero(fit_sim)
```

The gradient shading reflects strength of evidence. $x_1$ and $x_2$
are dark blue with diamond markers (significant heterogeneity);
$x_3$ is pale (no heterogeneity). This is the cleanest visual
confirmation that the test works.

### Subgroup comparison

```{r}
#| label: sim-subgroup-plot
#| fig-height: 4
z1_col <- fit_sim$Z[, "z1"]
plot_subgroup(
  fit_sim,
  subgroup = list("z1 > 0" = z1_col > 0, "z1 < 0" = z1_col < 0),
  title = "Subgroup AMCE: high-z1 vs low-z1 respondents"
)
```

The high-$z_1$ group shows a larger $\hat\theta_{x_1}$ than the
low-$z_1$ group, exactly as the DGP prescribes ($\beta_1 = 0.5 +
0.3 z_1$). The other attributes show little or no subgroup difference,
which is correct: $\beta_2$ depends on $z_2$ (not $z_1$) and $\beta_3$
is constant.

### Beta ridgelines

```{r}
#| label: sim-ridgelines
#| fig-height: 4
plot(fit_sim, "beta_ridgelines")
```

The ridgeline for $x_1$ is wide and centered above zero (heterogeneous,
positive); $x_2$ is wide and centered below zero (heterogeneous,
negative); $x_3$ is narrow and centered at 0.3 (homogeneous). This
matches the DGP perfectly.

### Loss trace

```{r}
#| label: sim-loss
plot(fit_sim, "loss_trace")
```

All folds converge to a similar loss level, confirming that 200 epochs
is sufficient for this dataset.

## Summary

Every check aligns with the ground truth:

- **Population-average** $\hat\theta$ recovers signs and magnitudes.
- **Individual-level** $\hat\beta(\mathbf Z_i)$ correlates strongly
  with the true preference vector for heterogeneous attributes.
- **Heterogeneity test** correctly flags $x_1$, $x_2$ (Z-dependent)
  as heterogeneous and $x_3$ (constant) as homogeneous.
- **Subgroup analysis** recovers the known $z_1$-driven split in
  $\beta_1$.
- **Visualizations** produce the expected patterns across all six
  plot types.

With these checks in hand, we proceed to real-data applications.
