# Sanity Check {#sec-sanity}

```{r}
#| label: setup-sim
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(42)
torch::torch_manual_seed(42)
library(sconjoint)
library(ggplot2)
```

Before applying the structural estimator to real data, this chapter
runs the full workflow on a synthetic conjoint experiment with
**known ground truth**. Because we control the data-generating process,
every quantity the package produces can be checked against its
analytically correct value.

## Data-generating process

The package ships `simdata`, a synthetic conjoint with $M = 1{,}000$
respondents, $T = 6$ tasks per respondent, and two profiles per task
(12,000 rows). Three binary attributes $x_1, x_2, x_3$ are drawn
independently as Bernoulli(0.5). Two continuous moderators
$z_1, z_2 \sim N(0,1)$ determine each respondent's preference vector:

$$
\beta_1(\mathbf Z_i) = 0.5 + 0.3\,z_{1i}, \quad
\beta_2(\mathbf Z_i) = -0.8 + 0.2\,z_{2i}, \quad
\beta_3(\mathbf Z_i) = 0.3.
$$

Key features:

- $\beta_1$ and $\beta_2$ are **heterogeneous** (vary with $Z$);
  $\beta_3$ is **homogeneous** (constant across respondents).
- $\beta_1 > 0$ for most respondents; $\beta_2 < 0$ for most.
- True population averages: $E[\beta_1] = 0.5$, $E[\beta_2] = -0.8$,
  $E[\beta_3] = 0.3$.

```{r}
#| label: sim-load
data(simdata, package = "sconjoint")
dim(simdata)
head(simdata, 6)

## True beta matrix stored as attribute
beta_true <- attr(simdata, "beta_true")
dim(beta_true)
```

::: {.callout-note}
### Why does `scfit()` require at least one Z column?

The DML estimator is defined in terms of a nuisance function
$\hat\beta(\mathbf Z)$ that maps respondent-level covariates into
preference vectors. Without any moderators the mapping is trivial and
the DML correction degenerates to a plain plug-in estimator. Users who
truly want a homogeneous model should fit a standard
`glm(choice ~ deltaX, family = binomial)` instead.
:::

## Fitting

With 1,000 respondents and 6 tasks each, the estimator has enough signal
to recover both population-level and individual-level parameters
cleanly.

```{r}
#| label: sim-fit
fit_sim <- scfit(
  choice ~ x1 + x2 + x3 | z1 + z2,
  data       = simdata,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 5L,
  n_epochs   = 200L,
  seed       = 42
)
summary(fit_sim)
```

::: {.callout-tip}
### How do I know my `sc_fit` is well-trained?

Check two things: (1) `summary(fit)` reports the DML/iid SE ratio ---
values close to one mean the clustering correction is small and the
DNN's per-respondent $\hat\beta$ has little additional
within-respondent information beyond the population mean.
(2) `plot(fit, "loss_trace")` shows per-fold training loss curves ---
they should visually level off before the last epoch. If they are
still descending, increase `n_epochs`.
:::

::: {.callout-tip}
### Performance tips

- Set `device = "cuda"` on a GPU machine; bit-exact determinism is
  lost but wall-clock time drops substantially.
- Use `parallel = TRUE` with `n_cores = parallelly::availableCores()`
  on CPU for large datasets --- the package guarantees bit-exact
  identity between parallel and sequential runs.
- Prefer `hidden = "auto"` as a starting point and tune only when
  clearly needed.
:::

```{r}
#| label: sim-loss
plot(fit_sim, "loss_trace")
```

All folds converge to a similar loss level, confirming that 200 epochs
is sufficient for this dataset.

## Population-average recovery

The DML point estimates $\hat\theta$ should approximate the true
population means.

```{r}
#| label: sim-theta
theta_hat <- coef(fit_sim)
theta_true <- c(x1 = 0.5, x2 = -0.8, x3 = 0.3)
data.frame(
  attribute  = names(theta_hat),
  true       = theta_true,
  estimated  = round(unname(theta_hat), 3),
  row.names  = NULL
)
```

All three signs match and the magnitudes are close. The AMCE coefficient
plot makes the pattern visual:

```{r}
#| label: sim-amce
#| fig-height: 4
plot_amce(fit_sim)
```

## Individual-level recovery

The per-respondent $\hat\beta(\mathbf Z_i)$ matrix is the structural
payoff --- it gives a complete preference vector for each respondent.

```{r}
#| label: sim-beta-recovery
beta_hat_all <- predict(fit_sim)

## De-duplicate to respondent level
first_row <- !duplicated(fit_sim$respondent_id)
beta_hat <- beta_hat_all[first_row, ]

## Correlation between true and estimated beta
data.frame(
  attribute   = c("x1", "x2", "x3"),
  correlation = round(c(
    cor(beta_true[, 1], beta_hat[, 1]),
    cor(beta_true[, 2], beta_hat[, 2]),
    cor(beta_true[, 3], beta_hat[, 3])
  ), 3),
  row.names = NULL
)
```

High correlations for $x_1$ and $x_2$ confirm that the DNN recovers
the individual-level mapping $\beta(Z)$. The correlation for $x_3$ is
low because $\beta_3$ is constant --- there is no heterogeneity to
recover.

```{r}
#| label: sim-scatter
#| fig-height: 5
df_compare <- data.frame(
  true      = beta_true[, 1],
  estimated = beta_hat[, 1]
)

ggplot(df_compare, aes(x = true, y = estimated)) +
  geom_point(alpha = 0.3, color = "#2166AC", size = 1.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "#E41A1C", linewidth = 0.8) +
  labs(x = expression("True " * beta[1](Z)),
       y = expression("Estimated " * hat(beta)[1](Z)),
       title = "Individual-level recovery: attribute x1") +
  theme_minimal(base_size = 12)
```

The point cloud tracks the 45-degree line, confirming recovery of both
**direction** and **magnitude**. The beta ridgelines below show the
full distribution:

```{r}
#| label: sim-ridgelines
#| fig-height: 4
plot(fit_sim, "beta_ridgelines")
```

The ridgeline for $x_1$ is wide and centered above zero (heterogeneous,
positive); $x_2$ is wide and centered below zero (heterogeneous,
negative); $x_3$ is narrow and centered at 0.3 (homogeneous).

## Direction and intensity

The decomposition should show: $x_1$ has strong positive direction,
$x_2$ has strong negative direction, $x_3$ has moderate positive
direction.

```{r}
#| label: sim-di
sc_direction_intensity(fit_sim)
```

This matches the DGP: $x_1$ and $x_3$ both have positive mean
direction, while $x_2$ has negative mean direction.

## Fraction preferring

We can verify against closed-form DGP values. For $x_1$:
$\Pr(\beta_1 > 0) = \Pr(z_1 > -5/3) \approx 0.95$. For $x_2$:
$\Pr(\beta_2 > 0) = \Pr(z_2 > 4) \approx 0$. For $x_3$: always 1.

```{r}
#| label: sim-fraction
frac_sim <- sc_fraction_preferring(fit_sim)

true_frac <- c(
  x1 = mean(beta_true[, 1] > 0),
  x2 = mean(beta_true[, 2] > 0),
  x3 = mean(beta_true[, 3] > 0)
)

data.frame(
  attribute     = names(true_frac),
  true_fraction = true_frac,
  estimated     = round(frac_sim$estimate$frac_positive, 3),
  row.names     = NULL
)
```

The diverging bar chart makes the asymmetry vivid:

```{r}
#| label: sim-fraction-bar
#| fig-height: 4
plot_fraction(fit_sim)
```

## Heterogeneity test

This is the most important validation. The test should flag $x_1$ and
$x_2$ as **significant** (their $\beta$ varies with $Z$) and $x_3$ as
**non-significant** (constant across respondents).

```{r}
#| label: sim-het
het_sim <- sc_heterogeneity_test(fit_sim, adjust = "bh")
het_sim$estimate[, c("dummy_name", "var_beta", "p_adjusted", "sig")]
```

The heterogeneity bar chart confirms this visually --- $x_1$ and $x_2$
are dark blue with diamond markers (significant), $x_3$ is pale
(homogeneous):

```{r}
#| label: sim-het-bar
#| fig-height: 4
plot_hetero(fit_sim)
```

## Subgroup analysis

By construction, respondents with $z_1 > 0$ have $\beta_1 > 0.5$
while those with $z_1 < 0$ have $\beta_1 < 0.5$.

```{r}
#| label: sim-subgroup
z1_col <- fit_sim$Z[, "z1"]
sub_sim <- sc_subgroup(fit_sim, list(
  "z1 > 0" = z1_col > 0,
  "z1 < 0" = z1_col < 0
))

z1_resp <- fit_sim$Z[first_row, "z1"]
true_high <- mean(beta_true[z1_resp > 0, 1])
true_low  <- mean(beta_true[z1_resp < 0, 1])

data.frame(
  group     = c("z1 > 0", "z1 < 0"),
  true_b1   = round(c(true_high, true_low), 3),
  est_b1    = round(c(
    sub_sim[["z1 > 0"]]$estimate$theta[1],
    sub_sim[["z1 < 0"]]$estimate$theta[1]
  ), 3),
  row.names = NULL
)
```

The subgroup comparison plot shows the split clearly --- the high-$z_1$
group has a larger $\hat\theta_{x_1}$, exactly as the DGP prescribes:

```{r}
#| label: sim-subgroup-plot
#| fig-height: 4
plot_subgroup(
  fit_sim,
  subgroup = list("z1 > 0" = z1_col > 0, "z1 < 0" = z1_col < 0),
  title = "Subgroup AMCE: high-z1 vs low-z1 respondents"
)
```

## Advanced quantities

### Baseline comparison

The homogeneous logit baseline should agree with the structural model
on population averages while lacking per-respondent heterogeneity.

```{r}
#| label: sim-baseline
bl <- sc_baseline_logit(fit_sim)
summary(bl)
```

```{r}
#| label: sim-baseline-compare
#| fig-height: 4
theta_struct <- coef(fit_sim)
theta_logit  <- coef(bl)
df_comp <- data.frame(
  dummy = rep(names(theta_struct), 2),
  estimate = c(theta_struct, theta_logit),
  model = rep(c("Structural (DML)", "Logit"), each = length(theta_struct))
)
ggplot(df_comp, aes(x = estimate, y = dummy, color = model)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(position = position_dodge(0.4), size = 3) +
  scale_color_manual(values = c("Structural (DML)" = "#E41A1C",
                                "Logit" = "#2166AC")) +
  labs(x = expression(hat(theta)), y = NULL, color = NULL,
       title = "Structural vs baseline logit") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

Both estimators agree on population averages --- the structural model's
advantage is the per-respondent $\hat\beta(\mathbf Z_i)$ that enables
all the heterogeneity analyses above.

### Average marginal effects (probability scale)

```{r}
#| label: sim-ame-prob
sc_average(fit_sim, scale = "probability")
```

Probability-scale AMEs are attenuated relative to logit-scale
$\hat\theta$ because $G'(x) = G(x)(1 - G(x)) \leq 0.25$.

### Consumer surplus and welfare change

```{r}
#| label: sim-surplus
sc_surplus(fit_sim, profiles = list(
  list(x1 = 1, x2 = 0, x3 = 1),
  list(x1 = 0, x2 = 1, x3 = 0)
))
```

Upgrading $x_1$ from 0 to 1 should increase welfare ($\beta_1 > 0$):

```{r}
#| label: sim-welfare
sc_welfare_change(fit_sim,
  old_set = list(list(x1 = 0)),
  new_set = list(list(x1 = 1))
)
```

The positive welfare change confirms that $x_1 = 1$ is valued, with
magnitude approximately $E[\beta_1] = 0.5$.

### Decisiveness

```{r}
#| label: sim-decisiveness
sc_decisiveness(fit_sim,
  A = list(x1 = 1, x3 = 1),
  B = list(x2 = 1)
)
```

Profile A ($\beta_1 + \beta_3 \approx 0.8$) vs profile B
($\beta_2 \approx -0.8$) gives $\Delta V \approx 1.6$, so most
respondents are decisive.

### Preference inequality

```{r}
#| label: sim-inequality
sc_inequality(fit_sim, measure = "variance")
```

$x_1$ and $x_2$ show higher variance ($\beta$ depends on $Z$);
$x_3$ has near-zero variance (constant $\beta_3 = 0.3$). This agrees
with the heterogeneity test above.

## Summary

Every check aligns with the ground truth:

- **Population-average** $\hat\theta$ recovers signs and magnitudes.
- **Individual-level** $\hat\beta(\mathbf Z_i)$ correlates strongly
  with the true preference vector for heterogeneous attributes.
- **Heterogeneity test** correctly flags $x_1$, $x_2$ (Z-dependent)
  as heterogeneous and $x_3$ (constant) as homogeneous.
- **Subgroup analysis** recovers the known $z_1$-driven split.
- **Baseline logit** agrees on population averages while lacking
  heterogeneity decomposition.
- **Advanced quantities** (AME, surplus, welfare, decisiveness,
  inequality) all produce results consistent with the DGP.

With these checks in hand, we proceed to real-data applications.
