# Example: Candidate Choice {#sec-sw}

```{r}
#| label: setup-sw
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Groups and labels for plot annotations --------------------------------
## Keyed by the attribute-dummy column names in the bundled data.
sw_groups <- c(
  cand_agendaModerate.Changes      = "Agenda",
  cand_agendaComplete.Overhaul     = "Agenda",
  cand_talentCollaborative         = "Talent",
  cand_talentDetermined.to.Succeed = "Talent",
  cand_talentEmpathetic            = "Talent",
  cand_talentGood.Communicator     = "Talent",
  cand_talentHard.Working          = "Talent",
  cand_talentTough.Negotiator      = "Talent",
  cand_child1.child                = "Children",
  cand_child2.children             = "Children",
  cand_child3.children             = "Children",
  cand_genderMale                  = "Gender",
  cand_runYes                      = "Prior run"
)

sw_labels <- c(
  cand_agendaModerate.Changes      = "Moderate changes",
  cand_agendaComplete.Overhaul     = "Complete overhaul",
  cand_talentCollaborative         = "Collaborative",
  cand_talentDetermined.to.Succeed = "Determined to succeed",
  cand_talentEmpathetic            = "Empathetic",
  cand_talentGood.Communicator     = "Good communicator",
  cand_talentHard.Working          = "Hard-working",
  cand_talentTough.Negotiator      = "Tough negotiator",
  cand_child1.child                = "1 child",
  cand_child2.children             = "2 children",
  cand_child3.children             = "3 children",
  cand_genderMale                  = "Male",
  cand_runYes                      = "Ran before"
)

## Attribute grouping for the variance-decomposition importance ----------
## The bundled formula enters each level as its own dummy, so we tell
## sc_importance() which dummies belong to the same attribute before
## computing the (paper-convention) share. sw_attr_map below is reused in
## the "Attribute importance" section.
sw_attr_map <- function(fit) {
  an  <- fit$attr_names
  map <- list(
    cand_gender = which(an == "cand_genderMale"),
    cand_run    = which(an == "cand_runYes"),
    cand_talent = which(grepl("^cand_talent",  an)),
    cand_agenda = which(grepl("^cand_agenda",  an)),
    cand_child  = which(grepl("^cand_child",   an))
  )
  fit$attr_map  <- map
  fit$attr_vars <- names(map)
  fit
}
```

This chapter walks through a complete `sconjoint` analysis of the
candidate-choice conjoint experiment of @saha2022ambitious. Respondents
evaluate pairs of hypothetical political candidates and select one. The
structural decomposition into direction and intensity lets us ask
whether a near-zero average effect for candidate gender masks genuine
heterogeneity --- opposing camps of voters who prefer women and men.

The bundled `sw2022` is the paper's full analysis frame, so the
quantities below reproduce the published estimates. The fit uses the
paper's production configuration (`K = 10` cross-fitting folds and the
default 1,000 Adam epochs); it is heavier than a quick demo and takes a
couple of minutes to run.

## Data preparation

Load the data and inspect its structure.

```{r}
#| label: sw-load
data(sw2022, package = "sconjoint")
dim(sw2022)
head(sw2022, 4)
```

The data are in long format: one row per respondent--task--profile. Each
respondent contributes three forced-choice tasks; within a task two
candidate profiles are presented and exactly one is chosen
(`choice == 1`).

```{r}
#| label: sw-structure
str(sw2022, list.len = 12)
table(table(sw2022$respondent))
```

The thirteen candidate attributes are already coded as 0/1 dummies, with
the omitted level of each attribute serving as its reference (Female for
gender, no prior run for ambition, and so on). The
respondent-level moderators in $\mathbf Z$ are the full set of nineteen
covariates from the paper --- gender, age, income, education, party,
region, employment status, ideology, vote choice, and a gender-attitudes
scale. The convenience factor `resp_party` (Democrat / Independent /
Republican) is carried alongside for subgroup labelling but is not part
of $\mathbf Z$.

The formula interface puts conjoint attributes to the left of `|` and
moderators to the right. Column names containing dots are backquoted:

```{r}
#| label: sw-formula
attrs <- c("cand_genderMale", "cand_runYes",
           "cand_talentCollaborative", "cand_talentDetermined.to.Succeed",
           "cand_talentEmpathetic", "cand_talentGood.Communicator",
           "cand_talentHard.Working", "cand_talentTough.Negotiator",
           "cand_agendaModerate.Changes", "cand_agendaComplete.Overhaul",
           "cand_child1.child", "cand_child2.children", "cand_child3.children")
Zvars <- c("gender_num", "age", "income", "educ_Middle", "educ_High",
           "party_Republican", "party_Independent", "region_NORTHEAST",
           "region_SOUTH", "region_WEST", "employ_parttime",
           "employ_homemaker", "employ_not_working", "employ_retired",
           "employ_student", "ideo_conservative", "vote_trump",
           "vote_clinton", "gender_att")

bt  <- function(x) paste0("`", x, "`")
sw_formula <- stats::as.formula(
  paste("choice ~", paste(bt(attrs), collapse = " + "),
        "|",       paste(bt(Zvars), collapse = " + "))
)
```

## Fitting

We fit the production configuration: `K = 10` cross-fitting folds and the
default 1,000 Adam epochs. This is the recipe that reproduces the paper,
so the chunk is cached --- rendering reuses the saved fit rather than
retraining each time. The loss trace later in the chapter shows the
per-fold training curves.

```{r}
#| label: sw-fit
#| cache: true
fit_sw <- scfit(
  sw_formula,
  data       = sw2022,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 10L,
  seed       = 42L,
  stage2     = "map_c5",
  parallel   = TRUE,
  n_cores    = 10L
)
summary(fit_sw)
```

::: {.callout-note}
### The Stage-2 default

`scfit()` runs an empirical-Bayes MAP refinement (paper `EnsC5`) on top
of the Stage-1 DNN by default (`stage2 = "map_c5"`). The summary above
reports this as `Stage 2: map_c5`. DML point estimates and clustered
standard errors continue to use the Stage-1 single-DNN prediction
(stored on `fit_sw$beta_hat_dnn`). The refined betas
(`fit_sw$beta_hat`) feed every distributional and individual-level
quantity below. Set `stage2 = "none"` to recover the unrefined
behaviour, or `stage2 = "mixed_logit"` for the BLUP alternative from the
paper's Supplementary Materials.
:::

## Population-average estimates

The DML point estimates $\hat\theta_k$ on the logit scale. Inspect
numerically, then visualize as a coefficient plot grouped by attribute.

```{r}
#| label: sw-coef
coef(fit_sw)
```

Agenda dominates: relative to keeping the status quo, both *moderate
changes* and a *complete overhaul* raise candidate utility by about 0.95
on the logit scale. Among the talent cues, *hard-working* is the
strongest (about +0.55). The average effect of a *male* candidate is
small and negative (about $-0.10$, confidence interval covering zero) ---
the near-zero average we return to below.

```{r}
#| label: sw-amce-plot
#| fig-height: 6
plot_amce(fit_sw, groups = sw_groups, labels = sw_labels)
```

### Structural AME vs reduced-form AMCE

The structural model's average marginal effect (AME) on the
*probability* scale nearly coincides with the standard
linear-probability-model AMCE, confirming that the structural estimator
nests reduced-form AMCE analysis. We compute both and overlay them as
point-ranges on the probability scale.

```{r}
#| label: sw-ame-vs-amce
#| fig-height: 6
## Structural AME on the probability scale (delta-method via plogis')
ame_struct <- sc_average(fit_sw, scale = "probability")
ame_df     <- ame_struct$estimate
ame_df$source <- "Structural AME"

## Linear-probability-model AMCE
lpm        <- sc_baseline_lpm(fit_sw)
amce_df    <- data.frame(
  dummy_name = names(stats::coef(lpm)),
  estimate   = unname(stats::coef(lpm)),
  se         = sqrt(diag(stats::vcov(lpm))),
  source     = "LPM AMCE",
  stringsAsFactors = FALSE
)
amce_df$ci_lo <- amce_df$estimate - 1.96 * amce_df$se
amce_df$ci_hi <- amce_df$estimate + 1.96 * amce_df$se

df_both <- rbind(
  ame_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")],
  amce_df[amce_df$dummy_name %in% names(sw_labels),
          c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")]
)
df_both$label <- sw_labels[df_both$dummy_name]
df_both$label <- factor(df_both$label, levels = rev(unname(sw_labels)))
df_both$group <- sw_groups[df_both$dummy_name]
df_both$group <- factor(df_both$group, levels = unique(sw_groups))

ggplot(df_both, aes(x = estimate, y = label, color = source)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  position = position_dodge(width = 0.5),
                  size = 0.35, fatten = 2) +
  scale_color_manual(values = c("Structural AME" = "#2166AC",
                                "LPM AMCE"        = "#B2182B")) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Effect on Pr(choice)",
       y = NULL, color = NULL,
       title = "Structural AME vs LPM AMCE (probability scale)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        strip.text.y    = element_text(angle = 0, hjust = 0, face = "bold"))
```

The two series overlap nearly point-for-point: the structural model
loses nothing the reduced-form AMCE provides. The
`sc_validate_amce()` check at the end of the chapter quantifies this
agreement.

## Individual-level preferences

Pull the respondent-level $\hat\beta(\mathbf Z_i)$ matrix. With the
default `stage2 = "map_c5"`, this is the MAP-refined hybrid view.

```{r}
#| label: sw-beta
beta <- predict(fit_sw)
dim(beta)
head(beta, 3)
```

Ridgeline densities show the distribution of $\hat\beta_k(\mathbf Z_i)$
per dummy across respondents. The vertical dashed line is at zero; the
solid line in each density is the median.

```{r}
#| label: sw-ridgelines
#| fig-height: 11
plot(fit_sw, "beta_ridgelines", groups = sw_groups, labels = sw_labels)
```

The gender density is centred just left of zero but spreads to both
sides --- the visual signature of two opposing camps rather than a
shared near-indifference, a pattern we make precise next.

## Structural quantities

### Attribute importance

The variance-decomposition share of utility carried by each *attribute*,
averaged across respondents. Importance answers: which attribute moves
the choice the most? Because the bundled formula enters each level as a
separate dummy, we first tell `sc_importance()` which dummies belong to
the same attribute (via the `sw_attr_map()` helper defined in the setup
chunk), then read off the paper's five-attribute decomposition.

```{r}
#| label: sw-importance
fit_grp <- sw_attr_map(fit_sw)
imp <- sc_importance(fit_grp, design = "uniform")
imp$estimate[, c("attribute", "share", "se")]
```

```{r}
#| label: sw-importance-plot
#| fig-height: 6
plot_importance(fit_grp, labels = c(
  cand_agenda = "Agenda", cand_talent = "Talent",
  cand_child = "Children", cand_gender = "Gender", cand_run = "Prior run"))
```

Policy agenda dominates the decision variance, carrying roughly half the
explained variation. Candidate gender accounts for about **16%** of the
variance --- the third-largest attribute, behind agenda and talent and
well ahead of children and prior office. This is the first sign that
gender matters far more than its near-zero average effect suggests: the
variance is large precisely because respondents disagree about its sign.

### Direction and intensity

For each dummy, decompose preference into a *direction* (the sign of
$\hat\beta_{ik}$) and an *intensity* (its magnitude). A near-zero
average effect can hide either consensus (everyone close to zero) or
polarization (large opposing intensities cancelling).

```{r}
#| label: sw-di
sc_direction_intensity(fit_sw)
```

Candidate gender is the paper's clearest example of cancellation: its
direction summary sits near zero (a slim majority lean toward female
candidates), but its intensity is substantial --- the opposing camps
hold their preferences with real strength, and they cancel in the
average.

### Fraction preferring

What share of respondents have a positive coefficient on each dummy?
This is the direction-only summary --- it ignores intensity.

```{r}
#| label: sw-fraction
frac <- sc_fraction_preferring(fit_sw, threshold = 0)
frac$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

```{r}
#| label: sw-fraction-bar
#| fig-height: 6
plot_fraction(fit_sw, groups = sw_groups, labels = sw_labels)
```

The directional finding for gender: the population-average effect for
*male* is near zero, yet a clear majority of respondents have
$\hat\beta_{\text{Male}} < 0$ --- about **54%** prefer female candidates
against **46%** who prefer men. The electorate is split close to evenly
on gender, not indifferent to it. The partisan structure behind that
split appears in the subgroup section below.

### Marginal rate of substitution

The respondent-averaged trimmed ratio $\hat\beta_j / \hat\beta_k$
between two dummies, with respondent-clustered SE. Trimming at
$\{0.01, 0.99\}$ guards the unbounded ratio when the denominator is
near zero.

```{r}
#| label: sw-mrs
sc_mrs(fit_sw, numerator = 1L, denominator = 2L)
```

::: {.callout-tip}
### Large trimmed mean?

The ratio $\hat\beta_j / \hat\beta_k$ is unbounded when the
denominator is near zero. Here the numerator is the *male* dummy, whose
recovered effect is near zero for many respondents, so the ratio is
noisy; inspect the ridgelines above to confirm the denominator column
is well-separated from zero before reading a MRS literally.
:::

### Counterfactual choice probability

Predicted choice probability for a *specific* head-to-head profile pair,
averaged across respondents. Each profile is a named list of dummy
settings; unmentioned dummies stay at zero (the reference level).

```{r}
#| label: sw-counterfactual
sc_counterfactual(
  fit_sw,
  ## Profile A: a Female, Empathetic candidate proposing a Complete
  ## Overhaul, with two children and a prior run for office. Female is
  ## the reference level, so cand_genderMale stays 0.
  ## Profile B: a Male, Tough-Negotiator candidate with no prior run and
  ## the omitted (status-quo) agenda.
  A = list(cand_talentEmpathetic = 1, cand_agendaComplete.Overhaul = 1,
           cand_child2.children = 1, cand_runYes = 1),
  B = list(cand_genderMale = 1, cand_talentTough.Negotiator = 1)
)
```

### Optimal profile

The single profile that maximizes the population-average utility (greedy
attribute-by-attribute selection over the discrete attribute levels).

```{r}
#| label: sw-optimal
sc_optimal_profile(fit_sw)
```

### Heterogeneity test

For each dummy, the one-sided test that $\mathrm{Var}(\beta_k(Z)) > 0$.
Useful for screening which attributes genuinely vary across respondents.

```{r}
#| label: sw-het
sc_heterogeneity_test(fit_sw, adjust = "bh")
```

```{r}
#| label: sw-het-bar
#| fig-height: 6
plot_hetero(fit_sw, groups = sw_groups, labels = sw_labels)
```

Candidate gender shows highly significant variance in $\beta(Z)$ --- the
formal counterpart to this cancellation: the spread around the
near-zero mean is real, not sampling noise.

### Subgroup analysis

Re-average the respondent-level betas over a subset of respondents, then
compare across subsets. The paper's headline result is partisan: the
null average effect of gender masks Democrats who prefer women and
Republicans who prefer men. We recover it by splitting on `resp_party`.
Subgroup masks are aligned to the rows of `fit_sw$beta_hat`, which `Z`
shares, so we build them from the party indicators in `fit_sw$Z`.

```{r}
#| label: sw-subgroup
Z   <- fit_sw$Z
dem <- Z[, "party_Republican"] == 0 & Z[, "party_Independent"] == 0
ind <- Z[, "party_Independent"] == 1
rep <- Z[, "party_Republican"] == 1

male_row <- function(mask) {
  s <- sc_subgroup(fit_sw, mask)
  s$estimate[s$estimate$dummy_name == "cand_genderMale", c("theta", "se")]
}
rbind(
  Democrat    = male_row(dem),
  Independent = male_row(ind),
  Republican  = male_row(rep)
)
```

The recovered party means show the paper's finding directly: the average
*male* coefficient is clearly **negative for Democrats** (they prefer
women), clearly **positive for Republicans** (they prefer men), and
**near zero for Independents**. Two camps of comparable strength pull in
opposite directions, and the overall average lands near zero by
cancellation, not by indifference.

```{r}
#| label: sw-subgroup-plot
#| fig-height: 6
plot_subgroup(
  fit_sw,
  subgroup = list(Democrat = dem, Independent = ind, Republican = rep),
  dummies  = "cand_genderMale",
  labels   = sw_labels,
  title    = "Recovered male-candidate preference, by party"
)
```

We can run the same split on the respondent's own gender, coded in the
standardized moderator `gender_num` (its two levels label the two
coded groups; the package does not store which sign is female, so we
keep the labels neutral):

```{r}
#| label: sw-subgroup-gender
g_lo <- fit_sw$Z[, "gender_num"] < 0
g_hi <- fit_sw$Z[, "gender_num"] > 0
rbind(
  `gender_num < 0` = male_row(g_lo),
  `gender_num > 0` = male_row(g_hi)
)
```

### Loss trace

```{r}
#| label: sw-loss
plot(fit_sw, "loss_trace")
```

### Design diagnostic (R²_Z + recovery-tier hint)

`sc_design_diagnostic()` estimates per-coefficient $\hat R^2_{Z,k}$ from
the MAP posterior and reports which recovery tiers the design supports
(mean / distributional / individual / ratio) per the paper's
design-guidance heuristics. On the Saha-Weeks design (six tasks per
respondent, the full nineteen-covariate set) the talent attributes are
well pinned by $\mathbf Z$, while gender, prior run, and children rely
most on the within-respondent task variation.

```{r}
#| label: sw-design-diag
sc_design_diagnostic(fit_sw)
```

### Validating against the homogeneous-logit AMCE {#sec-sw-validate}

The paper's Supplementary Materials (Section D) show that the structural
model nests the standard reduced-form AMCE: the DML population-average
parameter $\theta_k = \mathbb{E}[\beta_k(Z)]$ equals the pooled
homogeneous-logit coefficient on attribute $k$ when the logit
specification holds. The `sc_validate_amce()` export turns this into a
one-line sanity check.

```{r}
#| label: sw-validate
v <- sc_validate_amce(fit_sw)
v
```

At the production configuration used here (`K = 10`, 1,000 epochs) the
correlation across attribute levels is very high (close to 1), confirming
that the structural estimator reproduces the reduced-form benchmark. A
material drop would signal either insufficient training or a genuinely
non-logit data-generating process.

### Comparing the Stage-2 view to the raw DNN view

Every `sc_*` quantity accepts a `which_beta` argument with default
`"hybrid"` (Stage-2 refined). Passing `"dnn"` reads the unrefined
Stage-1 DNN mean, which lives in `fit_sw$beta_hat_dnn`.

```{r}
#| label: sw-which-beta
frac_hyb <- sc_fraction_preferring(fit_sw, which_beta = "hybrid")
frac_dnn <- sc_fraction_preferring(fit_sw, which_beta = "dnn")
data.frame(
  attribute   = frac_hyb$estimate$dummy_name,
  hybrid_pref = round(frac_hyb$estimate$frac_positive, 3),
  dnn_pref    = round(frac_dnn$estimate$frac_positive, 3)
)
```

The hybrid and raw-DNN fractions can differ because the Stage-2 MAP
update pulls each respondent's coefficients toward the mean stage. How
much a column's fraction moves --- and in which direction --- depends on
how much it leaned on noisy individual variation; compare the two
columns in the table above.
