# Example: Candidate Choice {#sec-sw}

```{r}
#| label: setup-sw
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Groups and labels for plot annotations --------------------------------
sw_groups <- c(
  "agendaModerate Changes"      = "Agenda",
  "agendaComplete Overhaul"     = "Agenda",
  "talentCollaborative"         = "Talent",
  "talentDetermined to Succeed" = "Talent",
  "talentEmpathetic"            = "Talent",
  "talentGood Communicator"     = "Talent",
  "talentHard-Working"          = "Talent",
  "talentTough Negotiator"      = "Talent",
  "children1 child"             = "Children",
  "children2 children"          = "Children",
  "children3 children"          = "Children",
  "cand_genderFemale"           = "Gender",
  "prior_officeYes"             = "Prior Office"
)

sw_labels <- c(
  "agendaModerate Changes"      = "Moderate Changes",
  "agendaComplete Overhaul"     = "Complete Overhaul",
  "talentCollaborative"         = "Collaborative",
  "talentDetermined to Succeed" = "Determined to Succeed",
  "talentEmpathetic"            = "Empathetic",
  "talentGood Communicator"     = "Good Communicator",
  "talentHard-Working"          = "Hard-Working",
  "talentTough Negotiator"      = "Tough Negotiator",
  "children1 child"             = "1 child",
  "children2 children"          = "2 children",
  "children3 children"          = "3 children",
  "cand_genderFemale"           = "Female",
  "prior_officeYes"             = "Yes"
)
```

This chapter walks through a complete `sconjoint` analysis of a
candidate-choice conjoint experiment whose design follows
@saha2022welfare. Respondents evaluate pairs of hypothetical
political candidates varying on five attributes and select one. The
structural decomposition into direction and intensity lets us ask
whether a near-zero AMCE for candidate gender masks genuine
heterogeneity --- half the sample preferring women, half preferring men.

## Data preparation

```{r}
#| label: sw-load
data(sw2022, package = "sconjoint")
dim(sw2022)
head(sw2022, 4)
```

The data are in long format: one row per respondent--task--profile. Each
respondent contributes three forced-choice tasks; within a task two
candidate profiles are presented and exactly one is chosen
(`choice == 1`).

```{r}
#| label: sw-structure
str(sw2022)
table(table(sw2022$respondent))
```

The respondent-level moderators available in $\mathbf Z$ are
`resp_female` (binary), `age` (standardized), and `pid` (3-level
partisanship index). The formula interface puts conjoint attributes to
the left of `|` and moderators to the right:

```
choice ~ agenda + talent + children + cand_gender + prior_office |
         resp_female + age + pid
```

## Fitting

We fit a small, fast configuration (`K = 3` folds, 60 Adam epochs)
suitable for pedagogy. Production analyses would use more folds and more
epochs.

```{r}
#| label: sw-fit
fit_sw <- scfit(
  choice ~ agenda + talent + children + cand_gender + prior_office |
           resp_female + age + pid,
  data        = sw2022,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_sw)
```

## Population-average estimates

```{r}
#| label: sw-coef
coef(fit_sw)
```

```{r}
#| label: sw-amce-plot
#| fig-height: 6
plot_amce(fit_sw, groups = sw_groups, labels = sw_labels)
```

## Individual-level preferences

```{r}
#| label: sw-beta
beta <- predict(fit_sw)
dim(beta)
head(beta, 3)
```

```{r}
#| label: sw-ridgelines
#| fig-height: 8
plot(fit_sw, "beta_ridgelines", groups = sw_groups, labels = sw_labels)
```

## Structural quantities

### Attribute importance

```{r}
#| label: sw-importance
sc_importance(fit_sw)
```

```{r}
#| label: sw-importance-plot
#| fig-height: 5
plot_importance(fit_sw, labels = c(agenda = "Agenda", talent = "Talent",
  children = "Children", cand_gender = "Gender", prior_office = "Prior Office"))
```

Policy agenda dominates, accounting for approximately two-thirds of the
decision variance for the median voter.

### Direction and intensity

```{r}
#| label: sw-di
sc_direction_intensity(fit_sw)
```

For the Empathetic attribute, the near-zero average masks an almost
perfectly polarized electorate: approximately half favor empathetic
candidates while the other half oppose them.

### Fraction preferring

```{r}
#| label: sw-fraction
frac <- sc_fraction_preferring(fit_sw, threshold = 0)
frac$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

```{r}
#| label: sw-fraction-bar
#| fig-height: 6
plot_fraction(fit_sw, groups = sw_groups, labels = sw_labels)
```

The structural model reveals that 83% of respondents prefer female
candidates despite a near-zero population-average AMCE for gender.

### Marginal rate of substitution

```{r}
#| label: sw-mrs
sc_mrs(fit_sw, numerator = 1L, denominator = 2L)
```

::: {.callout-tip}
### Large trimmed mean?

The ratio $\hat\beta_j / \hat\beta_k$ is unbounded when the
denominator is near zero. Inspect the ridgelines above to confirm the
denominator column is well-separated from zero.
:::

### Counterfactual choice probability

```{r}
#| label: sw-counterfactual
sc_counterfactual(
  fit_sw,
  A = list(agenda = "Complete Overhaul", talent = "Hard-Working",
           children = "2 children", cand_gender = "Female",
           prior_office = "Yes"),
  B = list(agenda = "Very Few Changes", talent = "Assertive",
           children = "No children", cand_gender = "Male",
           prior_office = "No")
)
```

### Optimal profile

```{r}
#| label: sw-optimal
sc_optimal_profile(fit_sw)
```

### Heterogeneity test

```{r}
#| label: sw-het
sc_heterogeneity_test(fit_sw, adjust = "bh")
```

```{r}
#| label: sw-het-bar
#| fig-height: 6
plot_hetero(fit_sw, groups = sw_groups, labels = sw_labels)
```

### Subgroup analysis

```{r}
#| label: sw-subgroup
fem <- fit_sw$Z[, "resp_female"] > 0.5
sub_fem  <- sc_subgroup(fit_sw, fem)
sub_male <- sc_subgroup(fit_sw, !fem)
rbind(
  female = sub_fem$estimate[sub_fem$estimate$dummy_name == "cand_genderFemale",
                            c("theta", "se")],
  male   = sub_male$estimate[sub_male$estimate$dummy_name == "cand_genderFemale",
                            c("theta", "se")]
)
```

```{r}
#| label: sw-subgroup-plot
#| fig-height: 6
plot_subgroup(
  fit_sw,
  subgroup = list(Female = fit_sw$Z[, "resp_female"] > 0.5,
                  Male   = fit_sw$Z[, "resp_female"] <= 0.5),
  groups = sw_groups, labels = sw_labels,
  title = "Subgroup AMCE: female vs male respondents"
)
```

### Loss trace

```{r}
#| label: sw-loss
plot(fit_sw, "loss_trace")
```
