---
title: "Democratic Norms"
---

```{r}
#| label: setup-gs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)
```

This chapter applies the structural estimator to a candidate-choice
conjoint whose design follows @graham2020democracy. Respondents
choose between pairs of hypothetical U.S. candidates; one attribute
captures the candidate's willingness to engage in undemocratic behavior
(ignoring a court ruling, restricting the opposition press, or banning
an opposition rally). The original study asks how much voters penalize
co-partisan candidates for violating democratic norms, and whether
strong partisanship leads to tolerance of such violations.

The structural model adds value here because it decomposes the
population-average penalty for undemocratic behavior into direction and
intensity, and traces the heterogeneity back to respondent-level
partisanship --- without re-estimating on subsets.

::: {.callout-note}
The `gs2020` dataset is a **seed-fixed synthetic fixture** for
illustration. Do not use it to draw substantive conclusions about the
Graham--Svolik population.
:::

## Data preparation

```{r}
#| label: gs-load
data(gs2020, package = "sconjoint")
dim(gs2020)
head(gs2020, 4)
str(gs2020)
```

Each respondent sees eight tasks with two candidate profiles each,
yielding 16 rows per respondent. The respondent-level moderators are
`resp_pid` (-1/0/+1 partisanship), `resp_ideo` (standardized ideology),
and standardized `age`.

## Fitting the structural model

```{r}
#| label: gs-fit
fit_gs <- scfit(
  choice ~ federal + immigration + tax + abortion + undem + cand_party |
           resp_pid + resp_ideo + age,
  data        = gs2020,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_gs)
```

## Heterogeneity test

Does the per-attribute dispersion of $\hat\beta_k(\mathbf Z_i)$ exceed
what sampling noise alone would explain? A significant result for
`undem` levels would confirm that the penalty for undemocratic behavior
varies meaningfully across respondents.

```{r}
#| label: gs-het
het_gs <- sc_heterogeneity_test(fit_gs, adjust = "bh")
het_gs$estimate
```

## Direction and intensity

The undemocratic levels should show a large negative average direction
(voters penalize them) coupled with high intensity. The structural
question is whether the *dispersion* of those coefficients concentrates
on co-partisans, as Graham and Svolik argue.

```{r}
#| label: gs-di
sc_direction_intensity(fit_gs)
```

## Subgroup analysis by partisanship

We re-average the held-out $\hat\beta(\mathbf Z_i)$ separately for
Democrats, Independents, and Republicans using `sc_subgroup()`. The
`resp_pid` column on the standardized $\mathbf Z$ matrix preserves the
sign of the underlying -1/0/+1 encoding.

```{r}
#| label: gs-subgroup
pid_col <- fit_gs$Z[, "resp_pid"]
sub <- sc_subgroup(fit_gs, list(
  dem = pid_col < -0.5,
  ind = abs(pid_col) <= 0.5,
  rep = pid_col >  0.5
))
do.call(rbind, lapply(names(sub), function(g) {
  e <- sub[[g]]$estimate
  rows <- grepl("^undem", e$dummy_name)
  data.frame(group = g, e[rows, c("dummy_name", "theta", "se")],
             row.names = NULL)
}))
```

The ability to read subgroup-specific $\hat\theta$ off the same fit
--- without re-estimating on subsets --- is the operational advantage
of the structural model over stacked subgroup AMCE regressions.

## Preference clustering

A complementary view on latent preference types: run k-means on the
per-respondent $\hat\beta(\mathbf Z_i)$ matrix to find clusters of
respondents with similar preference profiles.

```{r}
#| label: gs-clusters
cl_gs <- sc_clusters(fit_gs, k = 3L, seed = 2024)
cl_gs$estimate$sizes
round(cl_gs$estimate$centers, 2)
```

## Visualization

### AMCE coefficient plot

```{r}
#| label: gs-amce-plot
#| fig-height: 5
theta_gs <- coef(fit_gs)
se_gs <- sqrt(diag(vcov(fit_gs)))
df_amce_gs <- data.frame(
  dummy = names(theta_gs), estimate = theta_gs,
  ci_lo = theta_gs - 1.96 * se_gs, ci_hi = theta_gs + 1.96 * se_gs,
  row.names = NULL
)

ggplot(df_amce_gs, aes(x = estimate, y = dummy)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  color = "#E41A1C", size = 0.4) +
  labs(x = expression(hat(theta)[k]), y = NULL,
       title = "Population-average AMCE") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())
```

### Fraction favor / oppose

```{r}
#| label: gs-fraction-bar
#| fig-height: 5
frac_gs <- sc_fraction_preferring(fit_gs, threshold = 0)
df_frac_gs <- frac_gs$estimate
df_long_gs <- rbind(
  data.frame(dummy = df_frac_gs$dummy_name, value = df_frac_gs$frac_positive,
             direction = "Favor", stringsAsFactors = FALSE),
  data.frame(dummy = df_frac_gs$dummy_name, value = -df_frac_gs$frac_negative,
             direction = "Oppose", stringsAsFactors = FALSE)
)

ggplot(df_long_gs, aes(x = value, y = dummy, fill = direction)) +
  geom_col() +
  geom_vline(xintercept = 0, color = "gray30") +
  scale_fill_manual(values = c(Favor = "#377EB8", Oppose = "#E41A1C")) +
  labs(x = "Fraction of respondents", y = NULL, fill = NULL,
       title = "Fraction favoring / opposing each level") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        legend.position = "bottom")
```

### Beta ridgelines

```{r}
#| label: gs-ridgelines
#| fig-height: 6
plot(fit_gs, "beta_ridgelines")
```

### Loss trace

```{r}
#| label: gs-loss
plot(fit_gs, "loss_trace")
```
