---
title: "Example: Democratic Norms (GS)"
---

```{r}
#| label: setup-gs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
```

This chapter applies the structural estimator to a candidate-choice
conjoint whose design follows Graham and Svolik (2020). Respondents
choose between pairs of hypothetical U.S. candidates; one attribute
captures the candidate's willingness to engage in undemocratic behavior
(ignoring a court ruling, restricting the opposition press, or banning
an opposition rally). The original study asks how much voters penalize
co-partisan candidates for violating democratic norms, and whether
strong partisanship leads to tolerance of such violations.

The structural model adds value here because it decomposes the
population-average penalty for undemocratic behavior into direction and
intensity, and traces the heterogeneity back to respondent-level
partisanship --- without re-estimating on subsets.

::: {.callout-note}
The `gs2020` dataset is a **seed-fixed synthetic fixture** for
illustration. Do not use it to draw substantive conclusions about the
Graham--Svolik population.
:::

## Data preparation

```{r}
#| label: gs-load
data(gs2020, package = "sconjoint")
dim(gs2020)
head(gs2020, 4)
str(gs2020)
```

Each respondent sees eight tasks with two candidate profiles each,
yielding 16 rows per respondent. The respondent-level moderators are
`resp_pid` (-1/0/+1 partisanship), `resp_ideo` (standardized ideology),
and standardized `age`.

## Fitting the structural model

```{r}
#| label: gs-fit
fit_gs <- scfit(
  choice ~ federal + immigration + tax + abortion + undem + cand_party |
           resp_pid + resp_ideo + age,
  data        = gs2020,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_gs)
```

## Heterogeneity test

Does the per-attribute dispersion of $\hat\beta_k(\mathbf Z_i)$ exceed
what sampling noise alone would explain? A significant result for
`undem` levels would confirm that the penalty for undemocratic behavior
varies meaningfully across respondents.

```{r}
#| label: gs-het
het_gs <- sc_heterogeneity_test(fit_gs, adjust = "bh")
het_gs$estimate
```

## Direction and intensity

The undemocratic levels should show a large negative average direction
(voters penalize them) coupled with high intensity. The structural
question is whether the *dispersion* of those coefficients concentrates
on co-partisans, as Graham and Svolik argue.

```{r}
#| label: gs-di
sc_direction_intensity(fit_gs)
```

## Subgroup analysis by partisanship

We re-average the held-out $\hat\beta(\mathbf Z_i)$ separately for
Democrats, Independents, and Republicans using `sc_subgroup()`. The
`resp_pid` column on the standardized $\mathbf Z$ matrix preserves the
sign of the underlying -1/0/+1 encoding.

```{r}
#| label: gs-subgroup
pid_col <- fit_gs$Z[, "resp_pid"]
sub <- sc_subgroup(fit_gs, list(
  dem = pid_col < -0.5,
  ind = abs(pid_col) <= 0.5,
  rep = pid_col >  0.5
))
do.call(rbind, lapply(names(sub), function(g) {
  e <- sub[[g]]$estimate
  rows <- grepl("^undem", e$dummy_name)
  data.frame(group = g, e[rows, c("dummy_name", "theta", "se")],
             row.names = NULL)
}))
```

The ability to read subgroup-specific $\hat\theta$ off the same fit
--- without re-estimating on subsets --- is the operational advantage
of the structural model over stacked subgroup AMCE regressions.

## Preference clustering

A complementary view on latent preference types: run k-means on the
per-respondent $\hat\beta(\mathbf Z_i)$ matrix to find clusters of
respondents with similar preference profiles.

```{r}
#| label: gs-clusters
cl_gs <- sc_clusters(fit_gs, k = 3L, seed = 2024)
cl_gs$estimate$sizes
round(cl_gs$estimate$centers, 2)
```

## Visualization

```{r}
#| label: gs-ridgelines
#| fig-height: 6
plot(fit_gs, "beta_ridgelines")
```

```{r}
#| label: gs-loss
plot(fit_gs, "loss_trace")
```
