# Example: Candidate Choice {#sec-sw}

```{r}
#| label: setup-sw
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)
```

This chapter walks through a complete `sconjoint` analysis of a
candidate-choice conjoint experiment whose design follows
@saha2022welfare. Respondents evaluate pairs of hypothetical
political candidates varying on five attributes and select one. The
structural decomposition into direction and intensity lets us ask
whether a near-zero AMCE for candidate gender masks genuine
heterogeneity --- half the sample preferring women, half preferring men.

::: {.callout-note}
The `sw2022` dataset is a **seed-fixed synthetic fixture** for
illustration. Do not use it to draw substantive inferences about the
Saha--Weeks population.
:::

## Data preparation

```{r}
#| label: sw-load
data(sw2022, package = "sconjoint")
dim(sw2022)
head(sw2022, 4)
```

The data are in long format: one row per respondent--task--profile. Each
respondent contributes three forced-choice tasks; within a task two
candidate profiles are presented and exactly one is chosen
(`choice == 1`).

```{r}
#| label: sw-structure
str(sw2022)
table(table(sw2022$respondent))
```

The respondent-level moderators available in $\mathbf Z$ are
`resp_female` (binary), `age` (standardized), and `pid` (3-level
partisanship index). The formula interface puts conjoint attributes to
the left of `|` and moderators to the right:

```
choice ~ agenda + talent + children + cand_gender + prior_office |
         resp_female + age + pid
```

## Fitting the structural model

We fit a small, fast configuration (`K = 3` folds, 60 Adam epochs)
suitable for pedagogy. Production analyses would use more folds and more
epochs.

```{r}
#| label: sw-fit
fit_sw <- scfit(
  choice ~ agenda + talent + children + cand_gender + prior_office |
           resp_female + age + pid,
  data        = sw2022,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_sw)
```

The summary table reports the DML point estimate ($\hat\theta$),
respondent-clustered standard error, z-statistic, p-value, and 95%
confidence interval for every attribute dummy. The **DML/iid SE ratio**
diagnostic quantifies how much the clustering correction matters: values
meaningfully above one indicate economically large within-respondent
correlation.

## Interpretation

The `coef()` method extracts the population-average $\hat\theta$ vector.
`predict()` returns the $N \times p$ matrix of per-respondent
$\hat\beta(\mathbf Z_i)$ --- the complete preference vector for each
respondent.

```{r}
#| label: sw-coef
coef(fit_sw)
```

```{r}
#| label: sw-beta
beta <- predict(fit_sw)
dim(beta)
head(beta, 3)
```

Each row of the $\hat\beta$ matrix is a respondent's full marginal
utility vector across all attribute dummies. The structural model's
payoff is that all downstream quantities are functions of this matrix.

## Structural quantities

### Attribute importance

Which attributes drive choice, on average? `sc_importance()` returns
the respondent-averaged share of predicted-utility variance contributed
by each attribute block. Shares sum to one by construction.

```{r}
#| label: sw-importance
sc_importance(fit_sw)
```

### Direction and intensity

The structural decomposition separates *direction*
($\overline{\operatorname{sign}(\beta_k)}$) from *intensity*
($\overline{|\beta_k|}$). For `cand_gender` a near-zero AMCE could
either mean "no one cares" or "half the sample strongly prefers women,
half strongly prefers men and they cancel." The direction/intensity
decomposition resolves the ambiguity.

```{r}
#| label: sw-di
sc_direction_intensity(fit_sw)
```

### Fraction preferring

The fraction of respondents whose $\hat\beta_k(\mathbf Z_i)$ exceeds
zero --- i.e., who prefer each non-reference level, holding other
attributes fixed.

```{r}
#| label: sw-fraction
frac <- sc_fraction_preferring(fit_sw, threshold = 0)
frac$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

### Marginal rate of substitution

The trimmed mean of $\hat\beta_j(\mathbf Z_i) / \hat\beta_k(\mathbf Z_i)$
across respondents: by how many units of attribute $k$ must we
compensate a respondent to offset a one-unit change in attribute $j$?

```{r}
#| label: sw-mrs
sc_mrs(fit_sw, numerator = 1L, denominator = 2L)
```

::: {.callout-tip}
### Large trimmed mean?

The ratio $\hat\beta_j / \hat\beta_k$ is unbounded when the
denominator is near zero. The default trimming at the 1st and 99th
percentiles handles ordinary outliers, but a pathological fit can
still produce a large trimmed mean. Inspect
`plot(fit, "beta_ridgelines")` to confirm the denominator column is
well-separated from zero before interpreting the MRS quantitatively.
:::

### Counterfactual choice probability

For two user-specified profiles $A$ and $B$, report the average
probability that respondents choose $A$ over $B$.

```{r}
#| label: sw-counterfactual
sc_counterfactual(
  fit_sw,
  A = list(agenda = "progressive", talent = "experienced",
           children = "has_children", cand_gender = "female",
           prior_office = "us_house"),
  B = list(agenda = "conservative", talent = "average",
           children = "no_children", cand_gender = "male",
           prior_office = "none")
)
```

### Optimal profile

The greedy per-attribute argmax over subgroup-averaged $\hat\beta$,
assembled into the profile that maximizes average predicted utility.

```{r}
#| label: sw-optimal
sc_optimal_profile(fit_sw)
```

### Heterogeneity test

Does the per-attribute dispersion of $\hat\beta_k(\mathbf Z_i)$ exceed
what sampling noise alone would explain? A significant result indicates
genuine preference heterogeneity along $\mathbf Z$.

```{r}
#| label: sw-het
sc_heterogeneity_test(fit_sw, adjust = "bh")
```

### Subgroup analysis

Do women respondents weight candidate gender differently from men?
`sc_subgroup()` re-averages $\hat\beta(\mathbf Z_i)$ over a
user-specified subset of respondents.

```{r}
#| label: sw-subgroup
fem <- fit_sw$Z[, "resp_female"] > 0.5
sub_fem  <- sc_subgroup(fit_sw, fem)
sub_male <- sc_subgroup(fit_sw, !fem)
rbind(
  female = sub_fem$estimate[sub_fem$estimate$dummy_name == "cand_genderfemale",
                            c("theta", "se")],
  male   = sub_male$estimate[sub_male$estimate$dummy_name == "cand_genderfemale",
                            c("theta", "se")]
)
```

On this fixture the *direction* of the `cand_genderfemale` coefficient
splits along respondent gender lines even when its average AMCE is
small. This is the headline interpretive payoff of the structural
framework: one number becomes an entire distribution over respondents,
indexed by observable $\mathbf Z$.

## Advanced quantities

### Average marginal effects (probability scale)

Converting the logit-scale $\hat\theta$ to probability-scale AMEs
accounts for the nonlinearity of the logistic link.

```{r}
#| label: sw-ame-prob
sc_average(fit_sw, scale = "probability")
```

The probability-scale AMEs are attenuated relative to the logit-scale
coefficients because the logistic CDF compresses effects away from the
linear region. Attributes with large logit-scale effects show the
largest attenuation.

### Preference inequality

Which attributes generate the most disagreement across respondents?

```{r}
#| label: sw-inequality
sc_inequality(fit_sw, measure = "variance")
```

Higher variance indicates more heterogeneous preferences. Attributes
where the heterogeneity test is significant should also show large
variance here --- the two diagnostics measure the same phenomenon
from different angles.

### Baseline logit comparison

The homogeneous logit baseline should agree with the structural model
on population-average $\hat\theta$ while lacking the ability to
decompose heterogeneity.

```{r}
#| label: sw-baseline
bl_sw <- sc_baseline_logit(fit_sw)
data.frame(
  dummy     = fit_sw$attr_names,
  structural = round(coef(fit_sw), 3),
  logit      = round(coef(bl_sw), 3),
  row.names  = NULL
)
```

The structural and baseline logit estimates agree on population
averages. The structural model's advantage is the per-respondent
$\hat\beta(\mathbf Z_i)$ matrix that enables all the heterogeneity,
subgroup, and counterfactual analyses above --- the logit baseline
cannot provide any of those.

## Visualization

### AMCE coefficient plot

```{r}
#| label: sw-amce-plot
#| fig-height: 5
plot_amce(fit_sw)
```

### Fraction favor / oppose

```{r}
#| label: sw-fraction-bar
#| fig-height: 5
plot_fraction(fit_sw)
```

### Heterogeneity bar chart

Which attributes exhibit significant preference heterogeneity across
respondents? Bars are colored red when the heterogeneity test rejects
at the 5% level.

```{r}
#| label: sw-het-bar
#| fig-height: 5
plot_hetero(fit_sw)
```

### Subgroup AMCE: gender

```{r}
#| label: sw-subgroup-plot
#| fig-height: 5
plot_subgroup(
  fit_sw,
  subgroup = list(Female = fit_sw$Z[, "resp_female"] > 0.5,
                  Male   = fit_sw$Z[, "resp_female"] <= 0.5),
  title = "Subgroup AMCE: female vs male respondents"
)
```

### Beta ridgelines

```{r}
#| label: sw-ridgelines
#| fig-height: 6
plot(fit_sw, "beta_ridgelines")
```

### Loss trace

```{r}
#| label: sw-loss
plot(fit_sw, "loss_trace")
```
