# Example: Candidate Choice {#sec-sw}

```{r}
#| label: setup-sw
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Groups and labels for plot annotations --------------------------------
sw_groups <- c(
  "agendaModerate Changes"      = "Agenda",
  "agendaComplete Overhaul"     = "Agenda",
  "talentCollaborative"         = "Talent",
  "talentDetermined to Succeed" = "Talent",
  "talentEmpathetic"            = "Talent",
  "talentGood Communicator"     = "Talent",
  "talentHard-Working"          = "Talent",
  "talentTough Negotiator"      = "Talent",
  "children1 child"             = "Children",
  "children2 children"          = "Children",
  "children3 children"          = "Children",
  "cand_genderMale"             = "Gender",
  "prior_officeYes"             = "Prior Office"
)

sw_labels <- c(
  "agendaModerate Changes"      = "Moderate Changes",
  "agendaComplete Overhaul"     = "Complete Overhaul",
  "talentCollaborative"         = "Collaborative",
  "talentDetermined to Succeed" = "Determined to Succeed",
  "talentEmpathetic"            = "Empathetic",
  "talentGood Communicator"     = "Good Communicator",
  "talentHard-Working"          = "Hard-Working",
  "talentTough Negotiator"      = "Tough Negotiator",
  "children1 child"             = "1 child",
  "children2 children"          = "2 children",
  "children3 children"          = "3 children",
  "cand_genderMale"             = "Male",
  "prior_officeYes"             = "Yes"
)
```

This chapter walks through a complete `sconjoint` analysis of a
candidate-choice conjoint experiment whose design follows
@saha2022welfare. Respondents evaluate pairs of hypothetical
political candidates varying on five attributes and select one. The
structural decomposition into direction and intensity lets us ask
whether a near-zero AMCE for candidate gender masks genuine
heterogeneity --- half the sample preferring women, half preferring men.

## Data preparation

Load the bundled Saha-Weeks subset and inspect its structure.

```{r}
#| label: sw-load
data(sw2022, package = "sconjoint")
dim(sw2022)
head(sw2022, 4)
```

The data are in long format: one row per respondent--task--profile. Each
respondent contributes three forced-choice tasks; within a task two
candidate profiles are presented and exactly one is chosen
(`choice == 1`).

```{r}
#| label: sw-structure
str(sw2022)
table(table(sw2022$respondent))
```

For consistency with the paper, we re-level `cand_gender` so that
**Female is the reference and Male is the non-reference dummy**. The
paper reports the Male-relative-to-Female coefficient throughout.

```{r}
#| label: sw-relevel
sw2022$cand_gender <- stats::relevel(factor(sw2022$cand_gender),
                                     ref = "Female")
levels(sw2022$cand_gender)
```

The respondent-level moderators available in $\mathbf Z$ are
`resp_female` (binary), `age` (standardized), and `pid` (3-level
partisanship index). The formula interface puts conjoint attributes to
the left of `|` and moderators to the right:

```
choice ~ agenda + talent + children + cand_gender + prior_office |
         resp_female + age + pid
```

## Fitting

We fit a moderate configuration (`K = 5` folds, 200 Adam epochs) that
balances render speed against paper-quality estimates. A production run
would use `K = 10` and `n_epochs >= 1000`. The loss trace below shows
the per-fold training curves; with `K = 5` cross-fitting the loss has
largely plateaued by epoch 150-200.

```{r}
#| label: sw-fit
fit_sw <- scfit(
  choice ~ agenda + talent + children + cand_gender + prior_office |
           resp_female + age + pid,
  data        = sw2022,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 5L,
  n_epochs    = 200L,
  seed        = 2024
)
summary(fit_sw)
```

::: {.callout-note}
### The Stage-2 default

As of v0.2, `scfit()` runs an empirical-Bayes MAP refinement (paper
`EnsC5`) on top of the Stage-1 DNN by default
(`stage2 = "map_c5"`). The summary above reports this as
`Stage 2: map_c5`. DML point estimates and clustered standard errors
are unchanged from v0.1 — they continue to use the Stage-1 single-DNN
prediction (now stored on `fit_sw$beta_hat_dnn`). The refined betas
(`fit_sw$beta_hat`) feed every distributional and individual-level
quantity below. Set `stage2 = "none"` to recover v0.1 behavior, or
`stage2 = "mixed_logit"` for the BLUP alternative from paper §A.4.
:::

## Population-average estimates

The DML point estimates $\hat\theta_k$ on the logit scale. Inspect
numerically, then visualize as a coefficient plot grouped by attribute.

```{r}
#| label: sw-coef
coef(fit_sw)
```

```{r}
#| label: sw-amce-plot
#| fig-height: 6
plot_amce(fit_sw, groups = sw_groups, labels = sw_labels)
```

### Structural AME vs reduced-form AMCE (paper Figure 1.B)

The paper notes that the structural model's average marginal effect
(AME) on the *probability* scale nearly coincides with the standard
linear-probability-model AMCE --- confirming that the structural
estimator nests reduced-form AMCE analysis. Below we compute both and
overlay them as point-ranges on the probability scale.

```{r}
#| label: sw-ame-vs-amce
#| fig-height: 6
## Structural AME on the probability scale (delta-method via plogis')
ame_struct <- sc_average(fit_sw, scale = "probability")
ame_df     <- ame_struct$estimate
ame_df$source <- "Structural AME"

## Linear-probability-model AMCE
lpm        <- sc_baseline_lpm(fit_sw)
amce_df    <- data.frame(
  dummy_name = names(stats::coef(lpm)),
  estimate   = unname(stats::coef(lpm)),
  se         = sqrt(diag(stats::vcov(lpm))),
  source     = "LPM AMCE",
  stringsAsFactors = FALSE
)
amce_df$ci_lo <- amce_df$estimate - 1.96 * amce_df$se
amce_df$ci_hi <- amce_df$estimate + 1.96 * amce_df$se

df_both <- rbind(
  ame_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")],
  amce_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")]
)
df_both$label <- ifelse(df_both$dummy_name %in% names(sw_labels),
                        sw_labels[df_both$dummy_name],
                        df_both$dummy_name)
df_both$label <- factor(df_both$label, levels = rev(unname(sw_labels)))
df_both$group <- sw_groups[df_both$dummy_name]
df_both$group <- factor(df_both$group, levels = unique(sw_groups))

ggplot(df_both, aes(x = estimate, y = label, color = source)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  position = position_dodge(width = 0.5),
                  size = 0.35, fatten = 2) +
  scale_color_manual(values = c("Structural AME" = "#2166AC",
                                "LPM AMCE"        = "#B2182B")) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Effect on Pr(choice)",
       y = NULL, color = NULL,
       title = "Structural AME vs LPM AMCE (probability scale)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        strip.text.y    = element_text(angle = 0, hjust = 0, face = "bold"))
```

The two series should overlap nearly point-for-point ---
the structural model loses nothing the reduced-form AMCE delivers.

## Individual-level preferences

Pull the respondent-level $\hat\beta(\mathbf Z_i)$ matrix. With the v0.2
default `stage2 = "map_c5"`, this is the MAP-refined hybrid view.

```{r}
#| label: sw-beta
beta <- predict(fit_sw)
dim(beta)
head(beta, 3)
```

Ridgeline densities show the distribution of $\hat\beta_k(\mathbf Z_i)$
per dummy across respondents. The vertical dashed line is at zero; the
solid line in each density is the median.

```{r}
#| label: sw-ridgelines
#| fig-height: 8
plot(fit_sw, "beta_ridgelines", groups = sw_groups, labels = sw_labels)
```

## Structural quantities

### Attribute importance

Variance-decomposition share of utility carried by each attribute
group, averaged across respondents. The "importance" answers: which
attribute moves the choice the most?

```{r}
#| label: sw-importance
sc_importance(fit_sw)
```

```{r}
#| label: sw-importance-plot
#| fig-height: 5
plot_importance(fit_sw, labels = c(agenda = "Agenda", talent = "Talent",
  children = "Children", cand_gender = "Gender", prior_office = "Prior Office"))
```

Policy agenda dominates, accounting for approximately two-thirds of the
decision variance for the median voter.

### Direction and intensity

For each dummy, decompose preference into a *direction* (the sign of
$\hat\beta_{ik}$) and an *intensity* (its magnitude). Near-zero average
effects can hide either consensus (everyone close to zero) or
polarization (large opposing intensities cancelling).

```{r}
#| label: sw-di
sc_direction_intensity(fit_sw)
```

For the Empathetic attribute, the near-zero average masks an almost
perfectly polarized electorate: approximately half favor empathetic
candidates while the other half oppose them.

### Fraction preferring

What share of respondents have a positive coefficient on each dummy?
This is the direction-only summary --- ignores intensity.

```{r}
#| label: sw-fraction
frac <- sc_fraction_preferring(fit_sw, threshold = 0)
frac$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

```{r}
#| label: sw-fraction-bar
#| fig-height: 6
plot_fraction(fit_sw, groups = sw_groups, labels = sw_labels)
```

The paper's directional finding for gender: the population-average
AMCE for Male is near zero, but the individual-level fractions reveal
that about 64% of respondents have $\hat\beta_{\text{Male}} < 0$,
i.e.\ prefer Female candidates. The fraction-preferring panel above
makes this visible per dummy. The exact percentage depends on training
configuration; the qualitative finding survives down to the
lighter pedagogical settings.

### Marginal rate of substitution

The respondent-averaged trimmed ratio $\hat\beta_j / \hat\beta_k$
between two dummies, with respondent-clustered SE. Trimming at
$\{0.01, 0.99\}$ guards the unbounded ratio when the denominator is
near zero.

```{r}
#| label: sw-mrs
sc_mrs(fit_sw, numerator = 1L, denominator = 2L)
```

::: {.callout-tip}
### Large trimmed mean?

The ratio $\hat\beta_j / \hat\beta_k$ is unbounded when the
denominator is near zero. Inspect the ridgelines above to confirm the
denominator column is well-separated from zero.
:::

### Counterfactual choice probability

Predicted choice probability for a *specific* head-to-head profile pair,
averaged across respondents. Each profile is a named list of attribute
levels; unmentioned attributes default to the reference level.

```{r}
#| label: sw-counterfactual
sc_counterfactual(
  fit_sw,
  ## Profile A: a Female Hard-Working candidate proposing a Complete
  ## Overhaul, against B: a Male Assertive candidate proposing the
  ## minimal change.  Female is the reference level so it does not
  ## appear in the dummy contrast; Male contributes +beta_Male to A's
  ## utility (negative in this design).
  A = list(agenda = "Complete Overhaul", talent = "Hard-Working",
           children = "2 children", cand_gender = "Female",
           prior_office = "Yes"),
  B = list(agenda = "Very Few Changes", talent = "Assertive",
           children = "No children", cand_gender = "Male",
           prior_office = "No")
)
```

### Optimal profile

The single profile that maximizes the population-average utility (greedy
attribute-by-attribute selection over the discrete attribute levels).

```{r}
#| label: sw-optimal
sc_optimal_profile(fit_sw)
```

### Heterogeneity test

For each dummy, the one-sided test that $\mathrm{Var}(\beta_k(Z)) > 0$
exceeds zero. Useful for screening which attributes genuinely vary
across respondents.

```{r}
#| label: sw-het
sc_heterogeneity_test(fit_sw, adjust = "bh")
```

```{r}
#| label: sw-het-bar
#| fig-height: 6
plot_hetero(fit_sw, groups = sw_groups, labels = sw_labels)
```

### Subgroup analysis

Re-average the respondent-level betas over a subset of respondents,
then compare across subsets. Below: compare the per-attribute means for
female vs male respondents on the gender attribute.

```{r}
#| label: sw-subgroup
fem <- fit_sw$Z[, "resp_female"] > 0.5
sub_fem  <- sc_subgroup(fit_sw, fem)
sub_male <- sc_subgroup(fit_sw, !fem)
rbind(
  female = sub_fem$estimate[sub_fem$estimate$dummy_name == "cand_genderMale",
                            c("theta", "se")],
  male   = sub_male$estimate[sub_male$estimate$dummy_name == "cand_genderMale",
                            c("theta", "se")]
)
```

```{r}
#| label: sw-subgroup-plot
#| fig-height: 6
plot_subgroup(
  fit_sw,
  subgroup = list(Female = fit_sw$Z[, "resp_female"] > 0.5,
                  Male   = fit_sw$Z[, "resp_female"] <= 0.5),
  groups = sw_groups, labels = sw_labels,
  title = "Subgroup AMCE: female vs male respondents"
)
```

### Loss trace

```{r}
#| label: sw-loss
plot(fit_sw, "loss_trace")
```

### Validating against the homogeneous-logit AMCE {#sec-sw-validate}

The paper's Appendix D shows that the structural model nests the
standard reduced-form AMCE: the DML population-average parameter
$\theta_k = \mathbb{E}[\beta_k(Z)]$ equals the pooled
homogeneous-logit coefficient on attribute $k$ when the logit
specification holds. The `sc_validate_amce()` export turns this into
a one-line sanity check.

```{r}
#| label: sw-validate
v <- sc_validate_amce(fit_sw)
v
```

In a full paper-quality fit (`K = 10`, `n_epochs >= 1000`), the
correlation across attribute levels is essentially $r = 0.998$. Any
material drop is a sign of either insufficient training or a
genuinely non-logit data-generating process.

### Comparing the Stage-2 view to the raw DNN view

Every `sc_*` quantity now accepts a `which_beta` argument with
default `"hybrid"` (Stage-2 refined). Passing `"dnn"` reads the
unrefined Stage-1 DNN mean, which lives in `fit_sw$beta_hat_dnn`.

```{r}
#| label: sw-which-beta
frac_hyb <- sc_fraction_preferring(fit_sw, which_beta = "hybrid")
frac_dnn <- sc_fraction_preferring(fit_sw, which_beta = "dnn")
data.frame(
  attribute   = frac_hyb$estimate$dummy_name,
  hybrid_pref = round(frac_hyb$estimate$frac_positive, 3),
  dnn_pref    = round(frac_dnn$estimate$frac_positive, 3)
)
```

The Stage-2 refinement typically shifts the more polarized columns
(e.g. `cand_genderMale`, `talentEmpathetic`) toward smaller
fractions and the consensus columns (`talentHard-Working`,
`agendaModerate Changes`) toward 100% --- this is the within-respondent
shrinkage at work.
