# Example: Democratic Norms {#sec-gs}

```{r}
#| label: setup-gs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Groups and labels for plot annotations (keyed by the diff_ contrast
## column names; reference level = good-governance "elect oversight board",
## already dropped in the bundled data). ------------------------------------
gs_groups <- c(
  "diff_respParty"                  = "Party",
  "diff_p1_num"                     = "Policy",
  "diff_p2_num"                     = "Policy",
  "diff_dem_code_g_committee"       = "Good governance",
  "diff_dem_code_g_officestructure" = "Good governance",
  "diff_dem_code_g_procedure"       = "Good governance",
  "diff_dem_code_g_progEval"        = "Good governance",
  "diff_dem_code_g_record"          = "Good governance",
  "diff_dem_code_g_schedule"        = "Good governance",
  "diff_dem_code_u_banProtest"      = "Undemocratic",
  "diff_dem_code_u_court"           = "Undemocratic",
  "diff_dem_code_u_execRule"        = "Undemocratic",
  "diff_dem_code_u_gerry2"          = "Undemocratic",
  "diff_dem_code_u_gerry10"         = "Undemocratic",
  "diff_dem_code_u_journalists"     = "Undemocratic",
  "diff_dem_code_u_limitVote"       = "Undemocratic",
  "diff_dem_code_v_affair"          = "Valence",
  "diff_dem_code_v_tax"             = "Valence",
  "diff_sex_Female"                 = "Sex",
  "diff_race_Asian"                 = "Race",
  "diff_race_Black"                 = "Race",
  "diff_race_Hispanic"              = "Race",
  "diff_pro_Farmer"                 = "Profession",
  "diff_pro_Lawyer"                 = "Profession",
  "diff_pro_Legislative_staffer"    = "Profession",
  "diff_pro_Police_officer"         = "Profession",
  "diff_pro_Served_in_the_army"     = "Profession",
  "diff_pro_Served_in_the_navy"     = "Profession",
  "diff_pro_Small_business_owner"   = "Profession",
  "diff_pro_Teacher"                = "Profession"
)

gs_labels <- c(
  "diff_respParty"                  = "Co-partisan",
  "diff_p1_num"                     = "Econ. policy (cons.)",
  "diff_p2_num"                     = "Social policy (cons.)",
  "diff_dem_code_g_committee"       = "Committee oversight",
  "diff_dem_code_g_officestructure" = "Office structure",
  "diff_dem_code_g_procedure"       = "Legislative procedure",
  "diff_dem_code_g_progEval"        = "Program evaluation",
  "diff_dem_code_g_record"          = "Record-keeping",
  "diff_dem_code_g_schedule"        = "Legislative schedule",
  "diff_dem_code_u_banProtest"      = "Ban protests",
  "diff_dem_code_u_court"           = "Ignore courts",
  "diff_dem_code_u_execRule"        = "Executive orders",
  "diff_dem_code_u_gerry2"          = "Gerrymander (2 seats)",
  "diff_dem_code_u_gerry10"         = "Gerrymander (10 seats)",
  "diff_dem_code_u_journalists"     = "Restrict journalists",
  "diff_dem_code_u_limitVote"       = "Limit voting",
  "diff_dem_code_v_affair"          = "Extramarital affair",
  "diff_dem_code_v_tax"             = "Underpaid taxes",
  "diff_sex_Female"                 = "Female",
  "diff_race_Asian"                 = "Asian",
  "diff_race_Black"                 = "Black",
  "diff_race_Hispanic"              = "Hispanic",
  "diff_pro_Farmer"                 = "Farmer",
  "diff_pro_Lawyer"                 = "Lawyer",
  "diff_pro_Legislative_staffer"    = "Legislative staffer",
  "diff_pro_Police_officer"         = "Police officer",
  "diff_pro_Served_in_the_army"     = "Army veteran",
  "diff_pro_Served_in_the_navy"     = "Navy veteran",
  "diff_pro_Small_business_owner"   = "Small business owner",
  "diff_pro_Teacher"                = "Teacher"
)
```

This chapter applies the structural estimator to a candidate-choice
conjoint whose design follows @graham2020democracy. Respondents
choose between pairs of hypothetical U.S. candidates; one attribute
captures the candidate's willingness to engage in undemocratic behavior
(ignoring a court ruling, restricting the opposition press, or banning
an opposition rally). The original study asks how much voters penalize
co-partisan candidates for violating democratic norms, and whether
strong partisanship leads to tolerance of such violations.

The structural model adds value here because it decomposes the
population-average penalty for undemocratic behavior into direction and
intensity, and traces the heterogeneity back to respondent-level
partisanship --- without re-estimating on subsets.

The bundled `gs2020` data is the paper's full analysis frame, and this
chapter reproduces the paper's **main (held-out) specification**: the
16-covariate moderator set that drops the six direct
democracy-attitude items (see [Fitting the structural model](#fitting),
below). The numbers reported here match the paper's Section 4.2.

## Data preparation

Load the Graham-Svolik analysis frame and inspect its structure. Each
respondent contributes about 13 forced-choice tasks against pairs of
hypothetical candidates.

```{r}
#| label: gs-load
data(gs2020, package = "sconjoint")
dim(gs2020)
gs2020[1:4, c("respondent", "task", "profile", "choice",
              "diff_respParty", "diff_dem_code_u_journalists")]
```

The frame is already in the paper's two-profile parameterization: the
choice outcome, the identifiers (`respondent`, `task`, `profile`), 30
attribute contrasts (the `diff_` columns, differenced across the two
profiles), and the respondent-level moderators $\mathbf{Z}$ (the `z_*`
and `*_ideal` columns plus six direct democracy-attitude items).
Because the design is differenced, the reference levels are already
absorbed (good governance is measured against an "elect oversight
board" baseline; the undemocratic and valence items against "absent"
and "no scandal"). No re-leveling is needed.

The 15 undemocratic, good-governance, and valence contrasts split into
three families: good-governance (`diff_dem_code_g_*`, 6 contrasts),
undemocratic actions (`diff_dem_code_u_*`, 7 contrasts), and valence
violations (`diff_dem_code_v_*`, 2 contrasts). Their raw conditional
choice rates make the families clear: candidates who endorse an
undemocratic action or carry a valence scandal are chosen far less
often.

```{r}
#| label: gs-dem-code-raw
u_cols <- grep("^diff_dem_code_u_", names(gs2020), value = TRUE)
v_cols <- grep("^diff_dem_code_v_", names(gs2020), value = TRUE)
## Mean choice rate for the profile carrying each contrast (diff == 1)
rate_when_present <- function(col) {
  mean(gs2020$choice[gs2020[[col]] == 1])
}
round(sort(sapply(c(u_cols, v_cols), rate_when_present)), 3)
```

Every undemocratic and valence contrast is chosen well below 0.5 when
present: voters penalize both democratic backsliding and personal
scandal.

## Fitting the structural model {#fitting}

We fit the paper's main specification: the 30 attribute contrasts in
the first part of the formula, and the **held-out 16-covariate**
moderator set in the second part. The held-out set drops the six direct
democracy-attitude items (`dem_better`, `dem_satisfied`, and the four
`dem_treat_*` items), which ask respondents directly how acceptable
each action is; holding them out of $\mathbf{Z}$ is the more defensible
specification, because it forces the model to recover preferences over
undemocratic actions from the conjoint choices alone rather than from a
near-restatement of the outcome. Attribute names are back-quoted
because they contain characters R would otherwise parse.

```{r}
#| label: gs-fit
#| cache: true
fit_gs <- scfit(
  choice ~ `diff_respParty` + `diff_p1_num` + `diff_p2_num` +
    `diff_dem_code_g_committee` + `diff_dem_code_g_officestructure` +
    `diff_dem_code_g_procedure` + `diff_dem_code_g_progEval` +
    `diff_dem_code_g_record` + `diff_dem_code_g_schedule` +
    `diff_dem_code_u_banProtest` + `diff_dem_code_u_court` +
    `diff_dem_code_u_execRule` + `diff_dem_code_u_gerry2` +
    `diff_dem_code_u_gerry10` + `diff_dem_code_u_journalists` +
    `diff_dem_code_u_limitVote` + `diff_dem_code_v_affair` +
    `diff_dem_code_v_tax` + `diff_sex_Female` + `diff_race_Asian` +
    `diff_race_Black` + `diff_race_Hispanic` + `diff_pro_Farmer` +
    `diff_pro_Lawyer` + `diff_pro_Legislative_staffer` +
    `diff_pro_Police_officer` + `diff_pro_Served_in_the_army` +
    `diff_pro_Served_in_the_navy` + `diff_pro_Small_business_owner` +
    `diff_pro_Teacher` |
    `z_ideo` + `z_pid7` + `z_trump` + `z_age` + `z_educ` + `z_hhi` +
    `z_auth` + `z_knowl` + `z_female` + `z_race_black` + `z_race_asian` +
    `z_race_other` + `E_ideal` + `I_ideal` + `M_ideal` + `T_ideal`,
  data       = gs2020,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 10L,
  seed       = 42L,
  stage2     = "map_c5",
  parallel   = TRUE,
  n_cores    = 10L
)
summary(fit_gs)
```

The population-average penalty for being a co-partisan candidate is a
large, precisely-estimated positive coefficient
($\hat\theta_{\text{co-partisan}} \approx +0.72$ on the logit scale),
and every undemocratic action carries a strongly negative coefficient
(range $-0.57$ to $-0.93$). Restricting journalists is the most heavily
penalized action ($\hat\theta \approx -0.93$); the valence scandals are
even more damaging (extramarital affair $\approx -0.98$).

::: {.callout-note}
### Stage 2 is on by default (v0.2)

The summary above reports `Stage 2: map_c5` --- the empirical-Bayes
MAP refinement of paper §3. The DML population coefficients
$\hat\theta_k$ (the co-partisan and undemocratic-action effects above)
are computed from the Stage-1 cross-fitted DNN; the refined
respondent-level betas (used by the distributional and structural
quantities below) sit on `fit_gs$beta_hat_resp`. Pass
`stage2 = "none"` to recover v0.1 behavior.
:::

```{r}
#| label: gs-loss
plot(fit_gs, "loss_trace")
```

## Population-average estimates

::: {.callout-note}
### How to read the AMCE plot

All `dem_code` contrasts are read relative to their absent / no-scandal
baseline:

* `g_*` (good-governance) items: near zero --- voters are largely
  indifferent across procedural good-governance features.
* `u_*` (undemocratic) items: **strongly negative** --- voters
  penalize every undemocratic action.
* `v_*` (valence violation) items: **most negative** --- voters reject
  affairs and tax evasion most sharply.

This is the paper's directional finding: voters broadly oppose
democratic backsliding.
:::

Plot the population-average coefficients (DML on the logit scale).

```{r}
#| label: gs-amce-plot
#| fig-height: 8
plot_amce(fit_gs, groups = gs_groups, labels = gs_labels)
```

### Structural AME vs reduced-form AMCE

Overlay the structural model's average marginal effects (probability
scale) against the linear-probability-model AMCE. Under correct
specification the two coincide.

```{r}
#| label: gs-ame-vs-amce
#| fig-height: 9
ame_struct <- sc_average(fit_gs, scale = "probability")
ame_df     <- ame_struct$estimate; ame_df$source <- "Structural AME"
lpm        <- sc_baseline_lpm(fit_gs)
amce_df    <- data.frame(
  dummy_name = names(stats::coef(lpm)),
  estimate   = unname(stats::coef(lpm)),
  se         = sqrt(diag(stats::vcov(lpm))),
  source     = "LPM AMCE",
  stringsAsFactors = FALSE
)
amce_df$ci_lo <- amce_df$estimate - 1.96 * amce_df$se
amce_df$ci_hi <- amce_df$estimate + 1.96 * amce_df$se
df_both <- rbind(
  ame_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")],
  amce_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")]
)
df_both$label <- ifelse(df_both$dummy_name %in% names(gs_labels),
                        gs_labels[df_both$dummy_name],
                        df_both$dummy_name)
df_both$label <- factor(df_both$label,
                        levels = rev(intersect(unname(gs_labels), df_both$label)))
df_both$group <- gs_groups[df_both$dummy_name]
df_both$group <- factor(df_both$group, levels = unique(gs_groups))
df_both <- df_both[!is.na(df_both$group), ]

ggplot(df_both, aes(x = estimate, y = label, color = source)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  position = position_dodge(width = 0.5),
                  size = 0.3, fatten = 2) +
  scale_color_manual(values = c("Structural AME" = "#2166AC",
                                "LPM AMCE"        = "#B2182B")) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Effect on Pr(choice)",
       y = NULL, color = NULL,
       title = "Structural AME vs LPM AMCE (probability scale)") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        strip.text.y    = element_text(angle = 0, hjust = 0, face = "bold"))
```

## Individual-level preferences

Ridgeline densities show how each per-respondent coefficient
$\hat\beta_k(\mathbf Z_i)$ is distributed across the sample. The
undemocratic-action densities sit clearly left of zero for nearly every
respondent --- opposition to backsliding is broad, not confined to a
subgroup --- while the valence densities sit furthest left (see the
callout above).

```{r}
#| label: gs-ridgelines
#| fig-height: 16
plot(fit_gs, "beta_ridgelines", groups = gs_groups, labels = gs_labels)
```

## Structural quantities

### Attribute importance

Decompose utility variance by attribute group. The package's
`sc_importance()` reports a per-contrast share; the paper groups the
contrasts into attributes (party, policy, the three democracy families,
and the candidate demographics) and reports the variance share of each
attribute. Computing the grouped shares from the recovered
per-respondent coefficients reproduces the paper's decomposition.

```{r}
#| label: gs-importance
B <- fit_gs$beta_hat_resp
attr_groups <- list(
  party    = "diff_respParty",
  policy   = c("diff_p1_num", "diff_p2_num"),
  good     = grep("_g_", colnames(B), value = TRUE),
  undem    = grep("_u_", colnames(B), value = TRUE),
  valence  = grep("_v_", colnames(B), value = TRUE),
  sex      = "diff_sex_Female",
  race     = grep("^diff_race_", colnames(B), value = TRUE),
  prof     = grep("^diff_pro_", colnames(B), value = TRUE)
)
## Per-respondent utility-variance per attribute over uniform level draws
## (reference contributes beta = 0, so L = #contrasts + 1).
imp <- sapply(attr_groups, function(cols) {
  ba <- B[, cols, drop = FALSE]; L <- ncol(ba) + 1
  rowSums(ba^2) / L - (rowSums(ba) / L)^2
})
shares <- 100 * colMeans(imp / rowSums(imp))
round(sort(shares, decreasing = TRUE), 1)
```

The leading shares are valence ($\approx 29\%$), policy
($\approx 26\%$), and party ($\approx 23\%$) clustered at the top, with
the undemocratic-action attribute ($\approx 9\%$) carrying a smaller but
non-trivial share. Voters care most about scandal, policy, and party;
the democracy dimension matters, but it is not what most differentiates
candidates in this design.

The package's per-contrast view (uniform design) is complementary --- it
shows which individual contrasts within each attribute carry the
variance:

```{r}
#| label: gs-importance-contrast
sc_importance(fit_gs, design = "uniform")$estimate[, c("attribute", "share")]
```

```{r}
#| label: gs-importance-plot
#| fig-height: 8
plot_importance(fit_gs, labels = gs_labels)
```

## Heterogeneity test

Does the per-attribute dispersion of $\hat\beta_k(\mathbf Z_i)$ exceed
what sampling noise alone would explain? A significant result for the
`u_*` levels confirms that the intensity of the penalty for
undemocratic behavior varies meaningfully across respondents.

```{r}
#| label: gs-het
het_gs <- sc_heterogeneity_test(fit_gs, adjust = "bh")
het_gs$estimate
```

```{r}
#| label: gs-het-bar
#| fig-height: 8
plot_hetero(fit_gs, groups = gs_groups, labels = gs_labels)
```

The penalty for each undemocratic action varies in intensity across
respondents, but its population average is strongly negative for every
action: on average, voters oppose democratic backsliding. The
opposition is also broad at the individual level --- the next
section shows that almost no respondents actively favor any undemocratic
action.

### Fraction preferring

```{r}
#| label: gs-fraction
fr <- sc_fraction_preferring(fit_gs)
fr_u <- fr$estimate[grepl("^diff_dem_code_u_", fr$estimate$dummy_name), ]
fr_u$action <- sub("diff_dem_code_u_", "", fr_u$dummy_name)
round(setNames(fr_u$frac_positive, fr_u$action), 3)
```

The fraction of respondents who actively *favor* any undemocratic
action is tiny. The least-opposed action, a 10-seat gerrymander, is
favored by only about 6% of respondents --- over 93% are opposed --- and
restricting journalists, ignoring courts, and limiting voting are
favored by essentially no one. This is the paper's individual-level
finding: opposition to undemocratic behavior is nearly universal, not a
property of one partisan camp.

```{r}
#| label: gs-fraction-bar
#| fig-height: 8
plot_fraction(fit_gs, groups = gs_groups, labels = gs_labels)
```

## Direction and intensity

The undemocratic levels show a large negative average direction (voters
penalize them) coupled with high intensity. The structural question is
whether the *dispersion* of those coefficients concentrates on
co-partisans, as Graham and Svolik argue.

```{r}
#| label: gs-di
sc_direction_intensity(fit_gs)
```

### Marginal rate of substitution: democracy vs party

How much co-partisan advantage would a candidate need to offset the
penalty for an undemocratic action? The population MRS is the ratio of
the action's coefficient to the co-partisan coefficient,
$\theta_{\text{action}} / \theta_{\text{co-partisan}}$, estimated with
the debiased orthogonal score (`estimand = "population"`).

```{r}
#| label: gs-mrs
mrs_actions <- c("diff_dem_code_u_journalists",
                 "diff_dem_code_u_court",
                 "diff_dem_code_u_gerry10")
do.call(rbind, lapply(mrs_actions, function(a) {
  m <- sc_mrs(fit_gs, a, "diff_respParty", estimand = "population")
  data.frame(action = sub("diff_dem_code_u_", "", a),
             abs_mrs = round(abs(m$estimate), 2),
             se      = round(m$se, 2),
             row.names = NULL)
}))
```

Restricting journalists costs about 1.3 co-partisan-advantages to
offset; ignoring courts about 1.0; a 10-seat gerrymander about 0.95. In
words, a candidate would have to be *more* than a co-partisan --- worth
roughly 1.3 party switches --- to make the average voter forgive
prosecuting journalists. Party loyalty does not buy a free pass on the
most severe undemocratic actions.

## Subgroup analysis by partisanship

We re-average the held-out $\hat\beta(\mathbf Z_i)$ separately for
Democrats, Independents, and Republicans using `sc_subgroup()`. The
convenience column `pid7` carries the raw 7-point party identification
($<0$ Democrat, $0$ Independent, $>0$ Republican). `sc_subgroup()`
expects a logical vector aligned to the fit's task rows (one per
respondent--task), so we map the respondent-level `pid7` onto
`fit_gs$respondent_id`.

```{r}
#| label: gs-subgroup-setup
## Respondent-level lookups, indexed onto the fit's task rows.
pid_lookup  <- tapply(gs2020$pid7,  gs2020$respondent, function(x) x[1])
ideo_lookup <- tapply(gs2020$ideo7, gs2020$respondent, function(x) x[1])
rid       <- as.character(fit_gs$respondent_id)
pid_row   <- unname(pid_lookup[rid])    # length = nrow(fit_gs$beta_hat)
ideo_row  <- unname(ideo_lookup[rid])
```

```{r}
#| label: gs-subgroup
sub <- sc_subgroup(fit_gs, list(
  Democrat    = pid_row < 0,
  Independent = pid_row == 0,
  Republican  = pid_row > 0
))
do.call(rbind, lapply(names(sub), function(g) {
  e <- sub[[g]]$estimate
  rows <- grepl("^diff_dem_code_u_", e$dummy_name)
  data.frame(group = g, action = sub("diff_dem_code_u_", "", e$dummy_name[rows]),
             theta = round(e$theta[rows], 2), se = round(e$se[rows], 2),
             row.names = NULL)
}))
```

```{r}
#| label: gs-subgroup-plot
#| fig-height: 5
plot_subgroup(
  fit_gs,
  subgroup = list(Democrat    = pid_row < 0,
                  Independent = pid_row == 0,
                  Republican  = pid_row > 0),
  dummies = grep("^diff_dem_code_u_", fit_gs$attr_names, value = TRUE),
  groups = gs_groups, labels = gs_labels,
  title = "Subgroup AMCE: undemocratic behavior"
)
```

The structural model reveals that strong partisans of both parties
penalize undemocratic behavior; the intensity of the penalty varies
across respondents, but its sign does not. We can sharpen this with
ideology terciles, built from the raw 7-point ideology column `ideo7`
(1--3 Liberal, 4 Moderate, 5--7 Conservative), again indexed onto the
task rows via `ideo_row`:

```{r}
#| label: gs-ideo-tercile
ideo_grp <- cut(ideo_row, c(0, 3, 4, 7),
                labels = c("Liberal", "Moderate", "Conservative"))
sub_ideo <- sc_subgroup(fit_gs, list(
  Liberal      = ideo_grp == "Liberal",
  Moderate     = ideo_grp == "Moderate",
  Conservative = ideo_grp == "Conservative"
))
do.call(rbind, lapply(names(sub_ideo), function(g) {
  e <- sub_ideo[[g]]$estimate
  i <- e$dummy_name == "diff_dem_code_u_journalists"
  data.frame(ideology = g, action = "journalists",
             theta = round(e$theta[i], 2), se = round(e$se[i], 2),
             row.names = NULL)
}))
```

Every ideological tercile assigns a strong negative coefficient to
restricting journalists: the penalty is broad across the ideological
spectrum.

## Preference clustering

A complementary view on latent preference types: run k-means on the
per-respondent $\hat\beta(\mathbf Z_i)$ matrix to find clusters of
respondents with similar preference profiles.

```{r}
#| label: gs-clusters
cl_gs <- sc_clusters(fit_gs, k = 3L, seed = 2024)
cl_gs$estimate$sizes
round(cl_gs$estimate$centers, 2)
```


## Validating against the homogeneous-logit AMCE

The structural model nests the standard pooled-logit AMCE: under
correct logit specification, $\theta_k = \mathbb{E}[\beta_k(Z)]$
equals the homogeneous-logit coefficient on attribute $k$. The
`sc_validate_amce()` export reports this comparison.

```{r}
#| label: gs-validate
v_gs <- sc_validate_amce(fit_gs)
v_gs
```

A correlation near $r = 1$ across the 30 attribute levels confirms
that nothing a reduced-form AMCE analysis would deliver is lost by
using the structural model.
