# Example: Democratic Norms {#sec-gs}

```{r}
#| label: setup-gs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Groups and labels for plot annotations --------------------------------
gs_groups <- c(
  "cand_partyRepublican"      = "Party",
  "dem_codeg_committee"       = "Good Governance",
  "dem_codeg_officestructure" = "Good Governance",
  "dem_codeg_procedure"       = "Good Governance",
  "dem_codeg_progEval"        = "Good Governance",
  "dem_codeg_record"          = "Good Governance",
  "dem_codeg_schedule"        = "Good Governance",
  "dem_codeu_banProtest"      = "Undemocratic",
  "dem_codeu_court"           = "Undemocratic",
  "dem_codeu_execRule"        = "Undemocratic",
  "dem_codeu_gerry2"          = "Undemocratic",
  "dem_codeu_gerry10"         = "Undemocratic",
  "dem_codeu_journalists"     = "Undemocratic",
  "dem_codeu_limitVote"       = "Undemocratic",
  "dem_codev_affair"          = "Valence",
  "dem_codev_tax"             = "Valence",
  "cand_sexFemale"            = "Sex",
  "cand_raceBlack"            = "Race",
  "cand_raceHispanic"         = "Race",
  "cand_raceAsian"            = "Race",
  "policy_prox"               = "Policy"
)

gs_labels <- c(
  "cand_partyRepublican"      = "Republican",
  "dem_codeg_committee"       = "Committee oversight",
  "dem_codeg_officestructure" = "Office structure",
  "dem_codeg_procedure"       = "Procedure",
  "dem_codeg_progEval"        = "Program evaluation",
  "dem_codeg_record"          = "Record-keeping",
  "dem_codeg_schedule"        = "Schedule",
  "dem_codeu_banProtest"      = "Ban protest",
  "dem_codeu_court"           = "Ignore court",
  "dem_codeu_execRule"        = "Executive rule",
  "dem_codeu_gerry2"          = "Gerrymander (2%)",
  "dem_codeu_gerry10"         = "Gerrymander (10%)",
  "dem_codeu_journalists"     = "Restrict journalists",
  "dem_codeu_limitVote"       = "Limit voting",
  "dem_codev_affair"          = "Affair",
  "dem_codev_tax"             = "Tax evasion",
  "cand_sexFemale"            = "Female",
  "cand_raceBlack"            = "Black",
  "cand_raceHispanic"         = "Hispanic",
  "cand_raceAsian"            = "Asian",
  "policy_prox"               = "Policy proximity"
)
```

This chapter applies the structural estimator to a candidate-choice
conjoint whose design follows @graham2020democracy. Respondents
choose between pairs of hypothetical U.S. candidates; one attribute
captures the candidate's willingness to engage in undemocratic behavior
(ignoring a court ruling, restricting the opposition press, or banning
an opposition rally). The original study asks how much voters penalize
co-partisan candidates for violating democratic norms, and whether
strong partisanship leads to tolerance of such violations.

The structural model adds value here because it decomposes the
population-average penalty for undemocratic behavior into direction and
intensity, and traces the heterogeneity back to respondent-level
partisanship --- without re-estimating on subsets.

::: {.callout-note}
The `gs2020` dataset is a **seed-fixed synthetic fixture** for
illustration. Do not use it to draw substantive conclusions about the
Graham--Svolik population.
:::

## Data preparation

```{r}
#| label: gs-load
data(gs2020, package = "sconjoint")
dim(gs2020)
head(gs2020, 4)
str(gs2020)
```

Each respondent sees eight tasks with two candidate profiles each,
yielding 16 rows per respondent. The respondent-level moderators are
`resp_pid` (-1/0/+1 partisanship), `resp_ideo` (standardized ideology),
and standardized `age`.

## Fitting the structural model

```{r}
#| label: gs-fit
fit_gs <- scfit(
  choice ~ cand_party + dem_code + cand_sex + cand_race + policy_prox |
           resp_ideo + resp_pid + resp_age,
  data        = gs2020,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_gs)
```

## Attribute importance

```{r}
#| label: gs-importance
sc_importance(fit_gs)
```

```{r}
#| label: gs-importance-plot
#| fig-height: 5
plot_importance(fit_gs, labels = c(cand_party = "Party", dem_code = "Democracy",
  cand_sex = "Sex", cand_race = "Race", policy_prox = "Policy"))
```

## Heterogeneity test

Does the per-attribute dispersion of $\hat\beta_k(\mathbf Z_i)$ exceed
what sampling noise alone would explain? A significant result for
`undem` levels would confirm that the penalty for undemocratic behavior
varies meaningfully across respondents.

```{r}
#| label: gs-het
het_gs <- sc_heterogeneity_test(fit_gs, adjust = "bh")
het_gs$estimate
```

## Direction and intensity

The undemocratic levels should show a large negative average direction
(voters penalize them) coupled with high intensity. The structural
question is whether the *dispersion* of those coefficients concentrates
on co-partisans, as Graham and Svolik argue.

```{r}
#| label: gs-di
sc_direction_intensity(fit_gs)
```

## Subgroup analysis by partisanship

We re-average the held-out $\hat\beta(\mathbf Z_i)$ separately for
Democrats, Independents, and Republicans using `sc_subgroup()`. The
`resp_pid` column on the standardized $\mathbf Z$ matrix preserves the
sign of the underlying -1/0/+1 encoding.

```{r}
#| label: gs-subgroup
pid_col <- fit_gs$Z[, "resp_pid"]
sub <- sc_subgroup(fit_gs, list(
  dem = pid_col < -0.5,
  ind = abs(pid_col) <= 0.5,
  rep = pid_col >  0.5
))
do.call(rbind, lapply(names(sub), function(g) {
  e <- sub[[g]]$estimate
  rows <- grepl("^dem_codeu_", e$dummy_name)
  data.frame(group = g, e[rows, c("dummy_name", "theta", "se")],
             row.names = NULL)
}))
```

The ability to read subgroup-specific $\hat\theta$ off the same fit
--- without re-estimating on subsets --- is the operational advantage
of the structural model over stacked subgroup AMCE regressions.

## Preference clustering

A complementary view on latent preference types: run k-means on the
per-respondent $\hat\beta(\mathbf Z_i)$ matrix to find clusters of
respondents with similar preference profiles.

```{r}
#| label: gs-clusters
cl_gs <- sc_clusters(fit_gs, k = 3L, seed = 2024)
cl_gs$estimate$sizes
round(cl_gs$estimate$centers, 2)
```

## Visualization

### AMCE coefficient plot

```{r}
#| label: gs-amce-plot
#| fig-height: 5
plot_amce(fit_gs, groups = gs_groups, labels = gs_labels)
```

### Fraction favor / oppose

```{r}
#| label: gs-fraction-bar
#| fig-height: 5
plot_fraction(fit_gs, groups = gs_groups, labels = gs_labels)
```

### Subgroup AMCE comparison

How does the penalty for undemocratic behavior differ across partisan
groups? We plot subgroup-specific $\hat\theta$ for the `undem` dummies.

```{r}
#| label: gs-subgroup-plot
#| fig-height: 5
pid_col <- fit_gs$Z[, "resp_pid"]
plot_subgroup(
  fit_gs,
  subgroup = list(Democrat    = pid_col < -0.5,
                  Independent = abs(pid_col) <= 0.5,
                  Republican  = pid_col > 0.5),
  dummies = grep("^dem_codeu_", fit_gs$attr_names, value = TRUE),
  groups = gs_groups, labels = gs_labels,
  title = "Subgroup AMCE: undemocratic behavior"
)
```

### Beta ridgelines

```{r}
#| label: gs-ridgelines
#| fig-height: 6
plot(fit_gs, "beta_ridgelines", groups = gs_groups, labels = gs_labels)
```

### Loss trace

```{r}
#| label: gs-loss
plot(fit_gs, "loss_trace")
```
