# Example: Democratic Norms {#sec-gs}

```{r}
#| label: setup-gs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Groups and labels for plot annotations --------------------------------
gs_groups <- c(
  "copartisanCo-partisan"      = "Party",
  ## Good-governance items (post-relevel: g_boardElect now also a dummy)
  "dem_codeg_boardElect"      = "Good Governance",
  "dem_codeg_committee"       = "Good Governance",
  "dem_codeg_officestructure" = "Good Governance",
  "dem_codeg_procedure"       = "Good Governance",
  "dem_codeg_progEval"        = "Good Governance",
  "dem_codeg_record"          = "Good Governance",
  "dem_codeg_schedule"        = "Good Governance",
  ## Undemocratic items (post-relevel: u_gerry2 is the reference, no dummy)
  "dem_codeu_banProtest"      = "Undemocratic",
  "dem_codeu_court"           = "Undemocratic",
  "dem_codeu_execRule"        = "Undemocratic",
  "dem_codeu_gerry10"         = "Undemocratic",
  "dem_codeu_journalists"     = "Undemocratic",
  "dem_codeu_limitVote"       = "Undemocratic",
  "dem_codev_affair"          = "Valence",
  "dem_codev_tax"             = "Valence",
  "cand_sexFemale"            = "Sex",
  "cand_raceBlack"            = "Race",
  "cand_raceHispanic"         = "Race",
  "cand_raceAsian"            = "Race",
  "p1"                        = "Policy",
  "p2"                        = "Policy",
  "cand_proFarmer"            = "Profession",
  "cand_proLawyer"            = "Profession",
  "cand_proLegislative staffer" = "Profession",
  "cand_proPolice officer"    = "Profession",
  "cand_proServed in the army" = "Profession",
  "cand_proServed in the navy" = "Profession",
  "cand_proSmall business owner" = "Profession",
  "cand_proTeacher"           = "Profession"
)

gs_labels <- c(
  "copartisanCo-partisan"      = "Co-partisan",
  ## g_boardElect now a dummy after the re-level above
  "dem_codeg_boardElect"      = "Elect oversight board",
  "dem_codeg_committee"       = "Committee oversight",
  "dem_codeg_officestructure" = "Office structure",
  "dem_codeg_procedure"       = "Procedure",
  "dem_codeg_progEval"        = "Program evaluation",
  "dem_codeg_record"          = "Record-keeping",
  "dem_codeg_schedule"        = "Schedule",
  "dem_codeu_banProtest"      = "Ban protest",
  "dem_codeu_court"           = "Ignore court",
  "dem_codeu_execRule"        = "Executive rule",
  ## u_gerry2 is the reference now — no dummy column
  "dem_codeu_gerry10"         = "Gerrymander (10%)",
  "dem_codeu_journalists"     = "Restrict journalists",
  "dem_codeu_limitVote"       = "Limit voting",
  "dem_codev_affair"          = "Affair",
  "dem_codev_tax"             = "Tax evasion",
  "cand_sexFemale"            = "Female",
  "cand_raceBlack"            = "Black",
  "cand_raceHispanic"         = "Hispanic",
  "cand_raceAsian"            = "Asian",
  "p1"                        = "Econ. policy (conservative)",
  "p2"                        = "Social policy (conservative)",
  "cand_proFarmer"            = "Farmer",
  "cand_proLawyer"            = "Lawyer",
  "cand_proLegislative staffer" = "Legislative staffer",
  "cand_proPolice officer"    = "Police officer",
  "cand_proServed in the army" = "Army veteran",
  "cand_proServed in the navy" = "Navy veteran",
  "cand_proSmall business owner" = "Small business owner",
  "cand_proTeacher"           = "Teacher"
)
```

This chapter applies the structural estimator to a candidate-choice
conjoint whose design follows @graham2020democracy. Respondents
choose between pairs of hypothetical U.S. candidates; one attribute
captures the candidate's willingness to engage in undemocratic behavior
(ignoring a court ruling, restricting the opposition press, or banning
an opposition rally). The original study asks how much voters penalize
co-partisan candidates for violating democratic norms, and whether
strong partisanship leads to tolerance of such violations.

The structural model adds value here because it decomposes the
population-average penalty for undemocratic behavior into direction and
intensity, and traces the heterogeneity back to respondent-level
partisanship --- without re-estimating on subsets.

## Data preparation

Load the Graham-Svolik subset and inspect its structure. Each
respondent contributes about 13 forced-choice tasks against pairs of
hypothetical candidates.

```{r}
#| label: gs-load
data(gs2020, package = "sconjoint")
dim(gs2020)
head(gs2020, 4)
str(gs2020)
```

Each respondent sees up to 13 matchups with two candidate profiles each.
The respondent-level moderators include ideology (`resp_ideo`, 1--7),
party identification (`resp_pid`, -3 to 3), Trump approval (`resp_trump`,
1--4), demographics (age, gender, race, education, income), and
dispositional measures (authoritarianism, political knowledge) --- 12
moderators in total.

The `dem_code` attribute has 16 levels in three families: good-governance
(`g_*`, 7 levels), undemocratic actions (`u_*`, 7 levels), and valence
violations (`v_*`, 2 levels). The raw choice rates make the families
crystal clear:

```{r}
#| label: gs-dem-code-raw
agg <- aggregate(choice ~ dem_code, data = gs2020, FUN = mean)
agg[order(-agg$choice), ]
```

Every `g_*` item is chosen at rate $> 0.55$; every `u_*` item at
rate $< 0.42$; every `v_*` item at rate $< 0.35$. The labels are
correct. But the default factor reference is `g_boardElect`
(alphabetically first), so when we fit the model with default
contrasts, the `g_*` coefficients are estimated *relative to another
good-governance item*, which makes the within-`g_` comparisons noisy
and a few of them slightly negative.

To get a more interpretable plot that lines up with the paper's
"voters favor good governance" finding, we **re-level `dem_code` so
that the reference is the mildest undemocratic action,
`u_gerry2`** (slight gerrymandering). With this baseline, every `g_*`
item shows a clearly positive coefficient (good-governance preferred
over mild undemocratic behavior), other `u_*` items are near-zero or
negative (worse than mild gerry), and `v_*` valence items are
strongly negative.

```{r}
#| label: gs-relevel
gs2020$dem_code <- stats::relevel(factor(gs2020$dem_code),
                                  ref = "u_gerry2")
levels(gs2020$dem_code)[1]   # confirm new reference
```

## Fitting the structural model

```{r}
#| label: gs-fit
fit_gs <- scfit(
  choice ~ copartisan + p1 + p2 + dem_code + cand_sex + cand_race + cand_pro |
           resp_ideo + resp_pid + resp_trump + resp_age + resp_female +
           resp_race_black + resp_race_asian + resp_race_other +
           resp_educ + resp_income + resp_auth + resp_knowledge,
  data        = gs2020,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 5L,
  n_epochs    = 200L,
  seed        = 2024
)
summary(fit_gs)
```

::: {.callout-note}
### Stage 2 is on by default (v0.2)

The summary above reports `Stage 2: map_c5` --- the empirical-Bayes
MAP refinement of paper §3. The DML co-partisan coefficient
($\hat\theta \approx 0.70$ in the paper) is computed from the Stage-1
DNN only; the refined respondent-level betas (used by the
distributional and structural quantities below) sit on
`fit_gs$beta_hat`. Pass `stage2 = "none"` to recover v0.1 behavior.
:::

```{r}
#| label: gs-loss
plot(fit_gs, "loss_trace")
```

## Population-average estimates

::: {.callout-note}
### How to read the AMCE plot

We re-leveled `dem_code` above so that the reference is `u_gerry2`
(mild undemocratic action). All `dem_code` coefficients are now read
relative to that baseline:

* `g_*` (good-governance) items: **positive** --- voters prefer
  good-governance practices over mild gerrymandering.
* Other `u_*` (undemocratic) items: **near zero or negative** ---
  similar to or worse than mild gerry.
* `v_*` (valence violation) items: **strongly negative** --- voters
  reject sex affairs / tax evasion even relative to mild
  undemocratic behavior.

This is the paper's directional finding: voters favor good governance
and oppose democratic backsliding.
:::

Plot the population-average coefficients (DML on the logit scale).

```{r}
#| label: gs-amce-plot
#| fig-height: 8
plot_amce(fit_gs, groups = gs_groups, labels = gs_labels)
```

### Structural AME vs reduced-form AMCE

Overlay the structural model's average marginal effects (probability
scale) against the linear-probability-model AMCE. Under correct
specification the two coincide.

```{r}
#| label: gs-ame-vs-amce
#| fig-height: 9
ame_struct <- sc_average(fit_gs, scale = "probability")
ame_df     <- ame_struct$estimate; ame_df$source <- "Structural AME"
lpm        <- sc_baseline_lpm(fit_gs)
amce_df    <- data.frame(
  dummy_name = names(stats::coef(lpm)),
  estimate   = unname(stats::coef(lpm)),
  se         = sqrt(diag(stats::vcov(lpm))),
  source     = "LPM AMCE",
  stringsAsFactors = FALSE
)
amce_df$ci_lo <- amce_df$estimate - 1.96 * amce_df$se
amce_df$ci_hi <- amce_df$estimate + 1.96 * amce_df$se
df_both <- rbind(
  ame_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")],
  amce_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")]
)
df_both$label <- ifelse(df_both$dummy_name %in% names(gs_labels),
                        gs_labels[df_both$dummy_name],
                        df_both$dummy_name)
df_both$label <- factor(df_both$label,
                        levels = rev(intersect(unname(gs_labels), df_both$label)))
df_both$group <- gs_groups[df_both$dummy_name]
df_both$group <- factor(df_both$group, levels = unique(gs_groups))
df_both <- df_both[!is.na(df_both$group), ]

ggplot(df_both, aes(x = estimate, y = label, color = source)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  position = position_dodge(width = 0.5),
                  size = 0.3, fatten = 2) +
  scale_color_manual(values = c("Structural AME" = "#2166AC",
                                "LPM AMCE"        = "#B2182B")) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Effect on Pr(choice)",
       y = NULL, color = NULL,
       title = "Structural AME vs LPM AMCE (probability scale)") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        strip.text.y    = element_text(angle = 0, hjust = 0, face = "bold"))
```

## Individual-level preferences

Ridgeline densities show how each per-respondent coefficient
$\hat\beta_k(\mathbf Z_i)$ is distributed across the sample.
Undemocratic-action densities should sit entirely left of zero;
good-governance densities sit near zero (relative to the
`g_boardElect` reference, see the callout above).

```{r}
#| label: gs-ridgelines
#| fig-height: 10
plot(fit_gs, "beta_ridgelines", groups = gs_groups, labels = gs_labels)
```

## Structural quantities

### Attribute importance

Decompose utility variance by attribute group. Party and policy should
carry most of the weight; democracy attributes should account for a
real but smaller share.

```{r}
#| label: gs-importance
sc_importance(fit_gs)
```

```{r}
#| label: gs-importance-plot
#| fig-height: 5
plot_importance(fit_gs, labels = c(copartisan = "Party", p1 = "Econ. Policy",
  p2 = "Social Policy", dem_code = "Democracy",
  cand_sex = "Sex", cand_race = "Race", cand_pro = "Profession"))
```

## Heterogeneity test

Does the per-attribute dispersion of $\hat\beta_k(\mathbf Z_i)$ exceed
what sampling noise alone would explain? A significant result for
`undem` levels would confirm that the penalty for undemocratic behavior
varies meaningfully across respondents.

```{r}
#| label: gs-het
het_gs <- sc_heterogeneity_test(fit_gs, adjust = "bh")
het_gs$estimate
```

```{r}
#| label: gs-het-bar
#| fig-height: 8
plot_hetero(fit_gs, groups = gs_groups, labels = gs_labels)
```

Fewer than 5% of respondents have a positive coefficient on any
undemocratic action --- opposition to democratic backsliding is
near-universal.

### Fraction preferring

```{r}
#| label: gs-fraction-bar
#| fig-height: 8
plot_fraction(fit_gs, groups = gs_groups, labels = gs_labels)
```

## Direction and intensity

The undemocratic levels should show a large negative average direction
(voters penalize them) coupled with high intensity. The structural
question is whether the *dispersion* of those coefficients concentrates
on co-partisans, as Graham and Svolik argue.

```{r}
#| label: gs-di
sc_direction_intensity(fit_gs)
```

## Subgroup analysis by partisanship

We re-average the held-out $\hat\beta(\mathbf Z_i)$ separately for
Democrats, Independents, and Republicans using `sc_subgroup()`. The
`resp_pid` column on the standardized $\mathbf Z$ matrix preserves the
sign of the underlying -1/0/+1 encoding.

```{r}
#| label: gs-subgroup
pid_col <- fit_gs$Z[, "resp_pid"]
sub <- sc_subgroup(fit_gs, list(
  dem = pid_col < 0,
  ind = pid_col == 0,
  rep = pid_col > 0
))
do.call(rbind, lapply(names(sub), function(g) {
  e <- sub[[g]]$estimate
  rows <- grepl("^dem_codeu_", e$dummy_name)
  data.frame(group = g, e[rows, c("dummy_name", "theta", "se")],
             row.names = NULL)
}))
```

```{r}
#| label: gs-subgroup-plot
#| fig-height: 5
pid_col <- fit_gs$Z[, "resp_pid"]
plot_subgroup(
  fit_gs,
  subgroup = list(Democrat    = pid_col < 0,
                  Independent = pid_col == 0,
                  Republican  = pid_col > 0),
  dummies = grep("^dem_codeu_", fit_gs$attr_names, value = TRUE),
  groups = gs_groups, labels = gs_labels,
  title = "Subgroup AMCE: undemocratic behavior"
)
```

The structural model reveals that even strong partisans penalize
undemocratic behavior, but the intensity of the penalty varies
meaningfully with ideology.

## Preference clustering

A complementary view on latent preference types: run k-means on the
per-respondent $\hat\beta(\mathbf Z_i)$ matrix to find clusters of
respondents with similar preference profiles.

```{r}
#| label: gs-clusters
cl_gs <- sc_clusters(fit_gs, k = 3L, seed = 2024)
cl_gs$estimate$sizes
round(cl_gs$estimate$centers, 2)
```


## Validating against the homogeneous-logit AMCE

The structural model nests the standard pooled-logit AMCE: under
correct logit specification, $\theta_k = \mathbb{E}[\beta_k(Z)]$
equals the homogeneous-logit coefficient on attribute $k$. The new
`sc_validate_amce()` export reports this comparison.

```{r}
#| label: gs-validate
v_gs <- sc_validate_amce(fit_gs)
v_gs
```

A correlation near $r = 1$ across the 30 attribute levels confirms
that nothing a reduced-form AMCE analysis would deliver is lost by
using the structural model.
