# Example: Democratic Norms {#sec-gs}

```{r}
#| label: setup-gs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)
```

This chapter applies the structural estimator to a candidate-choice
conjoint whose design follows @graham2020democracy. Respondents
choose between pairs of hypothetical U.S. candidates; one attribute
captures the candidate's willingness to engage in undemocratic behavior
(ignoring a court ruling, restricting the opposition press, or banning
an opposition rally). The structural model decomposes the
population-average penalty for undemocratic behavior into direction and
intensity, and traces the heterogeneity back to respondent-level
partisanship --- without re-estimating on subsets.

::: {.callout-note}
The `gs2020` dataset is a **seed-fixed synthetic fixture** for
illustration.
:::

## Data preparation

```{r}
#| label: gs-load
data(gs2020, package = "sconjoint")
dim(gs2020)
head(gs2020, 4)
```

## Fitting

```{r}
#| label: gs-fit
fit_gs <- scfit(
  choice ~ federal + immigration + tax + abortion + undem + cand_party |
           resp_pid + resp_ideo + age,
  data        = gs2020,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_gs)
```

```{r}
#| label: gs-loss
plot(fit_gs, "loss_trace")
```

## Population-average estimates

```{r}
#| label: gs-amce-plot
#| fig-height: 5
plot_amce(fit_gs)
```

## Individual-level preferences

```{r}
#| label: gs-ridgelines
#| fig-height: 6
plot(fit_gs, "beta_ridgelines")
```

## Structural quantities

### Direction and intensity

```{r}
#| label: gs-di
sc_direction_intensity(fit_gs)
```

### Fraction preferring

```{r}
#| label: gs-fraction-bar
#| fig-height: 5
plot_fraction(fit_gs)
```

### Heterogeneity test

```{r}
#| label: gs-het
het_gs <- sc_heterogeneity_test(fit_gs, adjust = "bh")
het_gs$estimate
```

```{r}
#| label: gs-het-bar
#| fig-height: 5
plot_hetero(fit_gs)
```

### Subgroup analysis by partisanship

```{r}
#| label: gs-subgroup
pid_col <- fit_gs$Z[, "resp_pid"]
sub <- sc_subgroup(fit_gs, list(
  dem = pid_col < -0.5,
  ind = abs(pid_col) <= 0.5,
  rep = pid_col >  0.5
))
do.call(rbind, lapply(names(sub), function(g) {
  e <- sub[[g]]$estimate
  rows <- grepl("^undem", e$dummy_name)
  data.frame(group = g, e[rows, c("dummy_name", "theta", "se")],
             row.names = NULL)
}))
```

```{r}
#| label: gs-subgroup-plot
#| fig-height: 5
plot_subgroup(
  fit_gs,
  subgroup = list(Democrat    = pid_col < -0.5,
                  Independent = abs(pid_col) <= 0.5,
                  Republican  = pid_col > 0.5),
  dummies = grep("^undem", fit_gs$attr_names, value = TRUE),
  title = "Subgroup AMCE: undemocratic behavior"
)
```

### Preference clustering

```{r}
#| label: gs-clusters
cl_gs <- sc_clusters(fit_gs, k = 3L, seed = 2024)
cl_gs$estimate$sizes
round(cl_gs$estimate$centers, 2)
```

## Advanced quantities

### Decisiveness

How decisively do voters reject a candidate who ignores a court ruling?

```{r}
#| label: gs-decisiveness
sc_decisiveness(fit_gs,
  A = list(undem = "none"),
  B = list(undem = "ignore_court")
)
```

### Preference inequality

```{r}
#| label: gs-inequality
sc_inequality(fit_gs, measure = "gini")
```

Higher Gini values indicate that a small fraction of respondents hold
much stronger preferences than the rest.
