---
title: "Climate Treaties"
---

```{r}
#| label: setup-bs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)
```

This chapter applies the structural estimator to a climate-treaty
conjoint experiment whose design follows Bechtel and Scheve (2013).
Respondents evaluate pairs of hypothetical international climate
agreements varying on six attributes --- including a numeric cost
attribute (`cost_usd`) --- and select the one they prefer. Because cost
is numeric, this dataset is a natural showcase for willingness-to-pay
(WTP) analysis via `sc_wtp()` and compensating differentials via
`sc_compensating()`.

::: {.callout-note}
The `bs2013` dataset is a **seed-fixed synthetic fixture** for
illustration. Do not use it to draw substantive inferences about the
Bechtel--Scheve population.
:::

## Data preparation

```{r}
#| label: bs-load
data(bs2013, package = "sconjoint")
dim(bs2013)
head(bs2013, 4)
```

```{r}
#| label: bs-structure
str(bs2013)
```

Each respondent contributes six forced-choice tasks with two treaty
profiles each (12 rows per respondent). The respondent-level moderators
are `resp_female` (binary), `age` (standardized), and `resp_ideo`
(standardized ideology).

## Fitting the structural model

```{r}
#| label: bs-fit
fit_bs <- scfit(
  choice ~ distribution + enforcement + monitoring + participation +
           sanctions + cost_usd |
           resp_female + age + resp_ideo,
  data        = bs2013,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_bs)
```

## Structural quantities

### Attribute importance

```{r}
#| label: bs-importance
sc_importance(fit_bs)
```

### Direction and intensity

```{r}
#| label: bs-di
sc_direction_intensity(fit_bs)
```

### Willingness to pay

Because `cost_usd` is a single numeric attribute (one dummy column),
`sc_wtp()` can compute the WTP for each non-cost attribute level. The
WTP is the negative MRS against the cost column, expressed in the same
units as `cost_usd`.

```{r}
#| label: bs-wtp
## WTP for selected treaty features
sc_wtp(fit_bs, attr = "distribution:ability_to_pay", cost_attr = "cost_usd")
sc_wtp(fit_bs, attr = "enforcement:trade_sanctions", cost_attr = "cost_usd")
sc_wtp(fit_bs, attr = "monitoring:international",    cost_attr = "cost_usd")
```

### Compensating differentials

How much cost would respondents accept to gain a preferred treaty
feature? `sc_compensating()` computes the per-respondent ratio and
reports the trimmed mean.

```{r}
#| label: bs-compensating
sc_compensating(fit_bs,
                benefit = "distribution:ability_to_pay",
                cost    = "cost_usd")
sc_compensating(fit_bs,
                benefit = "enforcement:trade_sanctions",
                cost    = "cost_usd")
```

### Fraction preferring

```{r}
#| label: bs-fraction
frac_bs <- sc_fraction_preferring(fit_bs, threshold = 0)
frac_bs$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

## Visualization

### AMCE coefficient plot

```{r}
#| label: bs-amce-plot
#| fig-height: 5
theta_bs <- coef(fit_bs)
se_bs <- sqrt(diag(vcov(fit_bs)))
df_amce_bs <- data.frame(
  dummy = names(theta_bs), estimate = theta_bs,
  ci_lo = theta_bs - 1.96 * se_bs, ci_hi = theta_bs + 1.96 * se_bs,
  row.names = NULL
)

ggplot(df_amce_bs, aes(x = estimate, y = dummy)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  color = "#E41A1C", size = 0.4) +
  labs(x = expression(hat(theta)[k]), y = NULL,
       title = "Population-average AMCE") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())
```

### Fraction favor / oppose

```{r}
#| label: bs-fraction-bar
#| fig-height: 5
df_frac_bs <- frac_bs$estimate
df_long_bs <- rbind(
  data.frame(dummy = df_frac_bs$dummy_name, value = df_frac_bs$frac_positive,
             direction = "Favor", stringsAsFactors = FALSE),
  data.frame(dummy = df_frac_bs$dummy_name, value = -df_frac_bs$frac_negative,
             direction = "Oppose", stringsAsFactors = FALSE)
)

ggplot(df_long_bs, aes(x = value, y = dummy, fill = direction)) +
  geom_col() +
  geom_vline(xintercept = 0, color = "gray30") +
  scale_fill_manual(values = c(Favor = "#377EB8", Oppose = "#E41A1C")) +
  labs(x = "Fraction of respondents", y = NULL, fill = NULL,
       title = "Fraction favoring / opposing each level") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        legend.position = "bottom")
```

### Beta ridgelines

```{r}
#| label: bs-ridgelines
#| fig-height: 6
plot(fit_bs, "beta_ridgelines")
```

### Loss trace

```{r}
#| label: bs-loss
plot(fit_bs, "loss_trace")
```

## Subgroup analysis by ideology

Do respondents on the ideological left value different treaty features
than those on the right? We split at the median of `resp_ideo`.

```{r}
#| label: bs-subgroup
ideo_col <- fit_bs$Z[, "resp_ideo"]
sub_left  <- sc_subgroup(fit_bs, ideo_col < 0)
sub_right <- sc_subgroup(fit_bs, ideo_col >= 0)

rbind(
  data.frame(group = "left",  sub_left$estimate[,  c("dummy_name", "theta", "se")],
             row.names = NULL),
  data.frame(group = "right", sub_right$estimate[, c("dummy_name", "theta", "se")],
             row.names = NULL)
)
```

The structural model lets us read off subgroup-specific AMCEs from the
same fit, without re-estimating on subsets. Ideological differences in
the valuation of enforcement and cost are visible in the coefficient
table above and can be further explored via per-subgroup ridgeline
plots.
