---
title: "Fitting scfit()"
---

# Fitting `scfit()` {#sec-fitting}

`scfit()` is the single entry point to the structural estimator. It
consumes a formula of the form

$$ \text{choice} \sim \text{attr}_1 + \text{attr}_2 + \cdots \mid z_1 + z_2 + \cdots $$

together with a long-format data frame and the column names of the
respondent, task, and within-task profile identifiers. The left-hand
side of `|` names the conjoint attributes; the right-hand side names
the respondent-level moderators $\mathbf Z$ that enter the DNN's
mapping to $\boldsymbol\beta$.

Internally `scfit()` (i) dummy-encodes the attribute columns against
their reference levels, (ii) builds the per-task $\Delta X$ matrix,
(iii) assigns respondents to `K` cross-fit folds, (iv) trains a
full-batch Adam DNN on each training fold and predicts
$\hat\beta(Z_i)$ on the held-out fold, (v) estimates the $\Lambda(Z)$
Hessian of the logit index, and (vi) assembles the DML influence
function, point estimates, and respondent-clustered variance.

## Bit-exact parallel determinism

When `seed` is supplied, `scfit()` guarantees that the entire
pipeline — including the cross-fitting step run in parallel across
workers — produces *bit-identical* output regardless of `n_cores`.
This property is enforced by L'Ecuyer-CMRG parallel streams and a
save/restore of the top-level R and torch RNG states. The demo below
(not evaluated on CI) confirms that `identical()` returns `TRUE`
across a single-core and a two-core run.

```{r}
library(sconjoint)
```

Tuning `n_epochs`, `K`, and the `hidden` layer widths trades
wall-clock time against the flexibility of the learned preference
function. The `"auto"` default for `hidden` adapts the architecture
to the number of observations and is a sensible starting point.

```{r}
cat("fitting guide\n")
```

See @sec-interpreting for how to read the resulting coefficient
table.
