# Example: Climate Treaties {#sec-bs}

```{r}
#| label: setup-bs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)
```

This chapter applies the structural estimator to a climate-treaty
conjoint experiment whose design follows @bechtel_scheve_2013.
Respondents evaluate pairs of hypothetical international climate
agreements varying on six attributes --- including a numeric cost
attribute (`cost_usd`) --- and select the one they prefer. Because cost
is numeric, this dataset is a natural showcase for willingness-to-pay
(WTP) analysis via `sc_wtp()` and compensating differentials via
`sc_compensating()`.

::: {.callout-note}
The `bs2013` dataset is a **seed-fixed synthetic fixture** for
illustration. Do not use it to draw substantive inferences about the
Bechtel--Scheve population.
:::

## Data preparation

```{r}
#| label: bs-load
data(bs2013, package = "sconjoint")
dim(bs2013)
head(bs2013, 4)
```

```{r}
#| label: bs-structure
str(bs2013)
```

Each respondent contributes six forced-choice tasks with two treaty
profiles each (12 rows per respondent). The respondent-level moderators
are `resp_female` (binary), `age` (standardized), and `resp_ideo`
(standardized ideology).

## Fitting the structural model

```{r}
#| label: bs-fit
fit_bs <- scfit(
  choice ~ distribution + enforcement + monitoring + participation +
           sanctions + cost_usd |
           resp_female + age + resp_ideo,
  data        = bs2013,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_bs)
```

## Structural quantities

### Attribute importance

```{r}
#| label: bs-importance
sc_importance(fit_bs)
```

### Direction and intensity

```{r}
#| label: bs-di
sc_direction_intensity(fit_bs)
```

### Willingness to pay

Because `cost_usd` is a single numeric attribute (one dummy column),
`sc_wtp()` can compute the WTP for each non-cost attribute level. The
WTP is the negative MRS against the cost column, expressed in the same
units as `cost_usd`.

```{r}
#| label: bs-wtp
## WTP for selected treaty features
sc_wtp(fit_bs, attr = "distribution:ability_to_pay", cost_attr = "cost_usd")
sc_wtp(fit_bs, attr = "enforcement:trade_sanctions", cost_attr = "cost_usd")
sc_wtp(fit_bs, attr = "monitoring:international",    cost_attr = "cost_usd")
```

### Compensating differentials

How much cost would respondents accept to gain a preferred treaty
feature? `sc_compensating()` computes the per-respondent ratio and
reports the trimmed mean.

```{r}
#| label: bs-compensating
sc_compensating(fit_bs,
                benefit = "distribution:ability_to_pay",
                cost    = "cost_usd")
sc_compensating(fit_bs,
                benefit = "enforcement:trade_sanctions",
                cost    = "cost_usd")
```

### Fraction preferring

```{r}
#| label: bs-fraction
frac_bs <- sc_fraction_preferring(fit_bs, threshold = 0)
frac_bs$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

## Advanced quantities

Because `bs2013` includes a numeric cost attribute (`cost_usd`), it is
a natural showcase for surplus, welfare, and demand analysis.

### Consumer surplus

Compare the surplus from a cheap treaty versus an expensive one.

```{r}
#| label: bs-surplus
sc_surplus(fit_bs, profiles = list(
  list(cost_usd = 10),
  list(cost_usd = 50)
))
```

The log-sum surplus over the two-profile choice set reflects the
aggregate welfare respondents derive. The cheap treaty ($10) dominates
for most respondents (since $\hat\beta_{\mathrm{cost}} < 0$), but the
surplus captures the option value of having both alternatives available.

### Welfare change

What is the welfare gain from replacing an expensive treaty with a
cheaper one that also includes international monitoring?

```{r}
#| label: bs-welfare
sc_welfare_change(
  fit_bs,
  old_set = list(list(cost_usd = 50)),
  new_set = list(list(cost_usd = 10, monitoring = "international"))
)
```

The positive welfare change confirms that respondents prefer cheaper
treaties with international monitoring. The magnitude combines the
cost savings ($\hat\beta_{\mathrm{cost}} \times (-40)$) and the
monitoring premium ($\hat\beta_{\mathrm{monitoring:international}}$).

### Demand curve

The aggregate demand curve traces the average choice probability as a
function of the cost attribute, holding other attributes at reference
levels.

```{r}
#| label: bs-demand
demand <- sc_demand_curve(fit_bs, cost_attr = "cost_usd")
demand
```

```{r}
#| label: bs-demand-plot
#| fig-height: 4
demand_df <- demand$estimate
ggplot(demand_df, aes(x = cost, y = demand)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.2,
              fill = "#2166AC") +
  geom_line(color = "#2166AC", linewidth = 1) +
  labs(x = "Cost (USD per household per month)",
       y = "Average choice probability",
       title = "Aggregate demand curve for climate treaties") +
  theme_minimal(base_size = 12)
```

The downward-sloping demand curve is consistent with $\hat\beta_{\mathrm{cost}} < 0$:
as the cost of a treaty rises, fewer respondents choose it over the
reference alternative. The confidence band widens at the extremes of
the cost grid where fewer observations anchor the estimate.

## Visualization

### AMCE coefficient plot

```{r}
#| label: bs-amce-plot
#| fig-height: 5
plot_amce(fit_bs)
```

### Fraction favor / oppose

```{r}
#| label: bs-fraction-bar
#| fig-height: 5
plot_fraction(fit_bs)
```

### Heterogeneity

```{r}
#| label: bs-het-bar
#| fig-height: 5
plot_hetero(fit_bs)
```

### Subgroup AMCE by ideology

Do respondents on the ideological left, center, and right value
treaty features differently?

```{r}
#| label: bs-ideo-plot
#| fig-height: 6
ideo_col <- fit_bs$Z[, "resp_ideo"]
ideo_cuts <- quantile(ideo_col, probs = c(1/3, 2/3))
plot_subgroup(
  fit_bs,
  subgroup = list(Left   = ideo_col <= ideo_cuts[1],
                  Center = ideo_col > ideo_cuts[1] & ideo_col <= ideo_cuts[2],
                  Right  = ideo_col > ideo_cuts[2]),
  title = "Subgroup AMCE by ideology tercile"
)
```

### Beta ridgelines

```{r}
#| label: bs-ridgelines
#| fig-height: 6
plot(fit_bs, "beta_ridgelines")
```

### Loss trace

```{r}
#| label: bs-loss
plot(fit_bs, "loss_trace")
```
