# Example: Climate Treaties {#sec-bs}

```{r}
#| label: setup-bs
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)
```

This chapter applies the structural estimator to a climate-treaty
conjoint experiment whose design follows @bechtel_scheve_2013.
Respondents evaluate pairs of hypothetical international climate
agreements varying on six attributes --- including a numeric cost
attribute (`cost_usd`). Because cost is numeric, this dataset is a
natural showcase for willingness-to-pay and demand analysis.

## Data preparation

```{r}
#| label: bs-load
data(bs2013, package = "sconjoint")
dim(bs2013)
head(bs2013, 4)
```

## Fitting

```{r}
#| label: bs-fit
fit_bs <- scfit(
  choice ~ cost_usd + distribution + participation + emissions +
           sanctions + monitoring |
           resp_female + resp_age + resp_ideo,
  data        = bs2013,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 2024
)
summary(fit_bs)
```

```{r}
#| label: bs-loss
plot(fit_bs, "loss_trace")
```

## Population-average estimates

```{r}
#| label: bs-amce-plot
#| fig-height: 5
plot_amce(fit_bs)
```

## Individual-level preferences

```{r}
#| label: bs-ridgelines
#| fig-height: 6
plot(fit_bs, "beta_ridgelines")
```

## Structural quantities

### Attribute importance

```{r}
#| label: bs-importance
sc_importance(fit_bs)
```

### Direction and intensity

```{r}
#| label: bs-di
sc_direction_intensity(fit_bs)
```

### Fraction preferring

```{r}
#| label: bs-fraction
frac_bs <- sc_fraction_preferring(fit_bs, threshold = 0)
frac_bs$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

```{r}
#| label: bs-fraction-bar
#| fig-height: 5
plot_fraction(fit_bs)
```

### Willingness to pay

Because `cost_usd` is numeric, `sc_wtp()` computes the WTP for each
non-cost attribute level --- the negative MRS against cost, expressed in
the same units as `cost_usd`.

```{r}
#| label: bs-wtp
sc_wtp(fit_bs, attr = "distribution:Rich pay more, shared",    cost_attr = "cost_usd")
sc_wtp(fit_bs, attr = "emissions:60% reduction",      cost_attr = "cost_usd")
sc_wtp(fit_bs, attr = "monitoring:Independent commission",     cost_attr = "cost_usd")
```

### Compensating differentials

```{r}
#| label: bs-compensating
sc_compensating(fit_bs,
                benefit = "distribution:Rich pay more, shared",
                cost    = "cost_usd")
```

### Heterogeneity test

```{r}
#| label: bs-het
sc_heterogeneity_test(fit_bs, adjust = "bh")
```

```{r}
#| label: bs-het-bar
#| fig-height: 5
plot_hetero(fit_bs)
```

### Subgroup analysis by ideology

```{r}
#| label: bs-ideo-plot
#| fig-height: 6
ideo_col <- fit_bs$Z[, "resp_ideo"]
ideo_cuts <- quantile(ideo_col, probs = c(1/3, 2/3))
plot_subgroup(
  fit_bs,
  subgroup = list(Left   = ideo_col <= ideo_cuts[1],
                  Center = ideo_col > ideo_cuts[1] & ideo_col <= ideo_cuts[2],
                  Right  = ideo_col > ideo_cuts[2]),
  title = "Subgroup AMCE by ideology tercile"
)
```

## Advanced quantities

### Consumer surplus

```{r}
#| label: bs-surplus
sc_surplus(fit_bs, profiles = list(
  list(cost_usd = 10),
  list(cost_usd = 50)
))
```

### Welfare change

What is the welfare gain from replacing an expensive treaty with a
cheaper one that includes international monitoring?

```{r}
#| label: bs-welfare
sc_welfare_change(
  fit_bs,
  old_set = list(list(cost_usd = 50)),
  new_set = list(list(cost_usd = 10, monitoring = "United Nations"))
)
```

### Demand curve

```{r}
#| label: bs-demand
demand <- sc_demand_curve(fit_bs, cost_attr = "cost_usd")
```

```{r}
#| label: bs-demand-plot
#| fig-height: 4
demand_df <- demand$estimate
ggplot(demand_df, aes(x = cost, y = demand)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.2,
              fill = "#2166AC") +
  geom_line(color = "#2166AC", linewidth = 1) +
  labs(x = "Cost (USD/month)",
       y = "Average choice probability",
       title = "Aggregate demand curve") +
  theme_minimal(base_size = 12)
```

The downward-sloping demand curve is consistent with
$\hat\beta_{\mathrm{cost}} < 0$: as treaty cost rises, fewer
respondents choose it over the reference alternative.
