# Example: Tax Preferences {#sec-br}

```{r}
#| label: setup-br
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Labels and groups for plot annotations -----------------------------------
br_labels <- c(
  "rate_L10"      = "Under $10k",
  "rate_10_35"    = "$10-35k",
  "rate_35_85"    = "$35-85k",
  "rate_85_175"   = "$85-175k",
  "rate_175_375"  = "$175-375k",
  "rate_375P"     = "Over $375k",
  "revenue_score" = "Revenue impact"
)

br_groups <- c(
  "rate_L10"      = "Tax Rates",
  "rate_10_35"    = "Tax Rates",
  "rate_35_85"    = "Tax Rates",
  "rate_85_175"   = "Tax Rates",
  "rate_175_375"  = "Tax Rates",
  "rate_375P"     = "Tax Rates",
  "revenue_score" = "Revenue"
)
```

This chapter applies the structural estimator to a tax-plan conjoint
experiment from @ballard2017structure. Respondents evaluate pairs of
hypothetical income tax plans varying on **six marginal tax rates** (one
per income bracket) plus a **revenue impact** indicator, and select the
plan they prefer. The bundled `br2017` is the paper's full analysis
frame (2,000 respondents, 32,000 rows), so the quantities below reproduce
the empirical results reported in the manuscript.

What makes this dataset unique among the bundled examples is that
**every attribute is numeric** (continuous). There are no factor
attributes --- the six bracket rates are percentages and the revenue
score is a centered integer scale. This showcases how `sconjoint`
handles all-numeric attribute designs.

## Data preparation

```{r}
#| label: br-load
data(br2017, package = "sconjoint")
dim(br2017)
head(br2017, 4)
```

```{r}
#| label: br-structure
str(br2017)
```

Each respondent contributes eight forced-choice tasks with two tax-plan
profiles each (16 rows per respondent). The fit conditions on **23
respondent-level moderators** ($\mathbf Z_i$): demographics (`age_std`,
`female`, `educ_std`, `race_white`, `income_std`, `children`),
partisanship and ideology (`pid7_std`, `conserv_ideo`), economic
attitudes and beliefs (`ineq_averse`, `work_vs_luck`, `taxes_harm_econ`,
`hardwork`, `high_econ_know`, `govt_serv`, `gov_assist`), economic
position (`employed_ft`, `gen_mobile`, `future_mobile`, `hardship`,
`risk_averse`), and engagement/competence (`newsint`, `numeracy`,
`trust`). The raw seven-point party identification is also retained as a
convenience column, `resp_pid7` (1 = strong Democrat to 7 = strong
Republican), which we use below to split respondents by party. It is not
part of the moderator set (`pid7_std` is its standardized counterpart).

## Fitting the structural model

Every attribute here is continuous, with the tax rates encoded as
percentage points (0--55). The paper handles this with the **`varref`**
Stage-2 prior and a low variance floor (**`varref_floor = 1e-3`**) on the
raw rates, rather than the default score-based `map_c5` prior. We follow
the paper's recipe exactly here. Coefficients are in "logit utility per
percentage point" units.

```{r}
#| label: br-fit
#| cache: true
fit_br <- scfit(
  choice ~ rate_L10 + rate_10_35 + rate_35_85 + rate_85_175 +
           rate_175_375 + rate_375P + revenue_score |
           age_std + female + pid7_std + educ_std + race_white +
           income_std + ineq_averse + work_vs_luck + taxes_harm_econ +
           hardwork + high_econ_know + employed_ft + conserv_ideo +
           govt_serv + newsint + numeracy + gen_mobile + future_mobile +
           gov_assist + risk_averse + hardship + children + trust,
  data         = br2017,
  respondent   = "respondent",
  task         = "task",
  profile      = "profile",
  K            = 10L,
  seed         = 42L,
  stage2       = "varref",
  varref_floor = 1e-3,
  parallel     = TRUE,
  n_cores      = 10L
)
summary(fit_br)
```

This is the paper's exact recipe, run on the corrected bracket-5 data
(the distributed file mis-stored the 45% level of the $175-375k bracket;
see `?br2017`) with `varref_floor = 1e-3`. The estimates below therefore
reproduce the manuscript's reported numbers.

::: {.callout-note}
### Why `stage2 = "varref"` for continuous attributes?

The default `map_c5` Stage-2 prior is calibrated assuming
$\mathrm{Var}(\Delta X_k) \approx 1$ (true for factor dummies under
typical randomization). BR's tax rates span 0--50 percentage points, so
the score-based prior becomes loose and the MAP solver leaves
per-respondent estimates with extreme tails. The paper instead uses the
`varref` prior --- $0.5\cdot\mathrm{Var}_i(\hat\beta_{\text{ens},i,k})$,
floored at `varref_floor` --- on the raw rates. The **low floor matters**:
the prior `0.01` default over-shrinks continuous coefficients and badly
degrades recovery of individual progressivity, so continuous designs use
`varref_floor = 1e-3` (the paper's setting). An alternative is
`normalize_deltaX = TRUE` with the default `map_c5` prior, which
standardizes $\Delta X$ internally; see the *Advanced options* chapter.
For factor-dummy designs (SW, GS) neither is needed.
:::

```{r}
#| label: br-loss
plot(fit_br, "loss_trace")
```

## Population-average estimates

DML coefficients per attribute, on the logit scale, with respondent-
clustered SEs. With numeric attributes the coefficient is "logit
utility per unit of the attribute" (per percentage-point of the
rate, or per unit of the centered revenue indicator).

The population pattern is the paper's main result: a **steep
progressive structure** in the bracket rates. The average voter dislikes
higher rates on low brackets ($\hat\theta_{<\$10k} \approx -0.06$) and
mildly favors higher rates on the top bracket
($\hat\theta_{>\$375k} \approx +0.02$), with the coefficients rising
monotonically across brackets and crossing zero around the $85-175k
bracket. The **revenue-impact** coefficient is **weak-positive**
($\hat\theta \approx 0.14$, $p < 0.01$, 95% CI $[0.05, 0.23]$): voters
prefer plans that raise more revenue, but it is the least important of
the seven attributes.

```{r}
#| label: br-amce-plot
#| fig-height: 5
plot_amce(fit_br, groups = br_groups, labels = br_labels)
```

### Structural AME vs reduced-form AMCE

For continuous attributes the AME is "average effect on Pr(choice) per
unit of the attribute". Overlay the structural and LPM versions on the
probability scale.

```{r}
#| label: br-ame-vs-amce
#| fig-height: 5
ame_struct <- sc_average(fit_br, scale = "probability")
ame_df     <- ame_struct$estimate; ame_df$source <- "Structural AME"
lpm        <- sc_baseline_lpm(fit_br)
amce_df    <- data.frame(
  dummy_name = names(stats::coef(lpm)),
  estimate   = unname(stats::coef(lpm)),
  se         = sqrt(diag(stats::vcov(lpm))),
  source     = "LPM AMCE",
  stringsAsFactors = FALSE
)
amce_df$ci_lo <- amce_df$estimate - 1.96 * amce_df$se
amce_df$ci_hi <- amce_df$estimate + 1.96 * amce_df$se
df_both <- rbind(
  ame_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")],
  amce_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")]
)
df_both$label <- ifelse(df_both$dummy_name %in% names(br_labels),
                        br_labels[df_both$dummy_name],
                        df_both$dummy_name)
df_both$label <- factor(df_both$label, levels = rev(unname(br_labels)))
df_both$group <- br_groups[df_both$dummy_name]
df_both$group <- factor(df_both$group, levels = unique(br_groups))

ggplot(df_both, aes(x = estimate, y = label, color = source)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  position = position_dodge(width = 0.5),
                  size = 0.35, fatten = 2) +
  scale_color_manual(values = c("Structural AME" = "#2166AC",
                                "LPM AMCE"        = "#B2182B")) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Effect on Pr(choice)",
       y = NULL, color = NULL,
       title = "Structural AME vs LPM AMCE (probability scale)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        strip.text.y    = element_text(angle = 0, hjust = 0, face = "bold"))
```

On the paper's reporting convention (the effect of a one-percentage-point
rate increase evaluated at a 50% baseline, $\hat\theta_k \times 25$), the
bottom-bracket and top-bracket AMEs are about $-1.6$ and $+0.5$
percentage points of choice probability. Voters punish a rate hike on the
poorest bracket **more than threefold** as hard as they reward an
equal-sized hike on the richest bracket --- a sharp asymmetry that the
homogeneous AMCE reproduces but cannot attribute to any subgroup.

## Individual-level preferences

Ridgeline densities of $\hat\beta_k(\mathbf Z_i)$ per attribute across
respondents. With the `varref` prior (the fit above) the MAP refinement
is properly scaled --- the hybrid view shows clean per-respondent
distributions for every bracket.

```{r}
#| label: br-ridgelines
#| fig-height: 8
plot(fit_br, "beta_ridgelines", groups = br_groups, labels = br_labels)
```

## Structural quantities

### Attribute importance

For continuous attributes the paper's importance share weights each
respondent's squared coefficient by the population variance of the
attribute's **design level set** (e.g. `c(0, 5, 15, 25)` for the bottom
bracket). This is `design = "levels"`, passing the six brackets' level
sets plus the revenue scale; the default `design = "design_variance"`
gives essentially the same shares here.

```{r}
#| label: br-importance
br_levels <- list(
  rate_L10     = c(0, 5, 15, 25),  rate_10_35   = c(5, 15, 25, 35),
  rate_35_85   = c(5, 15, 25, 35), rate_85_175  = c(5, 15, 25, 35),
  rate_175_375 = c(5, 15, 25, 35, 45),
  rate_375P    = c(5, 15, 25, 35, 45, 55),
  revenue_score = c(-2, -1, 0, 1, 2)
)
sc_importance(fit_br, design = "levels", levels = br_levels)
```

```{r}
#| label: br-importance-plot
#| fig-height: 6
plot_importance(fit_br, labels = c(
  rate_L10 = "Under $10k", rate_10_35 = "$10-35k",
  rate_35_85 = "$35-85k", rate_85_175 = "$85-175k",
  rate_175_375 = "$175-375k", rate_375P = "Over $375k",
  revenue_score = "Revenue"
))
```

The importance shares differ sharply by party, which is the paper's
second main finding. Among **Democrats**, the top bracket (>$375k) is the
single most important attribute (about 29%) and the low bracket ($10-35k)
matters much less (about 16%); among **Republicans** the ranking
inverts, with the $10-35k bracket the most important (about 26%) and the
top bracket far less so (about 18%). The $35-85k bracket is roughly twice
as important to Republicans (about 18%) as to Democrats (about 8%). In
short, Democrats weight what the rich pay, Republicans weight what the
middle and working class pay.

```{r}
#| label: br-importance-party
imp_by_party <- function(sel) {
  ip <- sc_importance(fit_br, design = "levels", levels = br_levels,
                      subgroup = sel)
  setNames(round(100 * ip$estimate$share, 1), ip$estimate$attribute)
}
## raw 7-point party id, aligned to the fit's task rows
pid7_resp <- tapply(br2017$resp_pid7, as.character(br2017$respondent),
                    function(x) x[1])
pid7_row  <- pid7_resp[as.character(fit_br$respondent_id)]
pid7_row[pid7_row >= 8] <- 4          # "not sure / other" -> Independent
rbind(
  Democrat    = imp_by_party(pid7_row <= 3),
  Independent = imp_by_party(pid7_row == 4),
  Republican  = imp_by_party(pid7_row >= 5)
)
```

### Direction and intensity

```{r}
#| label: br-di
sc_direction_intensity(fit_br)
```

### Fraction preferring

```{r}
#| label: br-fraction
#| fig-height: 5
plot_fraction(fit_br, groups = br_groups, labels = br_labels)
```

### Heterogeneity test

```{r}
#| label: br-het-test
sc_heterogeneity_test(fit_br, adjust = "bh")
```

```{r}
#| label: br-het-bar
#| fig-height: 5
plot_hetero(fit_br, groups = br_groups, labels = br_labels)
```

### Subgroup analysis by party

Party membership comes from the raw `resp_pid7` convenience column
(Democrat $\le 3$, Independent $= 4$, Republican $\ge 5$), mapped to the
fit's task rows. We reuse `pid7_row` from the importance chunk above.

```{r}
#| label: br-subgroup
#| fig-height: 6
plot_subgroup(
  fit_br,
  subgroup = list(
    Democrat    = pid7_row <= 3,
    Independent = pid7_row == 4,
    Republican  = pid7_row >= 5
  ),
  groups = br_groups, labels = br_labels,
  title = "Subgroup AMCE by party"
)
```

### Tax schedule preferences by party

This plot is unique to the tax-plan conjoint. It extracts the six
bracket coefficients $\hat\beta_k(\mathbf Z_i)$ and shows the
respondent-level preferred-rate schedule in three panels (one per
party). The dark ribbon is the interquartile range across respondents
within each party; the light ribbon is the 10th--90th percentile range.

```{r}
#| label: br-schedule
#| fig-height: 4
#| fig-width: 10

## One row per respondent on the hybrid (MAP-refined) scale
beta_hat <- predict(fit_br, type = "beta")
first_row <- !duplicated(fit_br$respondent_id)
beta_resp <- beta_hat[first_row, ]

## Party assignment from the raw resp_pid7 convenience column
pid <- pid7_row[first_row]
party <- ifelse(pid <= 3, "Democrat",
         ifelse(pid >= 5, "Republican", "Independent"))

## Build schedule data: 6 brackets as x-axis, beta as y-axis
brackets <- c("Under $10k", "$10-35k", "$35-85k",
              "$85-175k", "$175-375k", "Over $375k")
rate_cols <- c("rate_L10", "rate_10_35", "rate_35_85",
               "rate_85_175", "rate_175_375", "rate_375P")

quant_data <- do.call(rbind, lapply(unique(party), function(p) {
  do.call(rbind, lapply(seq_along(brackets), function(j) {
    vals <- beta_resp[party == p, rate_cols[j]]
    data.frame(
      bracket = brackets[j],
      party   = p,
      median  = median(vals),
      q25     = quantile(vals, 0.25),
      q75     = quantile(vals, 0.75),
      q10     = quantile(vals, 0.10),
      q90     = quantile(vals, 0.90)
    )
  }))
}))
quant_data$bracket <- factor(quant_data$bracket, levels = brackets)
quant_data$party   <- factor(quant_data$party,
                              levels = c("Democrat", "Independent", "Republican"))

party_palette <- c(Democrat = "#2166AC", Independent = "gray50",
                   Republican = "#B2182B")

ggplot(quant_data, aes(x = bracket, color = party, fill = party, group = party)) +
  geom_ribbon(aes(ymin = q10, ymax = q90), alpha = 0.15, color = NA) +
  geom_ribbon(aes(ymin = q25, ymax = q75), alpha = 0.30, color = NA) +
  geom_line(aes(y = median), linewidth = 1.0) +
  geom_point(aes(y = median), size = 1.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  facet_wrap(~ party, ncol = 3) +
  scale_color_manual(values = party_palette, guide = "none") +
  scale_fill_manual(values  = party_palette, guide = "none") +
  labs(x = "Income bracket",
       y = expression(hat(beta)[k](Z) ~ "(utility per % point)"),
       title = "Tax schedule preferences, faceted by party") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        strip.text  = element_text(face = "bold"))
```

About **93% of respondents** exhibit progressive tax preferences (a
positive slope from low to high brackets), and **91%** prefer a higher
top-bracket rate than bottom-bracket rate. Progressivity remains the
**majority pattern even among Republicans** (about 90% have a positive
slope). The partisan gap is real but a matter of degree: the Democratic
mean slope (about $+0.021$) is roughly **48% steeper** than the
Republican mean slope (about $+0.014$). Progressivity is the consensus;
the parties differ on how steep, and --- as the importance shares show
--- on which end of the schedule they care about most.

## Validating against the homogeneous-logit AMCE

```{r}
#| label: br-validate
v_br <- sc_validate_amce(fit_br)
v_br
```

Under correct logit specification, the DML $\hat\theta$ on each tax
bracket matches the pooled logit's coefficient on the same bracket
exactly. The pooled correlation is the cleanest summary that the
structural model loses nothing the AMCE provides.
