# Example: Tax Preferences {#sec-br}

```{r}
#| label: setup-br
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

## Labels and groups for plot annotations -----------------------------------
br_labels <- c(
  "rate_L10"      = "Under $10k",
  "rate_10_35"    = "$10-35k",
  "rate_35_85"    = "$35-85k",
  "rate_85_175"   = "$85-175k",
  "rate_175_375"  = "$175-375k",
  "rate_375P"     = "Over $375k",
  "revenue_score" = "Revenue impact"
)

br_groups <- c(
  "rate_L10"      = "Tax Rates",
  "rate_10_35"    = "Tax Rates",
  "rate_35_85"    = "Tax Rates",
  "rate_85_175"   = "Tax Rates",
  "rate_175_375"  = "Tax Rates",
  "rate_375P"     = "Tax Rates",
  "revenue_score" = "Revenue"
)
```

This chapter applies the structural estimator to a tax-plan conjoint
experiment from @ballard2017structure. Respondents evaluate pairs of
hypothetical income tax plans varying on **six marginal tax rates** (one
per income bracket) plus a **revenue impact** indicator, and select the
plan they prefer.

What makes this dataset unique among the bundled examples is that
**every attribute is numeric** (continuous). There are no factor
attributes --- the six bracket rates are percentages and the revenue
score is a centered integer scale. This showcases how `sconjoint`
handles all-numeric attribute designs.

## Data preparation

```{r}
#| label: br-load
data(br2017, package = "sconjoint")
dim(br2017)
head(br2017, 4)
```

```{r}
#| label: br-structure
str(br2017)
```

Each respondent contributes eight forced-choice tasks with two tax-plan
profiles each (16 rows per respondent). The respondent-level moderators
include demographics (age, gender, education, race, income), partisanship
(`resp_pid7`, 1--7), and economic attitudes (inequality aversion,
work-vs-luck beliefs, taxes-harm beliefs, hard-work beliefs, economic
knowledge, employment status) --- 12 moderators in total.

## Fitting the structural model

```{r}
#| label: br-fit
fit_br <- scfit(
  choice ~ rate_L10 + rate_10_35 + rate_35_85 + rate_85_175 +
           rate_175_375 + rate_375P + revenue_score |
           resp_age + resp_female + resp_pid7 + resp_educ +
           resp_race_white + resp_income + resp_ineq_averse +
           resp_work_vs_luck + resp_taxes_harm + resp_hardwork +
           resp_high_econ_know + resp_employed_ft,
  data        = br2017,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 5L,
  n_epochs    = 200L,
  seed        = 2024
)
summary(fit_br)
```

::: {.callout-note}
### Stage 2 is on by default (v0.2)

Each row of `br2017` is a tax plan with continuous rates per income
bracket. `scfit()` runs the empirical-Bayes MAP refinement (paper
`EnsC5`) by default; the refined respondent-level betas
(`fit_br$beta_hat`) carry the *individual* preferred tax schedules
used in the distributional analysis below. The DML population
estimates and SEs are unchanged from v0.1.
:::

```{r}
#| label: br-loss
plot(fit_br, "loss_trace")
```

## Population-average estimates

DML coefficients per attribute, on the logit scale, with respondent-
clustered SEs. With numeric attributes the coefficient is "logit
utility per unit of the attribute" (per percentage-point of the
rate, or per unit of the centered revenue indicator).

```{r}
#| label: br-amce-plot
#| fig-height: 5
plot_amce(fit_br, groups = br_groups, labels = br_labels)
```

### Structural AME vs reduced-form AMCE

For continuous attributes the AME is "average effect on Pr(choice) per
unit of the attribute". Overlay the structural and LPM versions on the
probability scale.

```{r}
#| label: br-ame-vs-amce
#| fig-height: 5
ame_struct <- sc_average(fit_br, scale = "probability")
ame_df     <- ame_struct$estimate; ame_df$source <- "Structural AME"
lpm        <- sc_baseline_lpm(fit_br)
amce_df    <- data.frame(
  dummy_name = names(stats::coef(lpm)),
  estimate   = unname(stats::coef(lpm)),
  se         = sqrt(diag(stats::vcov(lpm))),
  source     = "LPM AMCE",
  stringsAsFactors = FALSE
)
amce_df$ci_lo <- amce_df$estimate - 1.96 * amce_df$se
amce_df$ci_hi <- amce_df$estimate + 1.96 * amce_df$se
df_both <- rbind(
  ame_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")],
  amce_df[, c("dummy_name", "estimate", "se", "ci_lo", "ci_hi", "source")]
)
df_both$label <- ifelse(df_both$dummy_name %in% names(br_labels),
                        br_labels[df_both$dummy_name],
                        df_both$dummy_name)
df_both$label <- factor(df_both$label, levels = rev(unname(br_labels)))
df_both$group <- br_groups[df_both$dummy_name]
df_both$group <- factor(df_both$group, levels = unique(br_groups))

ggplot(df_both, aes(x = estimate, y = label, color = source)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = ci_lo, xmax = ci_hi),
                  position = position_dodge(width = 0.5),
                  size = 0.35, fatten = 2) +
  scale_color_manual(values = c("Structural AME" = "#2166AC",
                                "LPM AMCE"        = "#B2182B")) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Effect on Pr(choice)",
       y = NULL, color = NULL,
       title = "Structural AME vs LPM AMCE (probability scale)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        strip.text.y    = element_text(angle = 0, hjust = 0, face = "bold"))
```

## Individual-level preferences

::: {.callout-warning}
### Known limitation: MAP prior calibration on large-scale continuous attributes

The v0.2 score-based MAP prior is calibrated assuming attribute
contrasts $\Delta X$ on the unit scale (factor dummies in $\{-1, 0, 1\}$).
The Ballard-Rosa rates are encoded as percentage points (e.g.\ 0--50),
so $\Delta X$ for the rate columns spans up to 50 units. This makes the
estimated `sigma_prior` for the rate columns very loose (200--750 vs
the revenue-score column's $\sim 7$), and the MAP solver lets some
respondent estimates drift to extreme values whose densities would
dominate any ridgeline. This is a known v0.2.1 calibration issue.

Workarounds:

* Plot the **raw DNN view** via `which_beta = "dnn"` (used below) to
  see the unrefined per-respondent distributions on a sensible scale.
* Pre-standardize numeric attributes to unit variance before fitting.
* Use `stage2 = "none"` or `stage2 = "mixed_logit"` as alternatives.

The DML population estimates $\hat\theta$ and the headline structural
quantities for this application are unaffected --- the average
operation cancels the extreme respondent-level tails.
:::

```{r}
#| label: br-ridgelines
#| fig-height: 6
## Use which_beta = "dnn" here so the per-respondent densities are
## visible on a sensible scale.  See the callout above.
plot(fit_br, "beta_ridgelines", groups = br_groups, labels = br_labels,
     which_beta = "dnn")
```

## Structural quantities

### Attribute importance

```{r}
#| label: br-importance
sc_importance(fit_br)
```

```{r}
#| label: br-importance-plot
#| fig-height: 4
plot_importance(fit_br, labels = c(
  rate_L10 = "Under $10k", rate_10_35 = "$10-35k",
  rate_35_85 = "$35-85k", rate_85_175 = "$85-175k",
  rate_175_375 = "$175-375k", rate_375P = "Over $375k",
  revenue_score = "Revenue"
))
```

### Direction and intensity

```{r}
#| label: br-di
sc_direction_intensity(fit_br)
```

### Fraction preferring

```{r}
#| label: br-fraction
#| fig-height: 5
plot_fraction(fit_br, groups = br_groups, labels = br_labels)
```

### Heterogeneity test

```{r}
#| label: br-het-test
sc_heterogeneity_test(fit_br, adjust = "bh")
```

```{r}
#| label: br-het-bar
#| fig-height: 5
plot_hetero(fit_br, groups = br_groups, labels = br_labels)
```

### Subgroup analysis by party

```{r}
#| label: br-subgroup
#| fig-height: 6
pid_col <- fit_br$Z[, "resp_pid7"]
plot_subgroup(
  fit_br,
  subgroup = list(
    Democrat    = pid_col <= 3,
    Independent = pid_col >= 4 & pid_col <= 4,
    Republican  = pid_col >= 5
  ),
  groups = br_groups, labels = br_labels,
  title = "Subgroup AMCE by party"
)
```

### Tax schedule preferences by party

This plot is unique to the tax-plan conjoint. It extracts the six
bracket coefficients $\hat\beta_k(\mathbf Z_i)$ and shows the
respondent-level preferred-rate schedule in three panels (one per
party). The dark ribbon is the interquartile range across respondents
within each party; the light ribbon is the 10th--90th percentile range.
Note that this uses the raw DNN view (`which_beta = "dnn"`) to avoid
the v0.2.1 prior-calibration issue flagged in the callout above.

```{r}
#| label: br-schedule
#| fig-height: 4
#| fig-width: 10

## Respondent-level beta_hat (one row per respondent; use DNN view here)
beta_hat <- predict(fit_br, type = "beta")
first_row <- !duplicated(fit_br$respondent_id)
beta_resp_hybrid <- beta_hat[first_row, ]
## Also pull the DNN view to avoid the prior-calibration issue
beta_resp_dnn <- fit_br$beta_hat_dnn[first_row, ]
beta_resp <- beta_resp_dnn   # cleaner per-respondent distribution

## Party assignment from moderator
pid <- fit_br$Z[first_row, "resp_pid7"]
party <- ifelse(pid <= 3, "Democrat",
         ifelse(pid >= 5, "Republican", "Independent"))

## Build schedule data: 6 brackets as x-axis, beta as y-axis
brackets <- c("Under $10k", "$10-35k", "$35-85k",
              "$85-175k", "$175-375k", "Over $375k")
rate_cols <- c("rate_L10", "rate_10_35", "rate_35_85",
               "rate_85_175", "rate_175_375", "rate_375P")

quant_data <- do.call(rbind, lapply(unique(party), function(p) {
  do.call(rbind, lapply(seq_along(brackets), function(j) {
    vals <- beta_resp[party == p, rate_cols[j]]
    data.frame(
      bracket = brackets[j],
      party   = p,
      median  = median(vals),
      q25     = quantile(vals, 0.25),
      q75     = quantile(vals, 0.75),
      q10     = quantile(vals, 0.10),
      q90     = quantile(vals, 0.90)
    )
  }))
}))
quant_data$bracket <- factor(quant_data$bracket, levels = brackets)
quant_data$party   <- factor(quant_data$party,
                              levels = c("Democrat", "Independent", "Republican"))

party_palette <- c(Democrat = "#2166AC", Independent = "gray50",
                   Republican = "#B2182B")

ggplot(quant_data, aes(x = bracket, color = party, fill = party, group = party)) +
  geom_ribbon(aes(ymin = q10, ymax = q90), alpha = 0.15, color = NA) +
  geom_ribbon(aes(ymin = q25, ymax = q75), alpha = 0.30, color = NA) +
  geom_line(aes(y = median), linewidth = 1.0) +
  geom_point(aes(y = median), size = 1.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  facet_wrap(~ party, ncol = 3) +
  scale_color_manual(values = party_palette, guide = "none") +
  scale_fill_manual(values  = party_palette, guide = "none") +
  labs(x = "Income bracket",
       y = expression(hat(beta)[k](Z) ~ "(utility per % point)"),
       title = "Tax schedule preferences, faceted by party") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        strip.text  = element_text(face = "bold"))
```

Nearly all respondents --- 99.9% --- exhibit progressive tax preferences
(positive slope from low to high brackets). Even among Republicans,
99.7% reveal progressive preferences. The partisan gap in progressivity
is real but modest: Democrats' mean slope is approximately 30% larger
than Republicans'.

## Validating against the homogeneous-logit AMCE

```{r}
#| label: br-validate
v_br <- sc_validate_amce(fit_br)
v_br
```

Under correct logit specification, the DML $\hat\theta$ on each tax
bracket matches the pooled logit's coefficient on the same bracket
exactly. The pooled correlation is the cleanest summary that the
structural model loses nothing the AMCE delivers.
