---
title: "Interpreting the fit"
---

# Interpreting the fit {#sec-interpreting}

Before walking through `summary.sc_fit()` it helps to spell out the
conceptual frame this package operates in. The three paragraphs
below are excerpted verbatim from Acharya, Hainmueller, and Xu
(2026), §§2.3 and 3.1. They describe the headline contribution of
the structural model over the standard AMCE and motivate the
direction-versus-intensity decomposition that `sc_direction_intensity()`
reports.

> **The AMCE: one number per attribute.** The standard AMCE provides
> a single number for each attribute level — the average effect on
> the choice probability, pooled across all respondents and
> marginalized over the randomization distribution of all other
> attributes. For example, the AMCE might tell us that "being female
> increases the probability of being chosen by 2 percentage points on
> average." This is a useful summary, but it is one number describing
> the entire population. An AMCE near zero is ambiguous: it could
> mean that no one cares about the attribute, or that half the
> population strongly favors it and the other half strongly opposes
> it, with the effects canceling in the average.

> **The preference vector: a personal profile for each respondent.**
> Our structural model recovers $\hat{\boldsymbol\beta}(\mathbf Z_i)
> \in \mathbb{R}^p$ for each respondent — the complete vector of
> marginal utilities across all attribute levels simultaneously. The
> DNN learns a smooth mapping from respondent characteristics
> $\mathbf Z$ to preferences $\boldsymbol\beta$, capturing nonlinear
> relationships, interactions (e.g., young Republicans may differ
> from old Republicans in ways that young Democrats do not), and
> threshold effects that linear interaction models would miss. The
> key conceptual advance over AMCEs — both unconditional and
> conditional — is that $\hat{\boldsymbol\beta}(\mathbf Z_i)$ gives
> us the *joint* distribution of preferences across attributes for
> each respondent, enabling the structural quantities defined below:
> marginal rates of substitution, compensating differentials,
> counterfactual choice probabilities, and preference polarization —
> all of which require the full preference vector rather than
> one-attribute-at-a-time marginal effects.

> **Direction and intensity.** Polarization highlights a more general
> limitation of the AMCE: it conflates the *direction* of a
> preference (does the respondent favor or oppose the attribute?)
> with its *intensity* (how strongly?). A near-zero AMCE is
> three-ways ambiguous: (a) nobody cares, so there is low intensity
> and no directional preference; (b) most respondents lean in one
> direction but weakly; or (c) strong opposing preferences cancel in
> the average. The structural $\beta_k(\mathbf Z_i)$ decomposes
> these two dimensions: $\operatorname{sign}(\beta_k)$ gives
> direction and $|\beta_k|$ gives intensity, for each respondent
> individually. Preference polarization captures the direction
> dimension; attribute importance captures the intensity dimension.
> The structural model delivers both, and their interaction — for
> instance, an attribute on which most respondents agree in
> direction but a minority disagrees with unusual intensity.

> — Acharya, Hainmueller, and Xu (2026), *Learning Preferences from
> Conjoint Data: A Structural Deep Learning Approach*, §2.3 and §3.1.

## The `summary.sc_fit()` output

`summary(fit)` returns a coefficient table with the DML point
estimate, respondent-clustered standard error, z-statistic,
two-sided p-value, and 95% normal-approximation confidence interval
for every dummy column. The print method also reports the
**DML/iid SE-ratio diagnostic**: the ratio of the clustered SE to the
iid SE. Values meaningfully larger than one indicate that
within-respondent correlation is economically large; values near one
indicate the clustering correction is not doing much work.

```{r}
library(sconjoint)
```

To decompose the population-level average into direction and
intensity, pair the summary table with `sc_direction_intensity()`
from @sec-quantities.

```{r}
cat("interpreting\n")
```

## Subgroup analysis with `sc_subgroup()`

The per-respondent preference vectors $\hat{\boldsymbol\beta}(\mathbf
Z_i)$ support subgroup-conditional summaries at no extra fitting
cost: one simply re-averages the stored held-out $\hat\beta_{i,j}$
over the rows that fall into the subgroup, and uses the same
respondent-clustered SE formula that powers the headline $\hat\theta$.
This is the operational payoff of switching from AMCE to the
structural framework, and it is what `sc_subgroup()` returns.

We demonstrate on a tiny synthetic example so the chapter stays
data-agnostic; the case-study chapters in @sec-case-sw,
@sec-case-gs, and @sec-case-bs apply the same recipe to the bundled
datasets.

```{r sub-toy, message=FALSE}
set.seed(1)
n_resp <- 120L
n_tasks <- 3L
rid <- rep(seq_len(n_resp), each = n_tasks * 2L)
tid <- rep(rep(seq_len(n_tasks), each = 2L), n_resp)
pos <- rep(c(1L, 2L), n_resp * n_tasks)
x1 <- rbinom(length(rid), 1L, 0.5)
x2 <- rbinom(length(rid), 1L, 0.5)
z1 <- rep(rnorm(n_resp), each = n_tasks * 2L)
## True utility: preferences rotate with z1.
u <- (0.6 + 0.5 * z1) * x1 + (-0.3 + 0.4 * z1) * x2
toy <- data.frame(rid, tid, pos, x1, x2, z1)
toy$y <- 0L
for (i in seq_len(n_resp)) for (t in seq_len(n_tasks)) {
  idx <- which(toy$rid == i & toy$tid == t)
  pick <- which.max(u[idx] + -log(-log(runif(length(idx)))))
  toy$y[idx[pick]] <- 1L
}
```

A structural fit followed by `sc_subgroup()` immediately returns
subgroup-conditional $\hat\theta$ and standard errors. Compared to
a homogeneous baseline fitted via `glm()`, the structural model
exposes heterogeneity that the logit baseline averages away.

```{r sub-compare, eval=FALSE}
fit_toy <- scfit(y ~ x1 + x2 | z1,
                 data = toy, respondent = "rid",
                 task = "tid", profile = "pos",
                 K = 2L, n_epochs = 40L, seed = 1L)
## Compare to a homogeneous baseline.
glm_base <- glm(y ~ x1 + x2, data = toy, family = binomial())
coef(glm_base)
## Structural subgroup estimates along z1:
hi <- fit_toy$Z[, "z1"] >  0
sc_subgroup(fit_toy, list(low = !hi, high = hi))
```

The `glm()` baseline returns a single coefficient per attribute;
`sc_subgroup()` returns one set per subgroup, holding the rest of
the design fixed. This side-by-side comparison is the inline
structural-versus-AMCE demonstration we promise in the
introduction.

