---
title: "Plot Options"
---

The `plot.sc_fit()` method ships two plot types. Both return plain
`ggplot` objects, so users can layer additional geoms, facets, or themes
on top.

## Beta ridgelines

The default `"beta_ridgelines"` draws one density ridgeline per
attribute dummy over the held-out per-respondent
$\hat\beta_j(\mathbf Z_i)$. It is the most compact visual summary of
the direction-versus-intensity decomposition: a ridgeline centered away
from zero indicates directional consensus; a wide ridgeline indicates
high intensity heterogeneity.

```{r}
#| label: setup-plot
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)

data(sw2022, package = "sconjoint")
fit_sw <- scfit(
  choice ~ agenda + talent + children + cand_gender + prior_office |
           resp_female + age + pid,
  data       = sw2022,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 3L,
  n_epochs   = 60L,
  seed       = 2024
)
```

```{r}
#| label: plot-ridgelines
#| fig-height: 6
plot(fit_sw, "beta_ridgelines")
```

The ridgeline layer is drawn by `ggridges::geom_density_ridges` when
`ggridges` is available; a boxplot fallback kicks in otherwise so the
plot method works in minimal installations.

## Loss trace

`"loss_trace"` plots the per-fold training loss curves. The curves
should visually level off before the last epoch; if they are still
descending, increase `n_epochs`.

```{r}
#| label: plot-loss
plot(fit_sw, "loss_trace")
```

## Customization with ggplot2

Since both plot types return `ggplot` objects, standard `ggplot2`
modifications apply directly.

```{r}
#| label: plot-custom
library(ggplot2)
plot(fit_sw, "beta_ridgelines") +
  labs(title = "Per-respondent preference distributions",
       subtitle = "Saha-Weeks candidate choice") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```

```{r}
#| label: plot-loss-custom
plot(fit_sw, "loss_trace") +
  labs(title = "Training convergence") +
  theme_bw()
```
