---
title: "Simulation Check"
---

This chapter is a sanity check, not a simulation study. It builds a
small synthetic data-generating process with known true
$\boldsymbol\beta(\mathbf Z)$, fits `scfit()`, and compares the
recovered per-respondent preference vector to the truth. The goal is
pedagogical: to give readers confidence that the estimator works on data
whose ground truth they can inspect directly.

```{r}
#| label: setup-sim
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(42)
torch::torch_manual_seed(42)
library(sconjoint)
library(ggplot2)
```

## Data-generating process

We simulate a conjoint experiment with $M = 100$ respondents, $T = 4$
tasks per respondent, and two profiles per task (800 rows total). Three
binary attributes $x_1, x_2, x_3$ are drawn independently as
Bernoulli(0.5). Two continuous moderators $z_1, z_2 \sim N(0,1)$
determine each respondent's preference vector through a simple linear
map:

$$
\beta_1(\mathbf Z_i) = 0.5 + 0.3\,z_{1i}, \quad
\beta_2(\mathbf Z_i) = -0.8 + 0.2\,z_{2i}, \quad
\beta_3(\mathbf Z_i) = 0.3.
$$

For each task, the respondent sees two profiles $A$ and $B$. The
probability of choosing $A$ over $B$ is
$\Lambda(\Delta\mathbf X \cdot \boldsymbol\beta(\mathbf Z_i))$
where $\Delta\mathbf X = \mathbf X_A - \mathbf X_B$ and $\Lambda$
is the logistic CDF.

```{r}
#| label: sim-dgp
M <- 100L     # respondents
TT <- 4L     # tasks per respondent
set.seed(42)

## Respondent-level moderators
z1 <- rnorm(M)
z2 <- rnorm(M)

## True beta(Z) for each respondent
beta_true <- cbind(
  b1 = 0.5 + 0.3 * z1,
  b2 = -0.8 + 0.2 * z2,
  b3 = rep(0.3, M)
)

## Build long-format data frame
rows <- vector("list", M * TT * 2L)
idx <- 0L
for (i in seq_len(M)) {
  for (tt in seq_len(TT)) {
    ## Draw attribute profiles
    xA <- rbinom(3, 1, 0.5)
    xB <- rbinom(3, 1, 0.5)
    deltaX <- xA - xB
    V <- sum(deltaX * beta_true[i, ])
    choiceA <- rbinom(1, 1, plogis(V))
    ## Profile A
    idx <- idx + 1L
    rows[[idx]] <- data.frame(
      respondent = i, task = tt, profile = 1L,
      choice = choiceA,
      x1 = xA[1], x2 = xA[2], x3 = xA[3],
      z1 = z1[i], z2 = z2[i]
    )
    ## Profile B
    idx <- idx + 1L
    rows[[idx]] <- data.frame(
      respondent = i, task = tt, profile = 2L,
      choice = 1L - choiceA,
      x1 = xB[1], x2 = xB[2], x3 = xB[3],
      z1 = z1[i], z2 = z2[i]
    )
  }
}
sim_data <- do.call(rbind, rows)
dim(sim_data)
head(sim_data, 8)
```

## Fitting the structural model

```{r}
#| label: sim-fit
fit_sim <- scfit(
  choice ~ x1 + x2 + x3 | z1 + z2,
  data       = sim_data,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 3L,
  n_epochs   = 100L,
  seed       = 42
)
summary(fit_sim)
```

## Population-average recovery

The DML point estimates $\hat\theta$ should approximate the
population-average of the true $\beta(\mathbf Z)$: $E[\beta_1] = 0.5$,
$E[\beta_2] = -0.8$, $E[\beta_3] = 0.3$.

```{r}
#| label: sim-theta
theta_hat <- coef(fit_sim)
theta_true <- c(x1 = 0.5, x2 = -0.8, x3 = 0.3)
data.frame(
  attribute  = names(theta_hat),
  true_theta = theta_true,
  estimated  = round(unname(theta_hat), 3)
)
```

Signs and rough magnitudes should line up. Exact recovery is not
expected with only $M = 100$ respondents and $T = 4$ tasks.

## Individual-level recovery

The structural payoff is the per-respondent $\hat\beta(\mathbf Z_i)$
matrix. We compare the estimated values to the true DGP values.

```{r}
#| label: sim-beta-recovery
beta_hat_all <- predict(fit_sim)
dim(beta_hat_all)

## De-duplicate to respondent level (beta depends only on Z_i)
first_row <- !duplicated(fit_sim$respondent_id)
beta_hat <- beta_hat_all[first_row, ]

## Correlation between true and estimated beta for each attribute
cor_b1 <- cor(beta_true[, 1], beta_hat[, 1])
cor_b2 <- cor(beta_true[, 2], beta_hat[, 2])
data.frame(
  attribute   = c("x1 (beta1)", "x2 (beta2)"),
  correlation = round(c(cor_b1, cor_b2), 3)
)
```

Attribute $x_3$ has constant $\beta_3 = 0.3$ (no heterogeneity), so
per-respondent correlation is not informative for it.

## Scatter plot: true vs estimated $\beta_1(\mathbf Z)$

```{r}
#| label: sim-scatter
#| fig-height: 5
df_compare <- data.frame(
  true      = beta_true[, 1],
  estimated = beta_hat[, 1]
)

ggplot(df_compare, aes(x = true, y = estimated)) +
  geom_point(alpha = 0.5, color = "#2166AC") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "#E41A1C") +
  labs(x = expression("True " * beta[1](Z)),
       y = expression("Estimated " * beta[1](Z)),
       title = "Individual-level recovery: attribute x1") +
  theme_minimal(base_size = 12)
```

Points clustering around the 45-degree line confirm that the structural
estimator recovers the direction and magnitude of individual-level
heterogeneity in $\beta_1(\mathbf Z)$.

## Loss trace

```{r}
#| label: sim-loss
plot(fit_sim, "loss_trace")
```
