# Plot Options {#sec-plots}

```{r}
#| label: setup-plot
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
set.seed(2024)
torch::torch_manual_seed(2024)
library(sconjoint)
library(ggplot2)

data(sw2022, package = "sconjoint")
fit <- scfit(
  choice ~ agenda + talent + children + cand_gender + prior_office |
           resp_female + age + pid,
  data       = sw2022,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 3L,
  n_epochs   = 60L,
  seed       = 2024
)
```

`sconjoint` ships six plot functions. All return plain `ggplot` objects
that can be further customized with `+`. This chapter is
**parameter-first**: it lists every customization option, grouped by
purpose, and demonstrates each in isolation on a single pre-fitted
object.

## Plot Functions

| Function | Purpose |
|----------|---------|
| `plot(fit, "beta_ridgelines")` | Per-respondent $\hat\beta(Z)$ density ridgelines |
| `plot(fit, "loss_trace")` | Per-fold training loss curves |
| `plot_amce(fit)` | Population-average $\hat\theta$ coefficient plot |
| `plot_fraction(fit)` | Fraction favor / oppose diverging bar chart |
| `plot_hetero(fit)` | Preference heterogeneity bar chart |
| `plot_subgroup(fit, subgroup)` | Subgroup AMCE comparison |

## Parameter Reference

All six functions share a common set of customization parameters.
Function-specific parameters are noted in parentheses.

| Parameter | Applies to | Purpose | Default |
|-----------|-----------|---------|---------|
| `dummies` | all | Character vector: which dummies to show and in what order | `NULL` (all) |
| `labels` | all | Named character vector: display names for dummies | `NULL` (raw names) |
| `title` | all | Main title | sensible default |
| `xlab` / `ylab` | all | Custom axis labels | `NULL` |
| `xlim` | all | Numeric(2): x-axis limits | `NULL` |
| `theme.bw` | all | Use `theme_bw()` instead of `theme_minimal()` | `FALSE` |
| `gridOff` | all | Remove grid lines | `FALSE` |
| `legendOff` | all | Hide the legend | `FALSE` |
| `legend.pos` | all | Legend position: `"bottom"`, `"right"`, `"none"`, or `c(x, y)` | varies |
| `cex.main` | all | Title font size multiplier | `NULL` |
| `cex.axis` | all | Tick label font size multiplier | `NULL` |
| `cex.lab` | all | Axis title font size multiplier | `NULL` |
| `color` | `plot_amce` | Point/line color | `"#E41A1C"` |
| `size` / `fatten` | `plot_amce`, `plot_subgroup` | Point size and fatten factor | `0.4` / `2` |
| `colors` | `plot_fraction`, `plot_subgroup` | Named color vector | defaults |
| `alpha` / `bar.width` | `plot_fraction` | Bar transparency and width | `0.85` / `0.65` |
| `gradient` | `plot_hetero` | `c(low, high)` fill gradient colors | blue scale |
| `sig.color` / `sig.shape` | `plot_hetero` | Marker for significant dummies | `"#B2182B"` / `18` |
| `dodge.width` | `plot_subgroup` | Spacing between groups | `0.6` |

## Selecting Variables

Use `dummies` to show only a subset of attribute dummies. The display
order follows the order of the `dummies` vector.

```{r}
#| label: plot-select-vars
#| fig-height: 3.5
plot_amce(fit, dummies = c("cand_genderfemale", "agendaprogressive",
                            "prior_officeus_house"))
```

This works identically across all plot functions:

```{r}
#| label: plot-select-fraction
#| fig-height: 3
plot_fraction(fit, dummies = c("cand_genderfemale", "agendaprogressive"))
```

## Custom Labels

Use `labels` to replace raw dummy names with human-readable text. Pass
a named character vector; any dummy not in `labels` keeps its original
name.

```{r}
#| label: plot-labels
#| fig-height: 4
my_labels <- c(
  agendaprogressive  = "Progressive agenda",
  agendaconservative = "Conservative agenda",
  cand_genderfemale  = "Female candidate",
  prior_officeus_house = "U.S. House experience"
)

plot_amce(fit, labels = my_labels)
```

Combined with `dummies` for a focused, publication-ready figure:

```{r}
#| label: plot-labels-focused
#| fig-height: 3
plot_amce(fit,
          dummies = c("cand_genderfemale", "agendaprogressive"),
          labels = c(cand_genderfemale = "Female candidate",
                     agendaprogressive = "Progressive agenda"),
          title = "Key attributes")
```

## Colors

### AMCE color

```{r}
#| label: plot-amce-color
#| fig-height: 4
plot_amce(fit, color = "#2166AC", title = "AMCE (blue)")
```

### Fraction colors

```{r}
#| label: plot-fraction-color
#| fig-height: 4
plot_fraction(fit,
              colors = c(Favor = "#1B9E77", Oppose = "#D95F02"),
              title = "Custom fraction colors")
```

### Heterogeneity gradient

```{r}
#| label: plot-hetero-gradient
#| fig-height: 4
plot_hetero(fit,
            gradient = c(low = "#FEE0D2", high = "#DE2D26"),
            sig.color = "black",
            title = "Red gradient with black markers")
```

### Subgroup colors

```{r}
#| label: plot-subgroup-color
#| fig-height: 4
fem <- fit$Z[, "resp_female"] > 0.5
plot_subgroup(fit,
              subgroup = list(Female = fem, Male = !fem),
              colors = c(Female = "#7570B3", Male = "#1B9E77"),
              title = "Custom subgroup colors")
```

## Titles and Axis Labels

```{r}
#| label: plot-titles
#| fig-height: 4
plot_amce(fit,
          title = "Saha-Weeks candidate choice",
          xlab  = "Logit-scale AMCE",
          ylab  = "Attribute level")
```

## Theme and Layout

### Black-and-white theme

```{r}
#| label: plot-theme-bw
#| fig-height: 4
plot_amce(fit, theme.bw = TRUE)
```

### Remove grid lines

```{r}
#| label: plot-grid-off
#| fig-height: 4
plot_fraction(fit, gridOff = TRUE)
```

### Font sizes

Use `cex.main`, `cex.axis`, and `cex.lab` as multipliers on the base
font sizes (14, 11, 12 pt respectively).

```{r}
#| label: plot-font-size
#| fig-height: 4
plot_amce(fit, cex.main = 1.4, cex.axis = 0.9, cex.lab = 1.1)
```

### Legend position

```{r}
#| label: plot-legend-pos
#| fig-height: 4
fem <- fit$Z[, "resp_female"] > 0.5
plot_subgroup(fit,
              subgroup = list(Female = fem, Male = !fem),
              legend.pos = "right")
```

### Hide legend

```{r}
#| label: plot-legend-off
#| fig-height: 4
plot_fraction(fit, legendOff = TRUE)
```

### Axis limits

```{r}
#| label: plot-xlim
#| fig-height: 4
plot_amce(fit, xlim = c(-1.5, 1.5))
```

## Ridgelines and Loss Trace

The `plot.sc_fit()` method accepts the same customization parameters
via `...`.

```{r}
#| label: plot-ridge-custom
#| fig-height: 5
plot(fit, "beta_ridgelines",
     dummies = c("agendaprogressive", "cand_genderfemale",
                  "prior_officeus_house"),
     labels = c(agendaprogressive = "Progressive",
                cand_genderfemale = "Female",
                prior_officeus_house = "U.S. House"),
     title = "Selected attributes")
```

```{r}
#| label: plot-loss-custom
plot(fit, "loss_trace", title = "Training convergence", theme.bw = TRUE)
```

## Composing with ggplot2

Since all plot functions return `ggplot` objects, standard `ggplot2`
modifications apply via `+`:

```{r}
#| label: plot-compose
#| fig-height: 4
plot_amce(fit) +
  labs(subtitle = "Saha-Weeks (2022) candidate choice") +
  theme(plot.subtitle = element_text(color = "gray40", size = 10))
```

This lets users add annotations, change coordinate systems, facet, or
compose multiple plots with `patchwork` without modifying the plot
functions themselves.
