---
title: "Structural quantities"
---

# Structural quantities {#sec-quantities}

The Tier A structural quantities shipped in v0.1 are all defined as
population averages of a per-respondent function of the estimated
preference vector $\hat{\boldsymbol\beta}(\mathbf Z_i)$. Every
quantity returns an `sc_quantity` object carrying a point estimate,
a respondent-clustered standard error, a normal-approximation 95%
confidence interval, and a per-function `details` list. All of the
examples below are illustrative; they are not executed during book
rendering.

```{r}
library(sconjoint)
```

## `sc_mrs()` — marginal rate of substitution

The trimmed mean over respondents of
$\hat\beta_j(Z_i) / \hat\beta_k(Z_i)$, clipped at the
user-specified quantiles of the raw ratio. Intuitively: by how many
units of attribute $k$ must we compensate a respondent so that she
is indifferent to a one-unit increase in attribute $j$?

## `sc_counterfactual()` — pairwise choice probability

For two user-specified profiles $A$ and $B$, reports
$|S|^{-1}\sum_{i\in S} G\!\bigl((X_A - X_B)^\top \hat\beta(Z_i)\bigr)$
— the average probability that respondents in the subgroup choose
$A$ over $B$.

## `sc_wtp()` — willingness-to-pay

A thin sign-flipped wrapper around `sc_mrs()` with a cost attribute
playing the role of the denominator. Positive WTP when the cost
attribute has negative average utility.

## `sc_importance()` — attribute importance

Per-respondent within-block variance decomposition
$\mathrm{share}_{ia} = V_{ia} / \sum_{a'} V_{ia'}$, averaged across
the subgroup. The shares sum to one by construction and answer
"how much of respondent $i$'s preference variability is driven by
attribute $a$?".

## `sc_polarization()` — polarization index

Per dummy, $1 - |\mathrm{frac}^+_j - \mathrm{frac}^-_j|$. Near 1 when
the population is evenly split in direction (polarized); near 0 when
respondents agree in direction.

## `sc_fraction_preferring()` — tail probabilities

Per dummy, $\Pr_{i\in S}\{\hat\beta_j(Z_i) > \tau\}$ and the
symmetric lower-tail version. With the default $\tau = 0$ this is
the fraction of respondents preferring each non-reference level.

## `sc_optimal_profile()` — greedy argmax

Greedy per-attribute argmax over subgroup-averaged
$\hat\beta_{aj}$, assembled into a dummy vector $X^\star$, plus the
mean choice probability of $X^\star$ against the all-reference
profile.

## `sc_direction_intensity()` — bivariate decomposition

Per dummy, reports both $d_j = \overline{\mathrm{sign}(\hat\beta_j)}$
(the direction) and $u_j = \overline{|\hat\beta_j|}$ (the
intensity). The return object is an `sc_quantity_bivariate` carrying
two `sc_quantity` sub-objects plus a scalar separation diagnostic.

## `sc_heterogeneity_test()` — per-dummy variance test

One-sided test of $\mathrm{Var}(\beta_j(\mathbf Z)) > 0$ per dummy
with optional Holm or BH multiplicity correction; the test statistic
and clustered SE of the variance estimator are taken from the
prototype (`07_quantities_of_interest.R` L215–251).

```{r}
cat("quantities overview\n")
```

## Tier B quantities

The Tier B quantities are advanced summaries of the per-respondent
$\hat\beta(\mathbf Z_i)$ matrix that users routinely reach for once
they are comfortable with the Tier A tool-set.

### `sc_subgroup()` — subgroup-conditional estimates

Re-averages $\hat\beta(\mathbf Z_i)$ over a user-specified subset
of respondents and returns respondent-clustered SEs computed on
that subset. The `subgroup` argument accepts a logical vector, a
row index, a length-1 character naming a $\mathbf Z$ column (the
greater-than-median half is selected), a function of $\mathbf Z$,
or a named list of any of the above.

$$
\hat\theta^{\text{sub}}_j = \frac{1}{|S|} \sum_{i\in S} \hat\beta_j(\mathbf Z_i),
\qquad
\hat{\mathrm{se}}^{\text{sub}}_j = \mathrm{cluster\_se}\big\{\hat\beta_j(\mathbf Z_i)\big\}_{i\in S}.
$$

```{r tierb-subgroup, eval=FALSE}
sc_subgroup(fit, subgroup = list(female = fit$Z[, "resp_female"] > 0.5,
                                 male   = fit$Z[, "resp_female"] < 0.5))
```

### `sc_compensating()` — compensating differentials

For a benefit dummy $j$ and a cost dummy $k$, the compensating
differential is the trimmed respondent-mean of $-\hat\beta_j(\mathbf
Z_i)/\hat\beta_k(\mathbf Z_i)$ (trimmed at the empirical 1% and 99%
quantiles to guard against division by small cost coefficients).
The returned object also reports `frac_compensated` — the share
of respondents for whom $\hat\beta_j + \hat\beta_k \ge 0$.

$$
\widehat{\mathrm{comp}}_{j,k} = \mathrm{trimmed\_mean}\!\left\{
  -\hat\beta_j(\mathbf Z_i)/\hat\beta_k(\mathbf Z_i)
\right\}_{i=1}^N.
$$

```{r tierb-compensating, eval=FALSE}
sc_compensating(fit, benefit = "benefit_attr:level",
                     cost    = "cost_attr:level")
```

When the cost attribute is numeric (as in the Bechtel-Scheve case
study) the compensating differential is reported on the original
scale of the cost attribute.

### `sc_clusters()` — preference clustering

Runs k-means on the (optionally column-standardized) $\hat\beta$
matrix and returns the length-$N$ cluster assignment, the
$k\times p$ matrix of cluster centers on the original $\hat\beta$
scale, the per-cluster within sum of squares, and the sizes. The
partition is reproducible whenever `seed` is supplied; `nstart`
controls the number of k-means restarts used to protect against
local optima. Hierarchical clustering (`method = "hclust"`) is
available as an alternative.

```{r tierb-clusters, eval=FALSE}
sc_clusters(fit, k = 3L, seed = 1L)
```

