---
title: "Structural quantities"
---

# Structural quantities {#sec-quantities}

The Tier A structural quantities shipped in v0.1 are all defined as
population averages of a per-respondent function of the estimated
preference vector $\hat{\boldsymbol\beta}(\mathbf Z_i)$. Every
quantity returns an `sc_quantity` object carrying a point estimate,
a respondent-clustered standard error, a normal-approximation 95%
confidence interval, and a per-function `details` list. All of the
examples below are illustrative; they are not executed during book
rendering.

```{r}
library(sconjoint)
```

## `sc_mrs()` — marginal rate of substitution

The trimmed mean over respondents of
$\hat\beta_j(Z_i) / \hat\beta_k(Z_i)$, clipped at the
user-specified quantiles of the raw ratio. Intuitively: by how many
units of attribute $k$ must we compensate a respondent so that she
is indifferent to a one-unit increase in attribute $j$?

## `sc_counterfactual()` — pairwise choice probability

For two user-specified profiles $A$ and $B$, reports
$|S|^{-1}\sum_{i\in S} G\!\bigl((X_A - X_B)^\top \hat\beta(Z_i)\bigr)$
— the average probability that respondents in the subgroup choose
$A$ over $B$.

## `sc_wtp()` — willingness-to-pay

A thin sign-flipped wrapper around `sc_mrs()` with a cost attribute
playing the role of the denominator. Positive WTP when the cost
attribute has negative average utility.

## `sc_importance()` — attribute importance

Per-respondent within-block variance decomposition
$\mathrm{share}_{ia} = V_{ia} / \sum_{a'} V_{ia'}$, averaged across
the subgroup. The shares sum to one by construction and answer
"how much of respondent $i$'s preference variability is driven by
attribute $a$?".

## `sc_polarization()` — polarization index

Per dummy, $1 - |\mathrm{frac}^+_j - \mathrm{frac}^-_j|$. Near 1 when
the population is evenly split in direction (polarized); near 0 when
respondents agree in direction.

## `sc_fraction_preferring()` — tail probabilities

Per dummy, $\Pr_{i\in S}\{\hat\beta_j(Z_i) > \tau\}$ and the
symmetric lower-tail version. With the default $\tau = 0$ this is
the fraction of respondents preferring each non-reference level.

## `sc_optimal_profile()` — greedy argmax

Greedy per-attribute argmax over subgroup-averaged
$\hat\beta_{aj}$, assembled into a dummy vector $X^\star$, plus the
mean choice probability of $X^\star$ against the all-reference
profile.

## `sc_direction_intensity()` — bivariate decomposition

Per dummy, reports both $d_j = \overline{\mathrm{sign}(\hat\beta_j)}$
(the direction) and $u_j = \overline{|\hat\beta_j|}$ (the
intensity). The return object is an `sc_quantity_bivariate` carrying
two `sc_quantity` sub-objects plus a scalar separation diagnostic.

## `sc_heterogeneity_test()` — per-dummy variance test

One-sided test of $\mathrm{Var}(\beta_j(\mathbf Z)) > 0$ per dummy
with optional Holm or BH multiplicity correction; the test statistic
and clustered SE of the variance estimator are taken from the
prototype (`07_quantities_of_interest.R` L215–251).

```{r}
cat("quantities overview\n")
```
