# Structural estimands {#sec-estimands}

```{r}
#| label: setup-estimands
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
library(sconjoint)
```

The application chapters fit one model and then read several quantities
off it: population averages, the distribution of preferences across
respondents, and individual-level summaries. This chapter is the
reference for those quantities. It lists each estimand, the function that
returns it, what it targets, and how its standard errors are computed,
so you can look up the right call without re-reading a worked example.

The quantities form a hierarchy by how much the design has to support to
recover them. Population averages (the AMCE and its relatives) are
available from almost any forced-choice design. Distributional quantities
(how preferences spread across the electorate) need enough tasks per
respondent and enough signal in the moderators $\mathbf Z$ to separate
real heterogeneity from noise. Individual-level quantities (one
respondent's preference vector, their personal trade-offs) ask for more
still. @acharya2026structural set out which quantities a given design can
recover; the [design diagnostic](#sec-design-diagnostic) at the end of
this chapter estimates which tier your own design reaches. The code below
is reference syntax (`eval: false`); see the application chapters for the
same calls run on real fits.

## A map of the quantities

| Estimand | Function | Targets | Inference | `which_beta` applies |
|----|----|----|----|:--:|
| AMCE $\hat\theta_k$ | `coef()`, `summary()`, `plot_amce()` | population part-worth (logit) | DML, clustered SE | no |
| Average marginal effect | `sc_average()` | population AME (logit or probability) | delta-method on $\hat\theta$ | --- |
| Structural AME | `sc_ame()` | population AME (probability) | debiased orthogonal score | no |
| Counterfactual vote share | `sc_counterfactual()`, `sc_voteshare_contrast()` | $\Pr(\text{choice})$ for a profile pair or contrast | debiased orthogonal score | no |
| Population MRS / WTP | `sc_mrs(estimand = "population")`, `sc_wtp(...)` | ratio $\theta_j/\theta_k$ | debiased, Fieller interval | no |
| Attribute importance | `sc_importance()`, `plot_importance()` | utility-variance share per attribute | plug-in, or debiased (`vartype = "orthogonal"`) | yes (plug-in) |
| Fraction preferring | `sc_fraction_preferring()`, `plot_fraction()` | share with $\hat\beta_k > 0$ | plug-in or wild bootstrap | yes |
| Polarization | `sc_polarization()` | dispersion of $\hat\beta_k(\mathbf Z)$ | plug-in or wild bootstrap | yes |
| Direction and intensity | `sc_direction_intensity()` | sign vs. magnitude split | plug-in | yes |
| Compensating differential | `sc_compensating()` | benefit/cost trade-off | plug-in, clustered SE | yes |
| Heterogeneity test | `sc_heterogeneity_test()`, `plot_hetero()` | $\mathrm{Var}(\hat\beta_k(\mathbf Z)) > 0$ | one-sided test | yes |
| Per-respondent betas | `predict()` | each respondent's $\hat\beta(\mathbf Z_i)$ | model-based | yes |
| Individual MRS / WTP | `sc_mrs()`, `sc_wtp()` (default) | respondent-averaged trimmed ratio | clustered SE | yes |
| Optimal profile | `sc_optimal_profile()` | utility-maximizing profile | model-based | yes |
| Preference clusters | `sc_clusters()` | latent preference types | k-means on $\hat\beta(\mathbf Z)$ | yes |
| Subgroup AMCE | `sc_subgroup()`, `plot_subgroup()` | $\hat\beta$ re-averaged over a subset | clustered SE | yes |

The `which_beta` column records whether the quantity reads the
per-respondent coefficients (and so can be computed from either the
Stage-2 hybrid view or the raw DNN view; see
[below](#sec-which-beta)). The population-average estimands marked "no"
are computed from the Stage-1 cross-fit and do not depend on that choice.

## Population averages

These are the quantities a standard AMCE analysis already reports, plus
their structural counterparts. They are population averages, computed
from the Stage-1 cross-fitted first stage, and they come with
double/debiased machine-learning (DML) inference.

The population-average parameter $\hat\theta$ (from `coef()` /
`summary()`) carries an orthogonal-score correction with a
respondent-clustered standard error. The same correction, implemented for
the paper's Supplementary Materials (Section C), extends to the other
average quantities, so several `sc_*` functions return debiased
confidence intervals rather than plug-in ones:

| Quantity | Debiased option |
|---|---|
| `sc_counterfactual()` | `vartype = "orthogonal"` (the **default**); `vartype = "plugin"` for the v0.1 plug-in |
| `sc_voteshare_contrast()` | always debiased; takes a raw contrast vector in attribute-dummy space |
| `sc_ame()` | always debiased (the structural AME on the probability scale) |
| `sc_mrs()`, `sc_wtp()` | `estimand = "population"` (debiased ratio `theta_j/theta_k` with a Fieller interval) |
| `sc_importance()` | `vartype = "orthogonal"` (opt-in; the plug-in shares stay the default) |

```{r}
#| label: population-example
#| eval: false
coef(fit)                                                # DML theta, clustered SE
sc_average(fit, scale = "probability")                   # population AME (delta-method)
sc_ame(fit, "attr_level")                                # debiased structural AME
sc_counterfactual(fit, A = list(...), B = list(...))     # debiased by default
sc_voteshare_contrast(fit, contrast = c(...))            # debiased vote share for a contrast
sc_mrs(fit, "attr_a", "attr_b", estimand = "population") # debiased theta_j/theta_k
sc_importance(fit, vartype = "orthogonal")               # debiased shares (opt-in)
```

The debiased estimates use the Stage-1 cross-fitted first stage and are
population averages, so `subgroup` and `which_beta` do not apply to them.
Applied to $\hat\theta$ itself the orthogonal score reproduces the
average-parameter estimate and its clustered standard error exactly. For
`sc_importance()`, the debiased shares are a ratio of orthogonal-score
numerators and can fall outside $[0, 1]$ when the first stage is noisy,
which is why the always-in-$[0, 1]$ plug-in shares remain the default.

## Distributional quantities

These describe how preferences are spread across respondents rather than
their average. They read the per-respondent coefficients
$\hat\beta_k(\mathbf Z_i)$, so they accept `which_beta`, and the
fixed-$T$ summaries (fraction preferring, polarization) accept a
respondent-cluster wild bootstrap for honest standard errors.

```{r}
#| label: distributional-example
#| eval: false
sc_fraction_preferring(fit)                              # share with beta_k > 0
sc_fraction_preferring(fit, se_method = "wild_bootstrap", n_boot = 1000)
sc_polarization(fit, se_method = "wild_bootstrap", n_boot = 1000)
sc_direction_intensity(fit)                              # sign vs. magnitude
sc_compensating(fit, benefit = "attr_a", cost = "attr_b")
sc_heterogeneity_test(fit, adjust = "bh")                # Var(beta_k(Z)) > 0
```

::: {.callout-warning}
### A fixed-$T$ shrinkage caveat

With few tasks per respondent, the Stage-2 MAP update pulls each
respondent's coefficients toward the mean stage, so a hybrid-view
fraction or polarization estimate is conservative: the recovered spread
is narrower than the truth. Sanity-check distributional claims with
`which_beta = "dnn"` (the unrefined view). If the qualitative conclusion
flips between the two, the design is too sparse to support an
individual-level reading, and you should report only the population
averages. The [design diagnostic](#sec-design-diagnostic) makes this
concrete.
:::

## Individual-level quantities

These summarize a single respondent's recovered preferences, or build on
the full per-respondent coefficient matrix. They are model-based: their
credibility rests on the first stage having learned
$\hat\beta(\mathbf Z)$ well, which needs both enough tasks per respondent
and moderators $\mathbf Z$ that explain a real share of the
heterogeneity.

```{r}
#| label: individual-example
#| eval: false
beta <- predict(fit)                                     # N_resp x p coefficient matrix
predict(fit, newdata = new_Z, type = "beta")            # forward pass on new respondents
sc_mrs(fit, "attr_a", "attr_b")                          # respondent-averaged trimmed ratio
sc_optimal_profile(fit)                                  # utility-maximizing profile
sc_clusters(fit, k = 3L)                                 # latent preference types
sc_subgroup(fit, list(GroupA = mask, GroupB = !mask))    # beta re-averaged over a subset
```

The trimmed ratio in `sc_mrs()` / `sc_wtp()` (the default
respondent-level estimand, as opposed to `estimand = "population"`) is
unbounded when the denominator coefficient is near zero, so it trims at
$\{0.01, 0.99\}$ by default; inspect the ridgelines before reading an
individual MRS literally.

## Choosing the coefficient view: `which_beta` {#sec-which-beta}

Every distributional and individual-level `sc_*` quantity accepts
`which_beta = c("hybrid", "dnn")`. The default `"hybrid"` reads
`fit$beta_hat` (whatever Stage 2 produced). Passing `"dnn"` reads
`fit$beta_hat_dnn` (the raw Stage-1 view). When `stage2 = "none"` the two
are numerically identical and the helper silently falls back.

```{r}
#| label: which-beta-example
#| eval: false
## Same fit, two views of the same distribution
sc_fraction_preferring(fit)                     # MAP-shrunk
sc_fraction_preferring(fit, which_beta = "dnn") # raw DNN
```

When the paper text says "the structural model reveals that X% of
respondents prefer Y", it means the MAP-shrunk fraction (the `"hybrid"`
default). The raw DNN fraction is usually wider because unrefined
respondent-level coefficients have more sampling noise. The two views are
the basis of the shrinkage sanity check described above.

## Survey-weighted estimands

By default, `sconjoint` targets the unweighted sample of respondents:
each respondent contributes equally after averaging over that
respondent's conjoint tasks. If the original survey supplies respondent
weights and the target is a weighted population, pass those weights to
`scfit()` with `respondent_weights`.

```{r}
#| label: survey-weight-example
#| eval: false
fit_w <- scfit(
  formula,
  data = dat,
  respondent = "respondent_id",
  task = "task_id",
  profile = "profile_id",
  respondent_weights = "survey_weight",
  seed = 42
)
```

The weight can be a column name in `data` or a numeric vector of length
`nrow(data)`. It must be finite, non-negative, and constant within
respondent. The package normalizes the weights across respondents and
then uses them in the respondent-level aggregation step:

$$
\hat\theta_w
=
\sum_{i=1}^{N} a_i\,\bar\psi_i,
\qquad
a_i=\frac{w_i}{\sum_{j=1}^{N}w_j},
\qquad
\bar\psi_i=\frac{1}{T_i}\sum_{t=1}^{T_i}\psi_{it}.
$$

The clustered variance uses the same weighted respondent contributions,
with the usual finite-cluster factor:

$$
\widehat V_w
=
\frac{N}{N-1}
\sum_{i=1}^{N}
a_i^2(\bar\psi_i-\hat\theta_w)(\bar\psi_i-\hat\theta_w)'.
$$

Survey weights do **not** reweight the first-stage learner. The DNN,
elastic net, or GRF still learns the conditional preference function
$f(\mathbf Z)$ from the observed respondent-task rows. The weights enter
only after the orthogonal scores or respondent-level plug-in quantities
have been formed. This is the right operation when the issue is
target-population composition rather than a different training loss.

The same weighted aggregation is used by the production-facing
quantities: `coef()` / `summary()`, `sc_average()`,
`sc_counterfactual()`, `sc_mrs()` / `sc_wtp()`, `sc_compensating()`,
`sc_fraction_preferring()`, `sc_polarization()`, `sc_importance()`,
`sc_subgroup()`, and `sc_direction_intensity()`.

```{r}
#| label: survey-weight-quantities
#| eval: false
coef(fit_w)
sc_counterfactual(fit_w, A = list(...), B = list(...))
sc_compensating(fit_w, benefit = "co_partisan", cost = "journalists")
sc_fraction_preferring(fit_w, se_method = "wild_bootstrap", n_boot = 1000)
```

For comparison or diagnostics, fit the model once without weights and
once with weights. A material difference means the estimand is sensitive
to survey composition; a small difference supports reporting the
unweighted structural estimate with a weighting robustness check.

## Design diagnostic: which tiers does my design support? {#sec-design-diagnostic}

`sc_design_diagnostic()` estimates per-coefficient $\hat R^2_{Z,k}$ from
the MAP posterior and maps the design to four recovery tiers per the
paper's design-guidance heuristics:

| Tier | Condition | Quantities recoverable |
|------|-----------|------------------------|
| mean & aggregate | any reasonable design | AMCE, average vote shares |
| distributional | T ≥ 5 and $\bar R^2_Z$ ≥ 0.35 | preference shares, importance rankings |
| individual-level | T ≥ 8 and $\bar R^2_Z$ ≥ 0.55 | respondent-specific quantities |
| ratio (MRS / WTP) | T ≥ 10 and $\bar R^2_Z$ ≥ 0.55 and N ≥ 5000 | individual MRS / WTP |

The tiers line up with the three sections above: the "mean & aggregate"
tier is the [population averages](#population-averages), and the
"distributional" and "individual-level" tiers gate the two sections that
follow it.

```{r}
#| label: design-diag-example
#| eval: false
diag <- sc_design_diagnostic(fit)
print(diag)
## Top / bottom R^2_Z attributes are also reported -- they show which
## dummies are best-pinned by Z and which rely most on T for recovery.
```

Requires `stage2 != "none"` (it needs the MAP Hessian's posterior
variance). The v0.2.1 estimator is flagged `experimental = TRUE` in print
output because it has not yet been validated against the paper's
1,872-cell simulation grid; @sec-sw shows it in action on the
Saha-Weeks design.
