---
title: "Case study: Saha-Weeks candidate choice"
---

# Case study: Saha-Weeks candidate choice {#sec-case-sw}

This chapter walks through a structural-conjoint analysis of a
candidate-choice experiment whose design follows
@saha_weeks_2022. Respondents evaluate pairs of hypothetical
political candidates varying on five attributes — the candidate's
policy agenda, talent description, whether they have children, their
gender, and their prior elected office — and select one. Saha and
Weeks' substantive interest is in how voters perceive the ambition
of women candidates; we will exploit the structural decomposition
into direction and intensity to bring that question into sharper
relief on the bundled fixture.

::: {.callout-note}
The `sw2022` dataset bundled with `sconjoint` is a **seed-fixed
synthetic fixture** whose column structure and attribute design
mirror the published study. It is intended only as a runnable
example; do not use it to draw substantive inferences about the
Saha-Weeks population. The full replication dataset is available
from the authors.
:::

```{r setup-sw, include=FALSE}
Sys.setenv(OMP_NUM_THREADS = "8")
options(Ncpus = 8L, mc.cores = 8L)
set.seed(20260409)
if (requireNamespace("torch", quietly = TRUE)) {
  torch::torch_manual_seed(20260409)
}
library(sconjoint)
```

## The bundled data

```{r sw-head}
data(sw2022, package = "sconjoint")
head(sw2022, 4)
dim(sw2022)
table(table(sw2022$respondent))  # tasks per respondent (x2 profiles)
```

Each respondent contributes three forced-choice tasks; within a
task, two candidate profiles are presented and exactly one is
chosen (`choice == 1`). The respondent-level moderators available
in $\mathbf Z$ are `resp_female`, `age` (standardized), and `pid`
(a 3-level partisanship index).

## Fitting the structural model

We fit a small, fast configuration (`K = 3` folds, 60 Adam epochs)
that is suitable for pedagogy. Production analyses would use more
folds and more epochs.

```{r sw-fit}
fit_sw <- scfit(
  choice ~ agenda + talent + children + cand_gender + prior_office |
           resp_female + age + pid,
  data        = sw2022,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 20260409
)
summary(fit_sw)
```

## Attribute importance

Which attributes drive choice, on average? `sc_importance()` returns
the respondent-averaged share of predicted-utility variance
contributed by each attribute block.

```{r sw-importance}
imp_sw <- sc_importance(fit_sw)
imp_sw
```

## Direction vs intensity for candidate gender

The structural decomposition separates *direction*
($\overline{\operatorname{sign}(\beta_k)}$) from *intensity*
($\overline{|\beta_k|}$). For `cand_gender` a near-zero AMCE could
either mean "no one cares" or "half the sample strongly prefers
women, half strongly prefers men and they cancel" — the
direction/intensity decomposition resolves the ambiguity.

```{r sw-di}
di_sw <- sc_direction_intensity(fit_sw)
di_sw
```

## Fraction preferring women

A complementary summary is the fraction of respondents whose
$\hat\beta_{\mathrm{cand\_genderfemale}}(Z_i)$ exceeds zero —
i.e. who, holding everything else fixed, prefer a female candidate.

```{r sw-fraction}
frac_sw <- sc_fraction_preferring(fit_sw, threshold = 0)
frac_sw$estimate[, c("dummy_name", "frac_positive", "frac_negative")]
```

## Subgroup analysis: women vs men respondents

Do women respondents weight candidate gender differently from men?
We can use `sc_subgroup()` to re-average $\hat\beta(Z_i)$ on the
subset of female respondents.

```{r sw-subgroup}
N <- nrow(fit_sw$beta_hat)
fem <- fit_sw$Z[, "resp_female"] > 0.5
sub_fem <- sc_subgroup(fit_sw, fem)
sub_male <- sc_subgroup(fit_sw, !fem)
rbind(
  female = sub_fem$estimate[sub_fem$estimate$dummy_name == "cand_genderfemale",
                            c("theta", "se")],
  male   = sub_male$estimate[sub_male$estimate$dummy_name == "cand_genderfemale",
                            c("theta", "se")]
)
```

## Takeaway

On this fixture, partisanship drives most of the heterogeneity in
agenda preferences, while the *direction* of the
`cand_genderfemale` coefficient splits cleanly along respondent
gender lines even when its average AMCE is small. This is the
headline interpretive payoff of switching from AMCE to the
structural framework: one number becomes an entire distribution
over respondents, indexed by observable $\mathbf Z$. When you
apply the same recipe to a real Saha-Weeks replication dataset,
you would expect the substantive direction to remain but the
magnitudes and standard errors to change.
