# Advanced options {#sec-advanced}

```{r}
#| label: setup-advanced
#| include: false
Sys.setenv(TORCH_SERIALIZE_VERSION = "2")
options(width = 78)
library(sconjoint)
```

This chapter is a reference for arguments to `scfit()` and the
`sc_*` quantity functions that most users will never need to change.
Read it when you need to deviate from defaults: alternative Stage-2
estimators, manual tuning of training hyperparameters, parallel
execution, or reproducibility audits. For the catalog of estimands
themselves --- which quantity to use, what it targets, and how its
inference works --- see @sec-estimands.

The four example chapters (Graham-Svolik, Ballard-Rosa, Saha-Weeks,
Bechtel-Scheve) all use the defaults. This chapter explains what those
defaults are and when you might override them.

## Stage 2: choosing the empirical-Bayes refinement

`scfit()` runs a Stage-2 refinement on top of the Stage-1 DNN cross-fit
by default. The Stage-2 output is what `sc_fit$beta_hat` carries and
what every `sc_*` quantity function reads. The DML population estimates
$\hat\theta$ and clustered SEs are *bit-exactly invariant* across
Stage-2 choices on the same `seed` --- they are always computed from
the Stage-1 DNN.

The four `stage2` choices:

| Value | What it does | When to use |
|-------|--------------|-------------|
| `"map_c5"` (default) | Paper EnsC5. Trains a 2nd DNN with `stage2_seed`, averages with the first DNN, runs a Newton MAP solver per respondent with a score-based diagonal prior scaled by $1/5$. | The paper's recommended default. Best individual-$\beta$ correlation under realistic conjoint designs. |
| `"none"` | Skip Stage 2. `beta_hat` is the raw Stage-1 cross-fitted DNN. | Reproduce v0.1 behavior, or when you want the raw DNN view for downstream code that does its own refinement. |
| `"varref"` | Use $0.5 \cdot \mathrm{Var}_i(\hat\beta_{\text{ens},i,k})$ as the diagonal prior variance (floored at `varref_floor`, default $10^{-3}$). Otherwise identical to MAP. | Recommended for continuous-attribute designs (e.g. Ballard-Rosa tax rates); see `varref_floor` below. |
| `"mixed_logit"` | DNN-offset BLUP via `lme4::glmer`. Treats the DNN prediction as a fixed offset and fits respondent-specific random slopes by Laplace approximation. | When you want a likelihood-based BLUP as a robustness check. The paper notes this is more competitive when $T$ is large, $p$ is modest, and $\mathbf{Z}$ already explains most heterogeneity. |

```{r}
#| label: stage2-example
#| eval: false
## v0.1 behavior
fit_dnn <- scfit(formula, data, ..., stage2 = "none")

## Paper EnsC5 default
fit_map <- scfit(formula, data, ..., stage2 = "map_c5")

## Mixed-logit alternative (requires lme4)
fit_ml  <- scfit(formula, data, ..., stage2 = "mixed_logit")
```

The mixed-logit option calls `lme4::glmer()` and is wrapped in a
deterministic fallback: on convergence failure it sets
`stage2_method = "mixed_logit_failed"`, emits a `warning()` with the
underlying glmer message, and sets `beta_hat <- beta_hat_dnn`. Quantity
functions continue to work; they simply read the unrefined DNN matrix.

### Inspecting what Stage 2 produced

```{r}
#| label: stage2-slots
#| eval: false
fit$stage2_method     # one of: none, map_c5, varref, mixed_logit, mixed_logit_failed
fit$beta_hat          # Stage-2-refined task-level betas (what quantities read)
fit$beta_hat_dnn      # Stage-1 single-DNN betas (always present in v0.2)
fit$beta_hat_dnn2     # 2nd-DNN cross-fit (NULL if stage2 = "none")
fit$beta_hat_ens      # ensemble average of dnn + dnn2 (NULL if stage2 = "none")
fit$beta_hat_resp     # respondent-level Stage-2 betas (N_resp x p)
fit$sigma_prior       # diagonal prior variance used in MAP (length p)
fit$sigma_post_diag   # mean posterior variance from MAP Hessian (length p)
fit$stage2_warnings   # character vector of glmer warnings, if any
```

## Reproducibility: `seed` and `stage2_seed`

`scfit()` ships a bit-exact determinism guarantee:

```{r}
#| label: determinism-example
#| eval: false
## All three return identical(beta_hat, ...) output
scfit(..., seed = 42, stage2_seed = 12345L, parallel = FALSE)
scfit(..., seed = 42, stage2_seed = 12345L, parallel = TRUE, n_cores = 2)
scfit(..., seed = 42, stage2_seed = 12345L, parallel = TRUE, n_cores = 4)
```

`seed` controls Stage 1; `stage2_seed` controls the 2nd DNN in the
Stage-2 ensemble. They are independent --- changing `stage2_seed`
does *not* perturb `theta` or `vcov` (those are computed from the
Stage-1 cross-fit alone), but it does shift the MAP `beta_hat`
through the 2nd DNN's contribution to the ensemble. The default
`stage2_seed = 12345L` matches the paper's prototype.

If `seed = NULL` (the default), Stage 1 is non-deterministic; we still
respect `stage2_seed` so the 2nd DNN's behavior remains predictable
given `seed`.

## `normalize_deltaX`: scale-aware MAP regularization

New in v0.2.1. The v0.2 score-based MAP prior is calibrated assuming
$\mathrm{Var}(\Delta X_k) \approx 1$. For factor dummies under typical
randomization this is approximately true. For designs with continuous
attributes on very different scales --- e.g. the Ballard-Rosa tax-rate
brackets in percentage points (0--50) --- the un-standardized prior
becomes loose by the same factor as $\mathrm{Var}(\Delta X_k)$, and
the MAP solver leaves per-respondent estimates with extreme tails.

`normalize_deltaX = TRUE` divides each `deltaX` column by its sample
SD before training / Lambda(Z) / DML / MAP, then un-standardizes all
coefficient-bearing slots at assembly time. The user-facing surface
(`coef(fit)`, `vcov(fit)`, `fit$beta_hat`, etc.) is on the original
units; only the internal computation runs on a common-variance scale.

```{r}
#| label: normalize-example
#| eval: false
## Recommended for designs with continuous attributes on different
## scales (e.g. price in USD + age in years + 0/1 dummies)
fit <- scfit(formula, data, ..., normalize_deltaX = TRUE)

## The SDs used are kept on the fit for diagnostics:
fit$sd_dx
```

Default is `FALSE`, matching the paper's stated runtime (which keeps
attributes on their original units, e.g. tax rates in percentage
points). On continuous-attribute designs the v0.2 score-based prior
can leave per-respondent MAP estimates with extreme tails when
$\mathrm{Var}(\Delta X_k) \gg 1$; pass `normalize_deltaX = TRUE` in
that case. For pure factor-dummy designs the option is a near-no-op.

## Training hyperparameters

`scfit()` exposes the DNN training loop. Defaults are tuned to handle
realistic conjoint sample sizes (1k--10k respondents). The arguments
worth knowing:

| Argument | Default | Effect |
|----------|---------|--------|
| `K` | `10L` | Number of respondent-clustered cross-fitting folds. Higher = less Stage-1 overfit; lower = faster. Below `K = 3` is not recommended. |
| `n_epochs` | `1000L` | Adam epochs per fold (paper v13 default). Tutorial chapters use 200--400 for render speed. Monitor the loss trace via `plot(fit, "loss_trace")`. |
| `learning_rate` | `0.01` | Adam step size. Lower if the loss diverges; higher if convergence is slow. |
| `weight_decay` | `"adaptive"` | L2 penalty on DNN weights, applied through the Adam optimizer. `"adaptive"` uses the paper's v13 rule $K_{adaptive}/NT$ with $K_{adaptive}=15$ if $NT/p<300$ else $25$; pass a fixed numeric (e.g. `1e-4`) to override. The resolved value is reported on `fit$weight_decay_used`. |
| `varref_floor` | `1e-3` | Lower bound on the prior variance when `stage2 = "varref"` (continuous-attribute designs). The default matches the paper's Ballard-Rosa setting; raising it over-shrinks per-respondent estimates. Ignored for other `stage2` choices. |
| `ridge_lambda` | `1e-4` | Ridge penalty on the local information matrix $\Lambda(Z)$ used in DML. Independent from `weight_decay`. |
| `hidden` | `"auto"` | Hidden-layer widths. `"auto"` picks the paper v13 base `c(32L, 32L, 16L)` for any $N \cdot T \ge 2000$ (with a `c(128L, 64L, 64L)` override at $p \ge 40$ and $NT \ge 80{,}000$); you can pass an integer vector directly. |

```{r}
#| label: hyperparam-example
#| eval: false
fit <- scfit(formula, data, ...,
             K = 10L, n_epochs = 1000L,
             learning_rate = 0.005, weight_decay = 5e-4,
             ridge_lambda = 1e-3,
             hidden = c(128L, 64L, 32L))
```

## Parallel execution

Cross-fitting parallelizes across folds. Use `parallel = TRUE` and set
`n_cores` to the number of physical cores you want to use. The
bit-exact determinism guarantee holds across worker counts (folds are
seeded by `(seed, fold_id)`, never by which worker picks them up).

```{r}
#| label: parallel-example
#| eval: false
fit <- scfit(formula, data, ...,
             parallel = TRUE, n_cores = 4L,
             seed = 42, stage2_seed = 12345L)
```

The bit-exact guarantee is CPU-only. `device = "cuda"` runs the DNN
on a GPU and is faster on large designs but is not deterministic.

## `keep_modules`: forward-pass prediction on new respondents

By default, `scfit()` retains the trained per-fold `torch::nn_module`
objects on the returned `sc_fit`, which enables forward-pass
prediction on new $\mathbf Z$:

```{r}
#| label: keep-modules-example
#| eval: false
fit <- scfit(formula, data, ..., keep_modules = TRUE)  # default

## Forward-pass on new respondents
new_Z <- matrix(stats::rnorm(50 * length(fit$z_names)),
                nrow = 50, ncol = length(fit$z_names))
beta_new <- predict(fit, newdata = new_Z, type = "beta")
```

Set `keep_modules = FALSE` to halve the in-memory size of the fit when
forward-pass prediction is not needed.

## When to override defaults: a quick decision guide

| Situation | Default | Override |
|-----------|---------|----------|
| Routine paper-quality conjoint analysis | use defaults | none |
| Code targeting a specific published v0.1 reproduction | `stage2 = "map_c5"` | `stage2 = "none"` |
| Comparing the MAP refinement to a likelihood-based alternative | use defaults | also run `stage2 = "mixed_logit"`; check `stage2_warnings` |
| Sample very small ($N < 100$ respondents) | `K = 10L` | drop to `K = 3L` (and accept higher Stage-1 variance) |
| Need GPU speedup | `device = "cpu"` | `device = "cuda"` (gives up bit-exact determinism) |
| Distributional claims about polarization or fraction-preferring | `which_beta = "hybrid"` | sanity-check with `which_beta = "dnn"`; if results flip qualitatively, design is too sparse for individual-level claims |
| Forward-pass on out-of-sample $\mathbf Z$ | `keep_modules = TRUE` | none --- the default does what you want |
