---
title: "Case study: Graham-Svolik democratic norms"
---

# Case study: Graham-Svolik democratic norms {#sec-case-gs}

This chapter reanalyses a candidate-choice experiment whose design
follows @graham_svolik_2020. Respondents choose between pairs of
hypothetical U.S. candidates; one attribute captures the candidate's
willingness to engage in undemocratic behavior (ignoring a court
ruling, restricting the opposition press, or banning an opposition
rally). The original study asks how much voters penalize
co-partisan candidates for violating democratic norms, and whether
strong partisanship leads to tolerance of such violations.

::: {.callout-note}
The `gs2020` dataset bundled with `sconjoint` is a **seed-fixed
synthetic fixture** whose column structure mirrors the published
design. It is intended only as a runnable example.
:::

```{r setup-gs, include=FALSE}
Sys.setenv(OMP_NUM_THREADS = "8")
options(Ncpus = 8L, mc.cores = 8L)
set.seed(20260409)
if (requireNamespace("torch", quietly = TRUE)) {
  torch::torch_manual_seed(20260409)
}
library(sconjoint)
```

## The bundled data

```{r gs-head}
data(gs2020, package = "sconjoint")
head(gs2020, 4)
dim(gs2020)
```

Each respondent sees eight tasks with two candidate profiles each,
yielding 16 rows per respondent. The respondent-level moderators
are `resp_pid` (-1/0/+1), `resp_ideo` (standardized), and
standardized `age`.

## Fit

```{r gs-fit}
fit_gs <- scfit(
  choice ~ federal + immigration + tax + abortion + undem + cand_party |
           resp_pid + resp_ideo + age,
  data        = gs2020,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 20260409
)
summary(fit_gs)
```

## Heterogeneity test

Does the per-attribute dispersion of $\hat\beta_k(Z_i)$ exceed what
sampling noise alone would explain?

```{r gs-het}
het_gs <- sc_heterogeneity_test(fit_gs, adjust = "bh")
het_gs$estimate
```

## Direction and intensity for undemocratic behavior

```{r gs-di}
di_gs <- sc_direction_intensity(fit_gs)
di_gs
```

The undemocratic levels should show a large negative average
direction (voters penalize them) coupled with high intensity; the
question the structural model asks is whether the *dispersion* of
those coefficients concentrates on co-partisans, as Graham and
Svolik argue.

## Subgroup analysis by partisanship

We re-average the held-out $\hat\beta(Z_i)$ separately for
Democrats, Independents, and Republicans using `sc_subgroup()`.
The `resp_pid` column on the standardized $\mathbf Z$ matrix
preserves the sign of the underlying -1/0/+1 encoding, so a
simple threshold recovers the partition.

```{r gs-subgroup}
pid_col <- fit_gs$Z[, "resp_pid"]
sub <- sc_subgroup(fit_gs, list(
  dem = pid_col < -0.5,
  ind = abs(pid_col) <= 0.5,
  rep = pid_col >  0.5
))
do.call(rbind, lapply(names(sub), function(g) {
  e <- sub[[g]]$estimate
  rows <- grepl("^undem", e$dummy_name)
  data.frame(group = g, e[rows, c("dummy_name", "theta", "se")],
             row.names = NULL)
}))
```

## Preference clustering

A complementary view on latent preference types is to run k-means
on the per-row $\hat\beta(Z_i)$ matrix.

```{r gs-clusters}
cl_gs <- sc_clusters(fit_gs, k = 3L, seed = 20260409)
cl_gs$estimate$sizes
round(cl_gs$estimate$centers, 2)
```

## Takeaway

Even on 200 synthetic respondents and 8 tasks, the structural fit
cleanly separates the (strong, negative) average disutility of
undemocratic behavior from the (partisan-driven) heterogeneity
around it. The ability to read subgroup-specific $\hat\theta$ off
the same fit — without re-estimating on subsets — is the
operational advantage of the structural model over stacked
subgroup AMCE regressions.
