---
title: "Case study: Bechtel-Scheve climate treaties (WTP)"
---

# Case study: Bechtel-Scheve climate treaties (WTP) {#sec-case-bs}

This chapter walks through a climate-treaty conjoint whose design
follows @bechtel_scheve_2013. Respondents evaluate pairs of
hypothetical international climate agreements varying on six
attributes: burden-sharing principle, enforcement, monitoring,
minimum participation, sanctions, and a numeric monthly cost per
household in U.S. dollars. The headline quantity for this study is
the respondent-level willingness to pay (WTP) for substantive
treaty features, which the structural model delivers directly via
`sc_wtp()` — the WTP is (minus) the marginal rate of substitution
between a benefit attribute and the numeric cost attribute.

::: {.callout-note}
The `bs2013` dataset bundled with `sconjoint` is a **seed-fixed
synthetic fixture** whose column structure mirrors the published
design. The cost attribute is drawn from a six-point dollar grid;
numerical WTP estimates here are illustrative only.
:::

```{r setup-bs, include=FALSE}
Sys.setenv(OMP_NUM_THREADS = "8")
options(Ncpus = 8L, mc.cores = 8L)
set.seed(20260409)
if (requireNamespace("torch", quietly = TRUE)) {
  torch::torch_manual_seed(20260409)
}
library(sconjoint)
```

## The bundled data

```{r bs-head}
data(bs2013)
head(bs2013, 4)
dim(bs2013)
range(bs2013$cost_usd)
```

## Fit

```{r bs-fit}
fit_bs <- scfit(
  choice ~ distribution + enforcement + monitoring +
           participation + sanctions + cost_usd |
           resp_female + age + resp_ideo,
  data        = bs2013,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 3L,
  n_epochs    = 60L,
  seed        = 20260409
)
summary(fit_bs)
```

## Attribute importance

```{r bs-importance}
imp_bs <- sc_importance(fit_bs)
imp_bs
```

## Willingness to pay

What is the average monthly dollar WTP for ambitious participation
(90% of countries joining, relative to the 50% reference) and for
international monitoring (relative to no monitoring)?

```{r bs-wtp}
wtp_part <- sc_wtp(fit_bs,
                   attr = "participation90pct",
                   cost_attr = "cost_usd")
wtp_mon  <- sc_wtp(fit_bs,
                   attr = "monitoringinternational",
                   cost_attr = "cost_usd")
rbind(
  participation_90 = c(wtp_part$estimate, wtp_part$se),
  monitoring_intl  = c(wtp_mon$estimate,  wtp_mon$se)
)
```

Units: USD per household per month. Positive values mean the
average respondent is willing to pay the listed amount to obtain
that feature.

## Compensating differential

`sc_compensating()` answers a complementary question: for each
respondent, how much of the cost attribute would offset a one-unit
gain in a benefit attribute on the utility scale? Here we look at
whether the `participation90pct` benefit is offset by cost for a
typical respondent, and report the share of respondents for whom
`beta_benefit + beta_cost >= 0`.

```{r bs-compensating}
comp <- sc_compensating(fit_bs,
                        benefit = "participation90pct",
                        cost    = "cost_usd")
comp
comp$details$frac_compensated
```

## Takeaway

The Bechtel-Scheve design is the natural home for WTP analysis in
this package: the numeric cost attribute lets `sc_wtp()` report
dollar-scale trade-offs directly, and `sc_compensating()` puts a
per-respondent distribution around the average. Applied to the
full published data one would expect high-income and ideologically
liberal respondents to carry most of the aggregate WTP; here the
heterogeneity is a pedagogical artefact of the synthetic DGP.
