# Integrated mixed logit {#sec-mixed-logit}

::: {.callout-warning}
## Experimental feature

`scmix()` lives on the `feat/mixed-logit` branch and its interface may
change. The estimator, its scores, and the simulation evidence behind it
are documented in the memo of 2026-08-10; this chapter shows how to run
it and how to read its output.
:::

The estimator in the rest of this book fits preferences in two stages: a
deep-network mean stage followed by a respondent-level refinement with a
calibrated prior. Its first stage targets the conditional-logit
projection, which understates coefficient magnitudes whenever residual
preference heterogeneity is present. `scmix()` removes that problem at
the source. It writes preferences as

$$
\beta_i = \mu(Z_i) + A u_i, \qquad u_i \sim N(0, I_q),
$$

with a low-dimensional residual factor $u_i$, and it estimates the mean
network $\mu(\cdot)$ and the loading matrix $A$ jointly through the
respondent-sequence marginal likelihood. Because the heterogeneity is
inside the objective, $\mu(z)$ estimates $\mathbb{E}[\beta_i \mid Z_i = z]$
directly, and none of the prior-calibration devices of the two-stage
pipeline is needed.

Three quantities come with debiased inference: the population mean
`scmix_theta()`, the sign shares `scmix_polarization()`, and the
heterogeneity-integrated vote share `scmix_counterfactual()`. The
latter two have no debiased counterpart under the two-stage estimator.

## A synthetic example with known truth

We simulate a small conjoint in which preferences have both a
covariate-driven part and genuine residual heterogeneity, so the true
population means are known and the projection bias is visible.

```{r}
#| label: mix-sim-data
#| cache: true
library(sconjoint)
set.seed(21)
N <- 1500; T_i <- 8
z <- matrix(runif(N * 2, -1, 1), N, 2)
mu_true <- cbind(0.8 + 0.5 * z[, 1],       # attribute a
                 -0.9 + 0.4 * z[, 2],      # attribute b
                 0.4)                      # attribute c (homogeneous mean)
A_true <- c(0.9, 0.5, 0)                   # residual SDs by attribute
beta <- mu_true + rnorm(N) %*% t(A_true)

a <- sample(c("no", "yes"), 2 * N * T_i, TRUE)
b <- sample(c("no", "yes"), 2 * N * T_i, TRUE)
cc <- sample(c("no", "yes"), 2 * N * T_i, TRUE)
odd <- seq(1L, 2 * N * T_i, by = 2L)
dx <- cbind((a[odd] == "yes") - (a[odd + 1] == "yes"),
            (b[odd] == "yes") - (b[odd + 1] == "yes"),
            (cc[odd] == "yes") - (cc[odd + 1] == "yes"))
rid <- rep(seq_len(N), each = T_i)
yA <- rbinom(N * T_i, 1, plogis(rowSums(dx * beta[rid, ])))
dat <- data.frame(
  respondent = rep(rid, each = 2), task = rep(rep(seq_len(T_i), N), each = 2),
  profile = rep(1:2, N * T_i), a = a, b = b, c = cc,
  z1 = rep(z[rid, 1], each = 2), z2 = rep(z[rid, 2], each = 2),
  choice = as.vector(rbind(yA, 1L - yA)))
theta_true <- colMeans(mu_true)
round(theta_true, 3)
```

The fit uses the same formula interface as `scfit()`. `q` is the
dimension of the residual factor; `q = 1` is the right starting point
for most designs.

```{r}
#| label: mix-sim-fit
#| cache: true
fit_mix <- scmix(choice ~ a + b + c | z1 + z2, dat,
                 respondent = "respondent", task = "task",
                 profile = "profile", q = 1, seed = 7)
fit_mix
```

The print method reports the estimated residual SDs, the square roots of
the diagonal of the rotation-invariant $AA^\top$. Compare them with the
true values $(0.9, 0.5, 0)$. The debiased population means come from
`scmix_theta()`, and a pooled conditional logit shows the projection
bias the mixed logit avoids:

```{r}
#| label: mix-sim-theta
#| cache: true
th <- scmix_theta(fit_mix, seed = 3)
pooled <- coef(glm(fit_mix$y ~ 0 + fit_mix$deltaX, family = binomial()))
data.frame(truth = round(theta_true, 3),
           mixed = round(unname(th$estimate), 3),
           se = round(unname(th$se), 3),
           pooled = round(unname(pooled), 3),
           row.names = fit_mix$attr_names)
```

The pooled logit understates the two large coefficients, while the
mixed-logit estimates sit near the truth with valid standard errors.
Sign shares and an integrated vote share follow the same pattern:

```{r}
#| label: mix-sim-shares
#| cache: true
scmix_polarization(fit_mix, seed = 3)
scmix_counterfactual(fit_mix, contrast = c(ayes = 1, byes = -1), seed = 3)
```

`scmix_polarization()` estimates $\Pr(\beta_{ik} > 0)$ under the fitted
mixing law, and `scmix_counterfactual()` integrates the head-to-head
choice probability over the full heterogeneity distribution. Standard
errors for these two quantities treat the estimated loading matrix as
fixed; the memo documents this limitation and the follow-up that
removes it.

## A real-data fit

On the bundled tax-conjoint data the call is identical. Continuous
attributes are handled internally by contrast standardization, so no
rescaling is needed on the user's side.

```{r}
#| label: mix-br-fit
#| cache: true
data(br2017)
fit_br <- scmix(
  choice ~ rate_L10 + rate_10_35 + rate_35_85 + rate_85_175 +
    rate_175_375 + rate_375P + revenue_score |
    age_std + female + pid7_std + educ_std + race_white + income_std +
    ineq_averse + work_vs_luck + taxes_harm_econ + hardwork +
    high_econ_know + employed_ft + conserv_ideo + govt_serv + newsint +
    numeracy + gen_mobile + future_mobile + gov_assist + risk_averse +
    hardship + children + trust,
  br2017, respondent = "respondent", task = "task", profile = "profile",
  q = 1, seed = 42)
round(scmix_theta(fit_br, seed = 7)$estimate, 4)
```

The estimates agree with the two-stage results of @sec-br within
one standard error. What changes is the interpretation and the residual
model: the coefficients now estimate latent population means rather than
projections, and the residual heterogeneity is estimated by the
likelihood rather than set by a calibrated prior. The fitted loading
concentrates the heterogeneity on the revenue score rather than the tax
rates, which the two-stage prior spread uniformly.

## When to use which estimator

The two-stage estimator remains the default for reproducing the paper's
published results, and it is robust to the shape of the residual
distribution because its probability-scale quantities do not depend on
it. `scmix()` is the right tool when the latent-scale magnitudes, the
sign shares, or the integrated vote shares are the quantities of
interest, at the price of a Gaussian assumption on the residual factor.
At $T = 3$ the loading matrix is weakly identified and the fold-to-fold
spread reported by the fit should be checked before any distributional
quantity is taken seriously.
