---
title: "Getting started with sconjoint"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with sconjoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`sconjoint` implements the structural deep-learning estimator of
Acharya, Hainmueller, and Xu (2026) for forced-choice conjoint
experiments.
This vignette walks through a minimal example; for a full tutorial see
the [online Quarto book](https://xuyiqing.github.io/sconjoint/).

## Setup

```{r setup}
library(sconjoint)
data(sw2022)
```

`sw2022` is a small synthetic candidate-choice conjoint dataset
(200 respondents, 3 tasks, 5 attributes, 3 moderators).

## Fit the model

```{r fit}
fit <- scfit(
  choice ~ agenda + talent + children + cand_gender + prior_office |
           resp_female + age + pid,
  data        = sw2022,
  respondent  = "respondent",
  task        = "task",
  profile     = "profile",
  K           = 2L,
  n_epochs    = 20L,
  seed        = 42L,
  verbose     = FALSE
)
```

## Inspect results

```{r summary}
summary(fit)
```

```{r coef}
head(coef(fit))
```

## Structural quantity: attribute importance

```{r importance}
imp <- sc_importance(fit)
imp
```

See `?sc_importance` and the companion quantities (`sc_mrs`, `sc_wtp`,
`sc_counterfactual`, `sc_polarization`, etc.) for the full toolkit.
The Quarto book provides detailed case studies with each quantity.
