---
title: "Getting started with sconjoint"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with sconjoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Overview

`sconjoint` implements the structural deep-learning estimator of
Acharya, Hainmueller, and Xu (2026) for forced-choice conjoint
experiments. The package is under active development.

This vignette will grow into a getting-started tour. At present
(milestone M1) the package ships only the skeleton; statistical
methods will land in subsequent releases.

```{r setup}
library(sconjoint)
```

# Coming soon

- `scfit()`: fit the structural DNN model with DML inference.
- A focused toolkit of structural quantities (MRS, counterfactuals,
  attribute importance, preference polarization, consumer surplus).
- Worked case studies bundled as small example datasets.

See the project README for the development roadmap.
