---
title: "Simulation sanity check"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation sanity check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette runs a quick synthetic-DGP sanity check to verify that
`scfit()` recovers the direction of known structural parameters.
It is a smoke test, not a formal simulation study.

## Generate synthetic data

```{r dgp}
set.seed(101)
N_resp <- 50L
N_task <- 3L

resp <- data.frame(
  respondent = seq_len(N_resp),
  z1         = rnorm(N_resp)
)

rows <- expand.grid(
  respondent = seq_len(N_resp),
  task       = seq_len(N_task),
  profile    = 1:2
)
rows <- merge(rows, resp, by = "respondent")

## Two binary attributes with uniform randomisation
rows$x1 <- sample(0:1, nrow(rows), replace = TRUE)
rows$x2 <- sample(0:1, nrow(rows), replace = TRUE)

## True beta(z): linear in moderator z1
## beta1(z) =  0.8 + 0.5 * z1   (positive main effect)
## beta2(z) = -0.6 + 0.3 * z1   (negative main effect)
beta1 <- 0.8 + 0.5 * rows$z1
beta2 <- -0.6 + 0.3 * rows$z1

logit <- beta1 * rows$x1 + beta2 * rows$x2
prob  <- 1 / (1 + exp(-logit))

## Generate choices within each task (exactly one chosen per pair)
rows$choice <- 0L
task_ids <- paste(rows$respondent, rows$task, sep = "_")
for (tid in unique(task_ids)) {
  idx <- which(task_ids == tid)
  p <- prob[idx]
  ## softmax across the two profiles
  ep <- exp(logit[idx])
  pr <- ep / sum(ep)
  chosen <- sample(idx, 1, prob = pr)
  rows$choice[chosen] <- 1L
}

## Convert attributes to factors for scfit
rows$x1 <- factor(rows$x1, levels = 0:1, labels = c("a", "b"))
rows$x2 <- factor(rows$x2, levels = 0:1, labels = c("lo", "hi"))
```

## Fit the model

```{r fit}
library(sconjoint)

fit <- scfit(
  choice ~ x1 + x2 | z1,
  data       = rows,
  respondent = "respondent",
  task       = "task",
  profile    = "profile",
  K          = 2L,
  n_epochs   = 20L,
  seed       = 101L,
  verbose    = FALSE
)
```

## Check recovered signs

The true main effects are positive for `x1` and negative for `x2`.
We verify the average theta direction matches.

```{r check}
theta <- coef(fit)
cat("Recovered average theta:\n")
print(colMeans(theta))
```

With only 20 epochs on 50 respondents the magnitudes will be noisy,
but the sign pattern should broadly align with the DGP.
A full Monte Carlo study with proper coverage analysis is beyond the
scope of this smoke test.
